### 2024-04-15 R/simulation_study/simulation_study_fjords.Rmd
Commit Latex and simulation study changes

---
title: "Simulation study in fjords"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
date: "2024-03-13"
---

# Import libraries and data

```{r import libraries}
#seed for reproducibility 
set.seed(1)

#for tex code
library(tinytex)

#for tex expression in plot legends
library(latex2exp)

#for pipeline operator
library(dplyr)

#plots
library(ggplot2)
library(plotly)
library(htmlwidgets)

#data managing
library(tidyr)

#progress bar
library(progress)

#for land polygons
library(sf)

#SDE fit
library(smoothSDE)

#to fit smooth function with tensor splines
library(mgcv)

#auxiliary function to simulate from RACVM with specified smooth function for omega
source("CVM_functions.R")
```

# Get land polygons

```{r get land polygons}
#land polygons

# Set the path to the directory containing the data
par_dir=dirname(dirname(getwd())) #parent directory 
greenland_data_path <- file.path(par_dir,"Data", "Greenland")

#get the coastline geometry from the geojson file
land<-st_read(file.path(greenland_data_path,"land_utm.shp"))
```

# SIMULATION

## Sample initial positions uniformly

```{r sample initial positions}

#number of tracks
n_samples=5

#initial velocity and location
v0=c(0,0)

#generate random initial points
x0=matrix(rep(NA,n_samples*2),ncol=2)
colnames(x0)=c("x1","x2")

i=1
while (i<=n_samples) {
  #choose location uniformly in the map
  x=c(runif(1,min=430,max=500),runif(1,min=7760,max=7900))
  p=nearest_shore_point(st_point(x),land)
  Dshore=(x[1]-p[1])^2+(x[2]-p[2])^2
  #keep it as initial position if it is at least 50 metres away from the shore
  if (Dshore>0.5) {
    x0[i,]=x
    i=i+1
  }
}
```

## Fix time steps for observations

```{r fix time steps}
#observation times
Tmax=3*24 #max time in hour

hf=1/60 #1 minute time step

#high frequency time steps
times_hf=seq(0,Tmax,by=hf)
n_hf=length(times_hf)-1

```

## Set observation error

```{r set observation error}
#no measurement error (we cannot put 0 because only the log of this quantity is used)
sigma_obs_0=0.003
H0_hf=array(rep(sigma_obs_0^2*diag(2),n_hf),dim=c(2,2,n_hf))
#GPS measurement error 
sigma_obs_1=0.03
H1_hf=array(rep(sigma_obs_1^2*diag(2),n_hf),dim=c(2,2,n_hf))

```

## Smooth parameters

### Define smooth functions for omega and tau

```{r define smooth parameters}

ftau=function(cov_data,D0=0.3,delta0=0.2,tau0=1,tau1=2,tau2=2,lambda=4,kappa=0.1) {
  Dshore=cov_data$DistanceShore
  theta=cov_data$theta
  coeff=exp(-kappa*(Dshore/D0)^2)
  tau=tau0*(1-coeff)+(delta0+(tau0-delta0)*exp(-theta^2/pi^2)+(tau1-delta0)*exp(-lambda*(theta-pi/2)^2)+(tau2-delta0)*exp(-lambda*(theta+pi/2)^2))*coeff
  return (tau)
}

#mixture of independent gaussian
ftau_gauss=function(cov_data,D0=0.3,tau0=1,tau1=3) {
  Dshore=cov_data$DistanceShore
  theta=cov_data$theta
  tau=tau0+(tau1-tau0)*exp(-((theta-pi/2)/(pi/10))^2-(Dshore/D0)^2)+(tau1-tau0)*exp(-((theta+pi/2)/(pi/10))^2-(Dshore/D0)^2)
  return (tau)
}

#constant tau
ftau_constant=function(Dshore,theta,tau=1) {
  return (tau)
}

fomega=function(cov_data,D0=0.3,omega0=60*pi/2,lambda=2,kappa=0.2) {
    Dshore=cov_data$DistanceShore
    theta=cov_data$theta
    if (is.null(Dshore)){
      Dshore=1/cov_data$ExpShore
    }
    coeff=exp(-kappa*(Dshore/D0)^2)
    omega=omega0/2*(tanh(lambda*(theta+pi/2-atanh(-0.9)/lambda))+tanh(lambda*(theta-(pi/2-atanh(-0.9)/lambda))))*coeff
    return(omega)
}

nu=4

```

### Plot smooth functions.

```{r plot smooth parameters }

Dshore=seq(from=0.1,to=2,length.out=30)
theta=seq(from=-pi,to=pi,length.out=30)
grid<- as.data.frame(expand.grid(Dshore,theta))
colnames(grid)=c("DistanceShore","theta")
true_par=grid
true_par$tau=ftau_constant(grid)
true_par$omega=fomega(grid)

plot_true_tau <- plot_ly(true_par,type = "mesh3d",
                         x = ~theta,y = ~DistanceShore,z=~tau,intensity=~tau,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("True tau"),scene=list(xaxis = list(title = "Theta",showgrid = F),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "tau")))

plot_true_omega <- plot_ly(true_par,type = "mesh3d",
                         x = ~theta,y = ~DistanceShore,z=~omega,intensity=~omega,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("True omega"),scene=list(xaxis = list(title = "Theta",showgrid = F),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "omega")))

plot_true_tau
plot_true_omega
```

### Approximate smooth functions with tensor splines of DistanceShore

```{r}
sp_df=5
#number of points
n <- 500

#sample theta and DistanceShore points in the domain
theta <- runif(n,-pi,pi);
DistanceShore <- runif(n,0.2,2);
samples=data.frame(theta=theta,DistanceShore=DistanceShore)

#define grid of values
theta_v <- seq(-pi,pi,length=30)
Dshore_v<- seq(0.2,2,length=30)
pr <- data.frame(theta=rep(theta_v,30),DistanceShore=rep(Dshore_v,rep(30,30)))

# true values of the function over this grid
truth <- matrix(fomega(pr),30,30)

#points on the surface perturbed by gaussian noise
f <- fomega(samples)
y <- f+0*rnorm(n)

#plot true function
persp(theta_v,Dshore_v,truth);
title("truth")

#fit with bivariate splines te
m1 <- gam(y~te(theta,DistanceShore,k=sp_df))

#visualize
vis.gam(m1);title("tensor product")

df=data.frame("x"=theta,"z"=DistanceShore,"y"=y)
fig <- plot_ly(df,type="scatter3d", x = ~x, y = ~z, z = ~y,mode="markers",marker=list(size=4))
fig
```

### Approximate smooth functions with tensor splines of ExpShore

```{r}
ExpShore=1/DistanceShore

#fit with bivariate splines te
m2 <- gam(y~te(theta,ExpShore,k=sp_df))

#visualize
vis.gam(m2);title("tensor product")

```

### Get estimated spline coefficients

```{r}
sp_coeff_Dshore=m1$coefficients
sp_coeff_ExpShore=m2$coefficients
```

## Simulation with measurement errors and high frequency

### Generate samples

```{r high frequency samples no error (hf ne)}

#initialize data frame
data_fjords_hf_ne=data.frame("y1"=NA,"y2"=NA,"time"=NA,"p1"=NA,"p2"=NA,"ID"=NA)

for (i in 1:n_samples) {

  res=sim_theta_CRCVM(ftau_constant,fomega,nu,log_sigma_obs=log(sigma_obs_0),v0,x0[i,],times_hf,land=land,verbose=FALSE)

  data_sim=res$sim
  data_sim$ID=factor(rep(i,length(data_sim$y1)))
  data_shore=res$shore[,c("p1","p2")]
  data_fjords_hf_ne=rbind(data_fjords_hf_ne,cbind(data_sim,data_shore))
}

data_fjords_hf_ne=data_fjords_hf_ne[-1,]

```

### percentage of samples that reached land

```{r percentage of hf ne samples that reached land}
count=0
for (id in unique(data_fjords_hf_ne$ID)){
  sub_data=data_fjords_hf_ne[data_fjords_hf_ne$ID==id,]
  if (nrow(sub_data) < n_hf) {
    print(id)
    count=count+1
  }
}
cat(count/n_samples*100,"percent of the samples reached land")
```

### Plot samples

```{r plot hf ne samples}

plot_samples_fjords_hf_ne=ggplot()+geom_sf(data=land$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_fjords_hf_ne[data_fjords_hf_ne$ID==2,],mapping=aes(y1,y2),size=0.1)+
  geom_point(data=data_fjords_hf_ne[data_fjords_hf_ne$ID==2,],mapping=aes(p1,p2),size=0.01,col="red",shape=5)+
  geom_path(data=data_fjords_hf_ne[data_fjords_hf_ne$ID==2,],mapping=aes(y1,y2),linewidth=0.1)+
  geom_point(data = data_fjords_hf_ne[data_fjords_hf_ne$ID==2,] %>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4,col = "blue")+
  xlab("x") + ylab("y")

ggsave(filename="samples_fjords_hf_ne_omega.png",plot=plot_samples_fjords_hf_ne)
ggplotly(plot_samples_fjords_hf_ne)
```

## sImulate with high frequency and measurement error

### Generate samples

```{r generate high frequency samples with error (hf e)}

#initialize data frame
data_fjords_hf_e=data.frame("y1"=NA,"y2"=NA,"time"=NA,"p1"=NA,"p2"=NA,"ID"=NA)

for (i in 1:n_samples) {

  res=sim_theta_CRCVM(ftau_constant,fomega,nu,log_sigma_obs=log(sigma_obs_1),v0,x0[i,],times_hf,land=land,verbose=FALSE)

  data_sim=res$sim
  data_sim$ID=factor(rep(i,length(data_sim$y1)))
  data_shore=res$shore[,c("p1","p2")]
  data_fjords_hf_e=rbind(data_fjords_hf_e,cbind(data_sim,data_shore))
}

data_fjords_hf_e=data_fjords_hf_e[-1,]

```

### Percentage of samples that reached land
```{r percentage of hf e samples that reached land}
count=0
for (id in unique(data_fjords_hf_e$ID)){
  sub_data=data_fjords_hf_e[data_fjords_hf_e$ID==id,]
  if (nrow(sub_data) < n_hf) {
    print(id)
    count=count+1
  }
}
cat(count/n_samples*100,"percent of the samples reached land")
```

### Plot one sample

```{r plot hf e samples}

plot_samples_fjords_hf_e=ggplot()+geom_sf(data=land$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_fjords_hf_e[data_fjords_hf_e$ID==2,],mapping=aes(y1,y2),size=0.1)+
   geom_path(data=data_fjords_hf_e[data_fjords_hf_e$ID==2,],mapping=aes(y1,y2),linewidth=0.1)+
  geom_point(data = data_fjords_hf_e[data_fjords_hf_e$ID==2,] %>% filter(!duplicated(ID)),
             aes(x = y1, y = y2 ), shape = 3, size = 4,col = "red")+
    xlab("x") + ylab("y")

ggsave(filename="samples_fjords_hf_e_omega.png",plot=plot_samples_fjords_hf_e)
ggplotly(plot_samples_fjords_hf_e)
```

# ESTIMATION

## Compute distance to shore and empirical deviation angles

```{r compute covariates}

signed_angle <- function(u, v) {
    #Compute signed angle in [-pi,pi] that rotates first vector into second vector 
    # as in 
    # https://math.stackexchange.com/questions/529555/signed-angle-between-2-vectors
    u <- matrix(u, ncol = 2)
    v <- matrix(v, ncol = 2)
    if (nrow(u) != nrow(v)) stop("u and v must have the same number of 
                                  rows")
    result <- as.numeric(atan2(v[,2], v[,1]) - atan2(u[,2], u[,1]))
    ind1 <- which(result > pi)
    ind2 <- which(result <= -pi)
    result[ind1] <- result[ind1] - 2*pi
    result[ind2] <- result[ind2] + 2*pi
    return(result) 
}

add_covs=function(data) {
  #data has at least columns "time","y1" "y2", "p1","p2"
  #this function returns a dataframe with three mùore columns : "theta", "DistanceShore" and "ExpShore"

  new_data=data
  #observed deviation angles
  n=length(new_data$time)

  new_data$theta=rep(NA,n)

  for (id in unique(new_data$ID)) {

    #filter data with specific id
    sub_ind=(new_data$ID==id)
    sub_data=new_data[sub_ind,]
    n_sub=length(sub_data$time)

    #time steps
    dtimes=sub_data[2:n_sub,"time"]-sub_data[1:(n_sub-1),"time"]

    #step lengths
    dx=sub_data[2:n_sub,"y1"]-sub_data[1:(n_sub-1),"y1"]
    dy=sub_data[2:n_sub,"y2"]-sub_data[1:(n_sub-1),"y2"]

    #matrix of empirical velocity
    vexp_df=cbind(dx/dtimes,dy/dtimes)

    #matrix of normal vectors
    normal=as.matrix(sub_data[2:n_sub,c("y1","y2")]-sub_data[2:n_sub,c("p1","p2")])

    #angle between velocity and normal vector
    theta_coast=signed_angle(normal,vexp_df)

    #adjust lengths 
    theta_coast=c(theta_coast,1)

    new_data[sub_ind,"theta"]=theta_coast
  }

  #distance to shore
  Dshore=sqrt((new_data$y1-new_data$p1)^2+(new_data$y2-new_data$p2)^2)
  new_data$DistanceShore=Dshore
  new_data$ExpShore=1/Dshore

  return(new_data)
}

data_fjords_hf_ne=add_covs(data_fjords_hf_ne)

data_fjords_hf_e=add_covs(data_fjords_hf_e)
```

## Estimate from high frequency data without error : ExpShore and tensor splines ti

### Use ExpShore covariate and tensor splines te

#### Fit sde

```{r estimate from hf e samples with cov ExpShore}

#estimation with tensor splines
formulas <- list(mu1=~1,mu2=~1,tau=~1,
                 nu=~1,omega=~te(theta,ExpShore,k=5,bs="cs"))
par0 <- c(0,0,1,4,0)

crcvm_fjords_hf_ne1 <- SDE$new(formulas = formulas,data = data_fjords_hf_ne[data_fjords_hf_ne$ID==2,],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2","nu","tau"),
                      other_data=list("H"=H0_hf))
crcvm_fjords_hf_ne1$fit(method="BFGS")
```

#### Plot estimated smooth parameter omega in 3D

```{r get plots from hf e estimates with Expshore}
xmin=list("theta"=-pi,"ExpShore"=1/2)
xmax=list("theta"=pi,"ExpShore"=1/0.3)
xlabel=list("ExpShore"="Distance to shore")
link=list("ExpShore"=(\(x) 1/x))
true_smooths=list("omega"=fomega)
res=crcvm_fjords_hf_ne1$get_all_plots(baseline=NULL,model_name="crcvm_fjords_hf_ne1",xmin=xmin,xmax=xmax,
                                    xlabel=xlabel,link=link,show_CI="pointwise",true_smooths=true_smooths,save=TRUE)
res$fe_omega_theta_ExpShore
```

#### Violin plot of the estimated spline coefficients

```{r}

true_coeff=data.frame(value=sp_coeff_ExpShore,name=paste("omega.",names(sp_coeff_ExpShore),sep=""))
rownames(true_coeff)=NULL

p=crcvm_fjords_hf_ne1$plot_violin_coeff(model_name="crcvm_fjords_hf_ne1",substring="omega",true_coeff=true_coeff,save=TRUE)

ggplotly(p)
```

#### Simulate from fitted model

```{r,cache=TRUE}

fExpShore=function(z,v,p) {

  #normal vector
  normal=c(z[1]-p[1],z[2]-p[2])

  #distance to shore
  Dshore=sqrt(normal[1]^2+normal[2]^2)
  if (Dshore>0.5){
    return (0)
  }
  else if (Dshore<0.01) {
    return (100)
  }
  else {
    return (1/Dshore)
  }
}

fangle=function(z,v,p) {

  #normal vector
  normal=c(z[1]-p[1],z[2]-p[2])

  #return angle
  return (signed_angle(normal,v))
}

atw=list("ExpShore"=fExpShore,"AngleNormal"=fangle)
sim_data=crcvm_fjords_hf_ne1$simulate(z0=x0[1,],data=crcvm_fjords_hf_ne1$data(),atw=atw,land=land)

```

#### Plot simulation from fitted model

```{r}

plot_sim=ggplot()+geom_sf(data=land$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=sim_data,mapping=aes(y1,y2),size=0.1)+
  geom_path(data=sim_data,mapping=aes(y1,y2),size=0.1)+
  geom_point(data = sim_data%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2 ), shape = 3, size = 4, col = "red")+
    xlab("x") + ylab("y")

ggsave("sim_crcvm_fjords_hf_ne1.png",plot=plot_sim,width=10,height=5,path="crcvm_fjords_hf_ne1")
ggplotly(plot_sim)
```

###  UseDistanceShore covariate and tensor splines ti

#### Fit sde

```{r,cache=TRUE}
#estimation with tensor splines
formulas <- list(mu1=~1,mu2=~1,tau=~1,
                 nu=~1,omega=~te(theta,DistanceShore,k=sp_df,bs="cs"))
par0 <- c(0,0,1,4,0)

crcvm_fjords_hf_ne2 <- SDE$new(formulas = formulas,data = data_rect_hf_ne[data_rect_hf_ne$ID==1,],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2","nu","tau"),
                      other_data=list("H"=H0_hf))
crcvm_fjords_hf_ne2$fit(method="BFGS")
```

#### Plots estimated smooth in 3D

```{r}
xmin=list("theta"=-pi,"DistanceShore"=0.3)
xmax=list("theta"=pi,"DistanceShore"=2)
res=crcvm_fjords_hf_ne2$get_all_plots(baseline=NULL,model_name="crcvm_fjords_hf_ne2",xmin=xmin,xmax=xmax,
                                    xlabel=xlabel,link=link,show_CI="pointwise",true_smooths=true_smooths,save=TRUE)
res$fe_omega_theta_DistanceShore
```

#### Violin plot of the estimated spline coefficients

```{r}

true_coeff=data.frame(value=sp_coeff_Dshore,name=paste("omega.",names(sp_coeff_Dshore),sep=""))
rownames(true_coeff)=NULL

p=crcvm_fjords_hf_ne2$plot_violin_coeff(model_name="crcvm_fjords_hf_ne2",substring="omega",true_coeff=true_coeff,save=TRUE)

ggplotly(p)
```

## Estimate from high frequency data with error :

### Use ExpShore and tensor splines te

#### Fit sde

```{r, cache=TRUE}
#estimation with tensor splines
formulas <- list(mu1=~1,mu2=~1,tau=~1,
                 nu=~1,omega=~te(theta,ExpShore,k=5,bs="cs"))
par0 <- c(0,0,1,4,0)

crcvm_fjords_hf_e1 <- SDE$new(formulas = formulas,data = data_fjords_hf_e[data_fjords_hf_e$ID==1,],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2","nu","tau"),
                      other_data=list("H"=H1_hf))
crcvm_fjords_hf_e1$fit(method="BFGS")
```

#### Plot estimated smooth in 3D

```{r}
xmin=list("theta"=-pi,"ExpShore"=1/2)
xmax=list("theta"=pi,"ExpShore"=1/0.3)
xlabel=list("ExpShore"="Distance to shore")
link=list("ExpShore"=(\(x) 1/x))
true_smooths=list("omega"=fomega)
res=crcvm_fjords_hf_e1$get_all_plots(baseline=NULL,model_name="crcvm_fjords_hf_e1",xmin=xmin,xmax=xmax,
                                    xlabel=xlabel,link=link,show_CI="pointwise",true_smooths=true_smooths,save=TRUE)
res$fe_omega_theta_ExpShore
```

#### Violin plot of the estimated spline coefficients

```{r}

true_coeff=data.frame(value=sp_coeff_ExpShore,name=paste("omega.",names(sp_coeff_ExpShore),sep=""))
rownames(true_coeff)=NULL

p=crcvm_fjords_hf_e1$plot_violin_coeff(model_name="crcvm_fjords_hf_e1",substring="omega",true_coeff=true_coeff,save=TRUE)

ggplotly(p)
```

### Use DistanceShore and tensor splines te

#### Fit sde

```{r,cache=TRUE}
#estimation with tensor splines
formulas <- list(mu1=~1,mu2=~1,tau=~1,
                 nu=~1,omega=~te(theta,DistanceShore,k=sp_df,bs="cs"))
par0 <- c(0,0,1,4,0)

crcvm_fjords_hf_e2 <- SDE$new(formulas = formulas,data = data_fjords_hf_e[data_fjords_hf_e$ID==1,],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2","nu","tau"),
                      other_data=list("H"=H1_hf))
crcvm_fjords_hf_e2$fit(method="BFGS")
```

#### Plot estimated smooth in 3D

```{r}
xmin=list("theta"=-pi,"DistanceShore"=0.3)
xmax=list("theta"=pi,"DistanceShore"=2)
res=crcvm_fjords_hf_e2$get_all_plots(baseline=NULL,model_name="crcvm_fjords_hf_e2",xmin=xmin,xmax=xmax,
                                    xlabel=xlabel,link=link,show_CI="pointwise",true_smooths=true_smooths,save=TRUE)
res$fe_omega_theta_DistanceShore
```

#### Violin plot of the estimated spline coefficients

```{r}

true_coeff=data.frame(value=sp_coeff_Dshore,name=paste("omega.",names(sp_coeff_Dshore),sep=""))
rownames(true_coeff)=NULL

p=crcvm_fjords_hf_e2$plot_violin_coeff(model_name="crcvm_fjords_hf_e2",substring="omega",true_coeff=true_coeff,save=TRUE)

ggplotly(p)
```