### 2024-04-16 R/simulation_study/simulation_study_rect.Rmd
parallelization in simulation_study_rect + other stuff

---
title: "Simulation study in rectangular domain"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
  date: "2024-03-13"
---

# SET UP

## Import libraries

```{r libraries}
#seed for reproducibility 
set.seed(1)

#parallel computing
library(foreach)
library(doParallel)

#for tex code
library(tinytex)

#for tex expression in plot legends
library(latex2exp)

#for pipeline operator
library(dplyr)

#plots
library(ggplot2)
library(plotly)
library(htmlwidgets)

#data managing
library(tidyr)

#progress bar
library(progress)

#for land polygons
library(sf)

#SDE fit
library(smoothSDE)

#to fit smooth function with tensor splines
library(mgcv)

#some auxiliary functions
source("CVM_functions.R")
```

## global variables

```{r global variables}
#number of batch
N_BATCH=10

#number of individual tracks per batch
N_ID=3

#duration of each track in hour
TMAX=1*24

#degree of freedom in tensor splines
SP_DF=5
```

# SIMULATION

## Create the rectangle

We create a square with sides of length 8km.

```{r create rectangle}

#rectangular domain 
large_rectangle_coords <- matrix(c(0, 0, 0, 10, 10, 10, 10, 0, 0, 0), ncol = 2, byrow = TRUE)
small_rectangle_coords <- matrix(c(1,1, 1, 9, 9, 9, 9, 1, 1, 1), ncol = 2, byrow = TRUE)

# Create an sf object representing the rectangles
large_rectangle <- st_sfc(st_polygon(list(large_rectangle_coords)))
large_rectangle<-st_sf(geometry = large_rectangle)

small_rectangle <- st_sfc(st_polygon(list(small_rectangle_coords)))
small_rectangle<-st_sf(geometry = small_rectangle)

#change its crs
domain<- st_sym_difference(large_rectangle, small_rectangle)

```

```{r plot rectangle}

plot_domain=ggplot()+geom_sf(data=domain$geometry,fill="blue")
plot_domain
```

## Sample initial positions uniformly in the domain

```{r initial positions}

#initial velocity and location
v0=c(0,0)

#generate random initial points
x0=matrix(rep(NA,N_ID*2),ncol=2)
colnames(x0)=c("x1","x2")

for (i in 1:N_ID) {
  #choose location uniformly in the domain
  x=c(runif(1,min=1.5,max=8.5),runif(1,min=1.5,max=8.5))
  x0[i,]=x
}

```

## Fix time steps for observations

```{r time steps}
#time step
hf=1/60

#high frequency time points
times_hf=seq(0,TMAX,by=hf)
n_hf=length(times_hf)-1

```

## Set observation error

```{r measurement error}
#no measurement error (we cannot put 0 because only the log of this quantity is used)
sigma_obs_0=0.005
H0_hf=array(rep(sigma_obs_0^2*diag(2),n_hf*N_ID),dim=c(2,2,n_hf*N_ID))
#GPS measurement error 
sigma_obs_1=0.03
H1_hf=array(rep(sigma_obs_1^2*diag(2),n_hf*N_ID),dim=c(2,2,n_hf*N_ID))

```

## Smooth parameters of the CRCVM

### Define smooth functions for tau and omega.

```{r smooth parameters}

#smooth parameters

ftau=function(cov_data,D0=0.3,delta0=0.2,tau0=1,tau1=2,tau2=2,lambda=4,kappa=0.1) {
  Dshore=cov_data$DistanceShore
  theta=cov_data$theta
  coeff=exp(-kappa*(Dshore/D0)^2)
  tau=tau0*(1-coeff)+(delta0+(tau0-delta0)*exp(-theta^2/pi^2)+(tau1-delta0)*exp(-lambda*(theta-pi/2)^2)+(tau2-delta0)*exp(-lambda*(theta+pi/2)^2))*coeff
  return (tau)
}

#mixture of independent gaussian
ftau_gauss=function(cov_data,D0=0.5,tau0=1,tau1=1.5) {
  Dshore=cov_data$DistanceShore
  theta=cov_data$theta
  tau=tau0+(tau1-tau0)*exp(-((theta-pi/2)/(pi/10))^2-(Dshore/D0)^2)+(tau1-tau0)*exp(-((theta+pi/2)/(pi/10))^2-(Dshore/D0)^2)
  return (tau)
}

#constant tau
ftau_constant=function(cov_data,tau=1,sigma_re=0.1) {
  b=rnorm(1,mean=0,sd=sigma_re)
  return (tau+b)
}

fomega=function(cov_data,D0=0.3,omega0=60*pi,lambda=1.5,kappa=0.5) {
    Dshore=cov_data$DistanceShore
    theta=cov_data$theta
    if (is.null(Dshore)){
      Dshore=1/cov_data$ExpShore
    }
    coeff=exp(-kappa*(Dshore/D0)^2)
    omega=omega0/2*(tanh(lambda*(theta+pi/2-atanh(-0.9)/lambda))+tanh(lambda*(theta-(pi/2-atanh(-0.9)/lambda))))*coeff
    return(omega)
}

#constant nu
fnu_constant=function(cov_data,nu=4,sigma_re=0.3) {
    b=rnorm(1,mean=0,sd=sigma_re)
  return (nu+b)
}

```

### Plot smooth functions.

```{r plot smooth parameters}

Dshore=seq(from=0.3,to=2,length.out=30)
theta=seq(from=-pi,to=pi,length.out=30)
grid<- as.data.frame(expand.grid(Dshore,theta))
colnames(grid)=c("DistanceShore","theta")
true_par=grid
true_par$tau=ftau_constant(grid)
true_par$omega=fomega(grid)

plot_true_tau <- plot_ly(true_par,type = "mesh3d",
                         x = ~theta,y = ~DistanceShore,z=~tau,intensity=~tau,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("True tau"),scene=list(xaxis = list(title = "Theta",showgrid = F),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "tau")))

plot_true_omega <- plot_ly(true_par,type = "mesh3d",
                         x = ~theta,y = ~DistanceShore,z=~omega,intensity=~omega,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("True omega"),scene=list(xaxis = list(title = "Theta",showgrid = F),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "omega")))

plot_true_tau
plot_true_omega
```

### Approximate smooth functions with tensor splines of DistanceShore

```{r approximate smooth omega with DistanceShore splines}

#number of points
n <- 1000

#sample theta and DistanceShore points in the domain
theta <- runif(n,-pi,pi);
DistanceShore <- runif(n,0.2,2);
samples=data.frame(theta=theta,DistanceShore=DistanceShore)

#define grid of values
theta_v <- seq(-pi,pi,length=30)
Dshore_v<- seq(0.2,2,length=30)
pr <- data.frame(theta=rep(theta_v,30),DistanceShore=rep(Dshore_v,rep(30,30)))

# true values of the function over this grid
truth <- matrix(fomega(pr),30,30)

#points on the surface perturbed by gaussian noise
f <- fomega(samples)
y <- f+0*rnorm(n)

#plot true function
persp(theta_v,Dshore_v,truth);
title("truth")

#fit with bivariate splines te
m1 <- gam(y~te(theta,DistanceShore,k=SP_DF))

#visualize
vis.gam(m1);title("tensor product")

df=data.frame("x"=theta,"z"=DistanceShore,"y"=y)
fig <- plot_ly(df,type="scatter3d", x = ~x, y = ~z, z = ~y,mode="markers",marker=list(size=4))
fig
```

### Approximate smooth functions with tensor splines of ExpShore

```{r approximate smooth omega with ExpShore splines}
ExpShore=1/DistanceShore

#fit with bivariate splines te
m2 <- gam(y~te(theta,ExpShore,k=SP_DF))

#visualize
vis.gam(m2);title("tensor product")

```

### Get estimated spline coefficients

```{r estimated spline coeffs}
sp_coeff_Dshore=m1$coefficients
sp_coeff_ExpShore=m2$coefficients
```

### Define spline smooth functions for omega based on the approximations

```{r spline smooth omega definition}

fomega_splines=function(cov_data) {

  if (is.null(cov_data$ExpShore)){
    omega=predict(m1,newdata=cov_data)
  }
  else if (is.null(cov_data$DistanceShore)){
    omega=predict(m2,newdata=cov_data)
  }

  return(as.numeric(omega))
}

```

### Plot

```{r plot spline smooth omega}

Dshore=seq(from=0.3,to=2,length.out=30)
theta=seq(from=-pi,to=pi,length.out=30)
grid<- as.data.frame(expand.grid(DistanceShore,theta))
colnames(grid)=c("DistanceShore","theta")
true_par=grid
true_par$omega=fomega_splines(grid)

plot_true_omega <- plot_ly(true_par,type = "mesh3d",
                         x = ~theta,y = ~DistanceShore,z=~omega,intensity=~omega,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("True omega"),scene=list(xaxis = list(title = "Theta",showgrid = F),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

plot_true_omega
```

## Simulation with low measurement errors and high frequency

### Generate sample

```{r parallel generate hf low errors samples}

#setup parallel backend to use many processors
cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)

data_rect_hf_ne <- foreach(i=1:N_BATCH,.packages=c("progress","MASS","sf","foreach")) %dopar% {
  data=foreach (i=1:N_ID,.combine='rbind') %do% {

    res=sim_theta_CRCVM(ftau_constant,fomega,fnu_constant,log_sigma_obs=log(sigma_obs_0),v0,x0[i,],times_hf,land=domain,verbose=FALSE)

    data_sim=res$sim
    data_sim$ID=factor(rep(i,length(data_sim$y1)))
    data_shore=res$shore[,c("p1","p2")]
    cbind(data_sim,data_shore)
  }
  data 
}

#stop cluster
stopCluster(cl)

names(data_rect_hf_ne)=paste("BATCH",1:N_BATCH,sep="")
```

### percentage of samples that reached land

```{r percentage of low error samples that reached land}
count=0
for (k in 1:length(names(data_rect_hf_ne))) {

  data=data_rect_hf_ne[[paste("BATCH",k,sep="")]]
  for (id in unique(data$ID)) {
    sub_data=data[data$ID==id,]
    if (nrow(sub_data) < n_hf) {
      count=count+1
      cat(paste("BATCH",k,sep=""),":",id,"\n",sep=" ")
    }
  }
}
cat(count/(N_BATCH*N_ID)*100,"percent of the samples reached land")
```

### plot one batch

```{r plot one batch low error samples}

plot_samples_rect_hf_ne=ggplot()+geom_sf(data=domain$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_rect_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_path(data=data_rect_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_point(data = data_rect_hf_ne$BATCH1%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

ggsave(filename="samples_rect_hf_ne_omega.png",plot=plot_samples_rect_hf_ne)
ggplotly(plot_samples_rect_hf_ne)
```

## Simulation with high measurement error and high frequency

### Generate samples

```{r high error samples, cache=TRUE,results='hide'}
#setup parallel backend to use many processors
cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)

data_rect_hf_e <- foreach(i=1:N_BATCH,.packages=c("progress","MASS","sf","foreach")) %dopar% {
  data=foreach (i=1:N_ID,.combine='rbind') %do% {

    res=sim_theta_CRCVM(ftau_constant,fomega,fnu_constant,log_sigma_obs=log(sigma_obs_1),v0,x0[i,],times_hf,land=domain,verbose=FALSE)

    data_sim=res$sim
    data_sim$ID=factor(rep(i,length(data_sim$y1)))
    data_shore=res$shore[,c("p1","p2")]
    cbind(data_sim,data_shore)
  }
  data 
}

#stop cluster
stopCluster(cl)

#change names of the list elements
names(data_rect_hf_e)=paste("BATCH",1:N_BATCH,sep="")
```

### percentage of samples that reached land

```{r percentage of high error samples that reached land}
count=0
for (k in 1:length(names(data_rect_hf_e))) {

  data=data_rect_hf_e[[paste("BATCH",k,sep="")]]
  for (id in unique(data$ID)) {
    sub_data=data[data$ID==id,]
    if (nrow(sub_data) < n_hf) {
      count=count+1
      cat("BATCH",k,id,"\n",sep="")
    }
  }
}
cat(count/(N_BATCH*N_ID)*100,"percent of the samples reached land")
```

### plot one sample

```{r plot one batch of high error samples}

plot_samples_rect_hf_e=ggplot()+geom_sf(data=domain$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_rect_hf_e$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_path(data=data_rect_hf_e$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_point(data = data_rect_hf_e$BATCH1%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

ggsave(filename="samples_rect_hf_e_omega.png",plot=plot_samples_rect_hf_e)
ggplotly(plot_samples_rect_hf_e)
```

# ESTIMATION

## Compute distance to shore and empirical deviation angles

```{r compute covs}

signed_angle <- function(u, v) {
    #Compute signed angle in [-pi,pi] that rotates first vector into second vector 
    # as in 
    # https://math.stackexchange.com/questions/529555/signed-angle-between-2-vectors
    u <- matrix(u, ncol = 2)
    v <- matrix(v, ncol = 2)
    if (nrow(u) != nrow(v)) stop("u and v must have the same number of 
                                  rows")
    result <- as.numeric(atan2(v[,2], v[,1]) - atan2(u[,2], u[,1]))
    ind1 <- which(result > pi)
    ind2 <- which(result <= -pi)
    result[ind1] <- result[ind1] - 2*pi
    result[ind2] <- result[ind2] + 2*pi
    return(result) 
}

add_covs=function(data) {
  #data has at least columns "time","y1" "y2", "p1","p2"
  #this function new data frame with three more columns "theta", "DistanceShore" and "ExpShore"

  new_data=data
  #observed deviation angles
  n=length(new_data$time)

  new_data$theta=rep(NA,n)

  for (id in unique(new_data$ID)) {

    #filter data with specific id
    sub_ind=(new_data$ID==id)
    sub_data=new_data[sub_ind,]
    n_sub=length(sub_data$time)

    #time steps
    dtimes=sub_data[2:n_sub,"time"]-sub_data[1:(n_sub-1),"time"]

    #step lengths
    dx=sub_data[2:n_sub,"y1"]-sub_data[1:(n_sub-1),"y1"]
    dy=sub_data[2:n_sub,"y2"]-sub_data[1:(n_sub-1),"y2"]

    #matrix of empirical velocity
    vexp_df=cbind(dx/dtimes,dy/dtimes)

    #matrix of normal vectors
    normal=as.matrix(sub_data[2:n_sub,c("y1","y2")]-sub_data[2:n_sub,c("p1","p2")])

    #angle between velocity and normal vector
    theta_coast=signed_angle(normal,vexp_df)

    #adjust lengths 
    theta_coast=c(theta_coast,1)

    new_data[sub_ind,"theta"]=theta_coast
  }

  #distance to shore
  Dshore=sqrt((new_data$y1-new_data$p1)^2+(new_data$y2-new_data$p2)^2)
  new_data$DistanceShore=Dshore
  new_data$ExpShore=1/Dshore

  return(new_data)
}

for (k in 1:N_BATCH) {

  #dataframes for this batch
  df_ne=data_rect_hf_ne[[paste("BATCH",k,sep="")]]
  df_e=data_rect_hf_e[[paste("BATCH",k,sep="")]]

  #add covariates columns
  df_ne=add_covs(df_ne)
  df_e=add_covs(df_e)

  #change in the list of batch
  data_rect_hf_ne[[paste("BATCH",k,sep="")]]=df_ne
  data_rect_hf_e[[paste("BATCH",k,sep="")]]=df_e
}

```

## Estimate from high frequency data with low error

### Use ExpShore covariate and tensor splines te

#### Fit sde

```{r estimate te ExpShore for low error samples,cache=TRUE}

#setup parallel backend to use many processors
cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)

crcvm_rect_hf_ne1<- foreach(i=1:N_BATCH,.packages=c("smoothSDE")) %dopar% {
  #estimation with tensor splines
  formulas <- list(mu1=~1,mu2=~1,tau=~s(ID,bs="re"),
                 nu=~s(ID,bs="re"),omega=~te(theta,ExpShore,k=SP_DF,bs="cs"))
  par0 <- c(0,0,1,1,0)
  m <- SDE$new(formulas = formulas,data = data_rect_hf_ne[[paste("BATCH",k,sep="")]],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2"),
                      other_data=list("H"=H0_hf))
  m$fit(method="BFGS")
  m 
}

#stop cluster
stopCluster(cl)

names(crcvm_rect_hf_ne1)=paste("BATCH",1:N_BATCH,sep="")

```

#### Plot estimated smooth parameter omega in 3D for one batch

```{r plot estimates te ExpShore for one batch of low error samples}
xmin=list("theta"=-pi,"ExpShore"=1/2)
xmax=list("theta"=pi,"ExpShore"=1/0.3)
xlabel=list("ExpShore"="Distance to shore")
link=list("ExpShore"=(\(x) 1/x))
true_smooths=list("omega"=fomega_splines)
res=crcvm_rect_hf_ne1$BATCH1$get_all_plots(baseline=NULL,model_name="crcvm_rect_hf_ne1",xmin=xmin,xmax=xmax,
                                    xlabel=xlabel,link=link,show_CI="pointwise",true_smooths=true_smooths,save=TRUE)
res$fe_omega_theta_ExpShore
```

#### Violin plot of the estimated spline coefficients for one batch

```{r violin plots spline coeff te ExpShore for one batch of low error samples}

true_coeff=data.frame(value=sp_coeff_ExpShore,name=c(paste("omega.",names(sp_coeff_ExpShore),sep="")))
rownames(true_coeff)=NULL

p=crcvm_rect_hf_ne1$BATCH1$plot_violin_coeff(model_name="crcvm_rect_hf_ne1",substring="omega",true_coeff=true_coeff,save=TRUE)

ggplotly(p)
```

#### Violin plot of the estimated spline coefficients for all batches

```{r violin plots spline coeff te ExpShore all batches with low measurement error}

n_coeffs=length(sp_coeff_ExpShore)
coeffs_df=data.frame("estimate"=NA,"coeff_name"=NA)

for (k in 1:N_BATCH) {
  m<-crcvm_rect_hf_ne1[[paste("BATCH",k,sep="")]]
  coeffs=m$coeff_re()
  coeff_names=rownames(coeffs)
  coeff_values=as.numeric(coeffs)
  sub_index=grepl("omega",coeff_names)
  sub_coeffs_df=data.frame("estimate"=coeff_values[sub_index],"coeff_name"=factor(coeff_names[sub_index]))
  coeffs_df=rbind(coeffs_df,sub_coeffs_df)
}

#remove initialization row
coeffs_df=coeffs_df[-1,]

p=ggplot()+geom_violin(data=coeffs_df,aes(x=coeff_name,y=estimate,fill=coeff_name))+
                xlab(" ")+labs(fill = "Estimated coefficients")+
                theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
                geom_point(data = true_coeff, aes(x = name, y = value), color = "red", size = 1,shape=4)

ggsave("all_batches_coeffs_crcvm_rect_hf_ne1.png",plot=p,path="crcvm_rect_hf_ne1")

ggplotly(p)
```

#### Simulate from fitted model

```{r simulate fitted te ExpShore low error samples,cache=TRUE}

fExpShore=function(z,v,p) {

  #normal vector
  normal=c(z[1]-p[1],z[2]-p[2])

  #distance to shore
  Dshore=sqrt(normal[1]^2+normal[2]^2)
  if (Dshore<0.01) {
    return (100)
  }
  else {
    return (1/Dshore)
  }
}

fangle=function(z,v,p) {

  #normal vector
  normal=c(z[1]-p[1],z[2]-p[2])

  #return angle
  return (signed_angle(normal,v))
}

atw=list("ExpShore"=fExpShore,"AngleNormal"=fangle)
sim_data=crcvm_rect_hf_ne1$BATCH1$simulate(z0=x0[1,],data=crcvm_rect_hf_ne1$BATCH1$data(),atw=atw,land=domain)

```

#### Plot simulation from fitted model

```{r plot sim te ExpSHore low error samples}

plot_sim=ggplot()+geom_sf(data=domain$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=sim_data,mapping=aes(y1,y2),size=0.1)+
  geom_path(data=sim_data,mapping=aes(y1,y2),size=0.1)+
  geom_point(data = sim_data%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2 ), shape = 3, size = 4, col = "red")+
    xlab("x") + ylab("y")

ggsave("sim_crcvm_rect_hf_ne1.png",plot=plot_sim,width=10,height=5,path="crcvm_rect_hf_ne1")
ggplotly(plot_sim)
```

###  Use DistanceShore covariate and tensor splines te

#### Fit sde

```{r estimate te DistanceShore for low error samples,cache=TRUE}
#estimation with tensor splines
formulas <- list(mu1=~1,mu2=~1,tau=~s(ID,bs="re"),
                 nu=~s(ID,bs="re"),omega=~te(theta,DistanceShore,k=SP_DF,bs="cs"))

crcvm_rect_hf_ne2=list()
for (k in 1:N_BATCH) {
  m <- SDE$new(formulas = formulas,data = data_rect_hf_ne[[paste("BATCH"),k,sep=""]],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2"),
                      other_data=list("H"=H0_hf))
  m$fit(method="BFGS")
  crcvm_rect_hf_ne2[[paste("BATCH",k,sep="")]]<-m
}
```

#### Plots estimated smooth in 3D

```{r plot estimates te DistanceShore for one batch of low error samples}
xmin=list("theta"=-pi,"DistanceShore"=0.3)
xmax=list("theta"=pi,"DistanceShore"=2)
res=crcvm_rect_hf_ne2$BATCH1$get_all_plots(baseline=NULL,model_name="crcvm_rect_hf_ne2",xmin=xmin,xmax=xmax,
                                    xlabel=xlabel,link=link,show_CI="pointwise",true_smooths=true_smooths,save=TRUE)
res$fe_omega_theta_DistanceShore
```

#### Violin plot of the estimated spline coefficients

```{r violin plots spline coeffs te DistanceShore for one batch of low error sample}

true_coeff=data.frame(value=sp_coeff_Dshore,name=c(paste("omega.",names(sp_coeff_Dshore),sep="")))
rownames(true_coeff)=NULL

p=crcvm_rect_hf_ne2$BATCH1$plot_violin_coeff(model_name="crcvm_rect_hf_ne2",substring="omega",
                                             true_coeff=true_coeff,save=TRUE)

ggplotly(p)
```

#### Violin plot of the estimated spline coefficients for all batches

```{r violin plots spline coeff te DistanceShore all batches with low measurement error}

n_coeffs=length(sp_coeff_DShore)
coeffs_df=data.frame("estimate"=NA,"coeff_name"=NA)

for (k in 1:N_BATCH) {
  m<-crcvm_rect_hf_ne2[[paste("BATCH",k,sep="")]]
  coeffs=m$coeff_re()
  coeff_names=rownames(coeffs)
  coeff_values=as.numeric(coeffs)
  sub_index=grepl("omega",coeff_names)
  sub_coeffs_df=data.frame("estimate"=coeff_values[sub_index],"coeff_name"=factor(coeff_names[sub_index]))
  coeffs_df=rbind(coeffs_df,sub_coeffs_df)
}

#remove initialization row
coeffs_df=coeffs_df[-1,]

p=ggplot()+geom_violin(data=coeffs_df,aes(x=coeff_name,y=estimate,fill=coeff_name))+
                xlab(" ")+labs(fill = "Estimated coefficients")+
                theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
                geom_point(data = true_coeff, aes(x = name, y = value), color = "red", size = 1,shape=4)

ggsave("all_batches_coeffs_crcvm_rect_hf_ne2.png",plot=p,width=10,height=5,path="crcvm_rect_hf_ne2")

```

## Estimate from high frequency data with error

### Use ExpShore and tensor splines te

#### Fit sde

```{r estimate te ExpShore for high error samples, cache=TRUE}
#estimation with tensor splines
formulas <- list(mu1=~1,mu2=~1,tau=~s(ID,bs="re"),
                 nu=~s(ID,bs="re"),omega=~te(theta,ExpShore,k=SP_DF,bs="cs"))

crcvm_rect_hf_e1=list()
for (k in 1:N_BATCH) {
  m <- SDE$new(formulas = formulas,data = data_rect_hf_e[[paste("BATCH",k,sep="")]],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2"),
                      other_data=list("H"=H0_hf))
  m$fit(method="BFGS")
  crcvm_rect_hf_e1[[paste("BATCH",k,sep="")]]<-m
}
```

#### Plot estimated smooth in 3D for one batch

```{r plots estimates te ExpShore for one batch of high error samples}
xmin=list("theta"=-pi,"ExpShore"=1/2)
xmax=list("theta"=pi,"ExpShore"=1/0.3)
xlabel=list("ExpShore"="Distance to shore")
link=list("ExpShore"=(\(x) 1/x))
true_smooths=list("omega"=fomega)
res=crcvm_rect_hf_e1$BATCH1$get_all_plots(baseline=NULL,model_name="crcvm_rect_hf_e1",xmin=xmin,xmax=xmax,
                                    xlabel=xlabel,link=link,show_CI="pointwise",true_smooths=true_smooths,save=TRUE)
res$fe_omega_theta_ExpShore
```

#### Violin plot of the estimated spline coefficients for one batch

```{r violin plots spline coeffs te ExpShore for one batch of high error sample}

true_coeff=data.frame(value=sp_coeff_ExpShore,name=paste("omega.",names(sp_coeff_ExpShore),sep=""))
rownames(true_coeff)=NULL

p=crcvm_rect_hf_e1$BATCH1$plot_violin_coeff(model_name="crcvm_rect_hf_e1",substring="omega",true_coeff=true_coeff,save=TRUE)

ggplotly(p)
```

#### Violin plot of the estimated spline coefficients for all batches

```{r violin plots spline coeff te ExpShore all batches with high measurement error}

n_coeffs=length(sp_coeff_ExpShore)
coeffs_df=data.frame("estimate"=NA,"coeff_name"=NA)

for (k in 1:N_BATCH) {
  m<-crcvm_rect_hf_e1[[paste("BATCH",k,sep="")]]
  coeffs=m$coeff_re()
  coeff_names=rownames(coeffs)
  coeff_values=as.numeric(coeffs)
  sub_index=grepl("omega",coeff_names)
  sub_coeffs_df=data.frame("estimate"=coeff_values[sub_index],"coeff_name"=factor(coeff_names[sub_index]))
  coeffs_df=rbind(coeffs_df,sub_coeffs_df)
}

#remove initialization row
coeffs_df=coeffs_df[-1,]

p=ggplot()+geom_violin(data=coeffs_df,aes(x=coeff_name,y=estimate,fill=coeff_name))+
                xlab(" ")+labs(fill = "Estimated coefficients")+
                theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
                geom_point(data = true_coeff, aes(x = name, y = value), color = "red", size = 1,shape=4)

ggsave("all_batches_coeffs_crcvm_rect_hf_e1.png",plot=p,width=10,height=5,path="crcvm_rect_hf_e1")

ggplotly(p)

```

### Use DistanceShore and tensor splines te

#### Fit sde

```{r estimate te DistanceShore for high error samples,cache=TRUE}
#estimation with tensor splines
formulas <- list(mu1=~1,mu2=~1,tau=~s(ID,bs="re"),
                 nu=~s(ID,bs="re"),omega=~te(theta,DistanceShore,k=SP_DF,bs="cs"))

crcvm_rect_hf_e2=list()

for (k in 1:N_BATCH) {
  m <- SDE$new(formulas = formulas,data = data_rect_hf_e[[paste("BATCH",k,sep="")]],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2"),
                      other_data=list("H"=H0_hf))
  m$fit(method="BFGS")
  crcvm_rect_hf_e2[[paste("BATCH",k,sep="")]]<-m
}
```

#### Plot estimated smooth in 3D for one batch

```{r plots estimates te DistanceShore for one batch of high error samples}
xmin=list("theta"=-pi,"DistanceShore"=0.3)
xmax=list("theta"=pi,"DistanceShore"=2)
res=crcvm_rect_hf_e2$BATCH1$get_all_plots(baseline=NULL,model_name="crcvm_rect_hf_e2",xmin=xmin,xmax=xmax,
                                    xlabel=xlabel,link=link,show_CI="pointwise",true_smooths=true_smooths,save=TRUE)
res$fe_omega_theta_DistanceShore
```

#### Violin plot of the estimated spline coefficients for one batch

```{r violin plots spline coeffs te DistanceShore for one batch of high error sample}

true_coeff=data.frame(value=sp_coeff_Dshore,name=paste("omega.",names(sp_coeff_Dshore),sep=""))
rownames(true_coeff)=NULL

p=crcvm_rect_hf_e2$BATCH1$plot_violin_coeff(model_name="crcvm_rect_hf_e2",substring="omega",true_coeff=true_coeff,save=TRUE)

ggplotly(p)
```

#### Violin plot of the estimated spline coefficients for all batches

```{r violin plots spline coeff te DistanceShore all batches with high measurement error}

n_coeffs=length(sp_coeff_DShore)
coeffs_df=data.frame("estimate"=NA,"coeff_name"=NA)

for (k in 1:N_BATCH) {
  m<-crcvm_rect_hf_e2[[paste("BATCH",k,sep="")]]
  coeffs=m$coeff_re()
  coeff_names=rownames(coeffs)
  coeff_values=as.numeric(coeffs)
  sub_index=grepl("omega",coeff_names)
  sub_coeffs_df=data.frame("estimate"=coeff_values[sub_index],"coeff_name"=factor(coeff_names[sub_index]))
  coeffs_df=rbind(coeffs_df,sub_coeffs_df)
}

#remove initialization row
coeffs_df=coeffs_df[-1,]

p=ggplot()+geom_violin(data=coeffs_df,aes(x=coeff_name,y=estimate,fill=coeff_name))+
                xlab(" ")+labs(fill = "Estimated coefficients")+
                theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
                geom_point(data = true_coeff, aes(x = name, y = value), color = "red", size = 1,shape=4)

ggsave("all_batches_coeffs_crcvm_rect_hf_e2.png",plot=p,width=10,height=5,path="crcvm_rect_hf_e2")
```