### 2024-11-08 R/application/CRCVM/fit_crcvm.Rmd

---
title: "fit_crcvm"
author: "Alexandre Delporte"
date: "2024-06-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up

```{r packages}
set.seed(42)

library(smoothSDE)
library(ggplot2)
library(dplyr)
library(plotly)
library(htmlwidgets)
library(mgcv)
```

```{r get data}

# Set the path to the directory containing the data
par_dir=dirname(dirname(dirname(getwd()))) #parent directory 
narwhal_data_path <- file.path(par_dir,"Data", "Narwhals")

#data before exposure with first 24h removed
dataBE1=read.csv(file.path(narwhal_data_path,"DataBE1.csv"), header = TRUE,dec = ".")

#data before exposure with first 12h removed
dataBE2=read.csv(file.path(narwhal_data_path,"DataBE2.csv"), header = TRUE,dec = ".")

#data after exposure
dataAE=read.csv(file.path(narwhal_data_path,"DataAE.csv"), header = TRUE,dec = ".")

```

We choose thresholds on the distance to the shore to define the ExpShore covariate.

```{r preprocess data}

D_low=0.073
D_up=3
dataBE1[dataBE1$DistanceShore> D_up,"ExpShore"]=0
dataBE1[dataBE1$DistanceShore< D_low,"ExpShore"]=1/D_low

dataBE2[dataBE2$DistanceShore> D_up,"ExpShore"]=0
dataBE2[dataBE2$DistanceShore< D_low,"ExpShore"]=1/D_low

dataAE[dataAE$DistanceShore> D_up,"ExpShore"]=0
dataAE[dataAE$DistanceShore< D_low,"ExpShore"]=1/D_low

N=length(unique(dataBE2$ID))

n_pre=length(dataBE2$time) #number of observations before exposure
n_post=length(dataAE$time) #number of observations after exposure
```

## Baseline models

We can estimate the measurement error standard deviation directly from the data. Based on accuracy studies about fastloc GPS, it should be a few tens of metres. The estimate we get from the data before exposure is 35m (std error 1m), which seems consistent with previous studies. However, when estimating this measurement error from the post exposure data, we get a higher value of 48m, which is weird : the same GPS is device is used during all the experiment, so measurement error before or after exposure should be about the same. If we fix the measurement error to 35m when estimating the response, then we can't estimate because TMB gets a non finite value of the negative log-likelihood at initial parameters. This cannot be solved by changing the other initial parameters in the optimisation. Thus we decided to fix the measurement error to different values in the baseline and the response and compare the results we get. Higher measurement error gives higher estimates for tau, whereas the estimate of nu is not affected by the changes in the measurement error.

### Baselines with constant parameters

```{r fit baseline0}

#initial parameters
par0 <- c(0,0,1,4,0)

#model formula
formulas <- list(mu1=~1,mu2=~1,tau =~s(ID,bs="re"),nu=~s(ID,bs="re"),omega=~1)

# Fit baseline 
baseline0<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("log_sigma_obs0"=log(0.05)),fixpar=c("mu1","mu2"))
baseline0$fit()

# Fit baseline with fixed 25m measurement error
sigma_obs=0.025
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline0_25m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"))
baseline0_25m$fit()

# Fit baseline with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline0_50m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"))
baseline0_50m$fit()

# Fit baseline with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline0_75m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"))
baseline0_75m$fit()

# Fit baseline with 100m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline0_100m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"))
baseline0_100m$fit()

```

```{r results baseline0}

estimates_bas0=as.list(baseline0$tmb_rep(),what="Est")
std_bas0=as.list(baseline0$tmb_rep(),what="Std")

estimates_bas0_25m=as.list(baseline0_25m$tmb_rep(),what="Est")
std_bas0_25m=as.list(baseline0_25m$tmb_rep(),what="Std")

estimates_bas0_50m=as.list(baseline0_50m$tmb_rep(),what="Est")
std_bas0_50m=as.list(baseline0_50m$tmb_rep(),what="Std")

estimates_bas0_75m=as.list(baseline0_75m$tmb_rep(),what="Est")
std_bas0_75m=as.list(baseline0_75m$tmb_rep(),what="Std")

estimates_bas0_100m=as.list(baseline0_100m$tmb_rep(),what="Est")
std_bas0_100m=as.list(baseline0_100m$tmb_rep(),what="Std")

```

### Baseline with AngleNormal covariate in omega

Fixing a high standard deviation (>50m) for the measurement error tends to increase the estimate of the parameter tau. Intuitively, for a low value of the measurement error, tortuous trajectory can only be caused by less persistent movement, whereas if the measurement error is high, a part of the tortuosity can be attributed to the measurement errors, which might give higher estimate of the persistence parameter.

```{r fit baseline1}

#initial parameters
par0 <- c(0,0,1,4,0)
init_lambda=c(1,1,1)

#model formula
formulas <- list(mu1=~1,mu2=~1,
                 tau =~s(ID,bs="re"),
                 nu=~s(ID,bs="re"),
                 omega=~s(AngleNormal,k=5,bs="cp"))

# Fit baseline 
baseline1<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("log_sigma_obs0"=log(0.05)),fixpar=c("mu1","mu2"))

baseline1$update_map(new_map=list(log_lambda=factor(c("tau.s(ID)","nu.s(ID)",NA))))
baseline1$fit()

# Fit baseline with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline1_50m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"))

baseline1_50m$update_lambda(init_lambda)

baseline1_50m$update_map(new_map=list(log_lambda=factor(c("tau.s(ID)","nu.s(ID)",NA))))

baseline1_50m$fit()

# Fit baseline with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline1_75m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"))

baseline1_75m$update_lambda(init_lambda)

baseline1_75m$update_map(new_map=list(log_lambda=factor(c("tau.s(ID)","nu.s(ID)",NA))))

baseline1_75m$fit()

# Fit baseline with 100m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline1_100m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"))

baseline1_100m$update_lambda(init_lambda)

baseline1_100m$update_map(new_map=list(log_lambda=factor(c("tau.s(ID)","nu.s(ID)",NA))))

baseline1_100m$fit()

```

```{r results baseline1}

#get estimates
estimates_bas1=as.list(baseline1$tmb_rep(),what="Est")
std_bas1=as.list(baseline1$tmb_rep(),what="Std")

estimates_bas1_50m=as.list(baseline1_50m$tmb_rep(),what="Est")
std_bas1_50m=as.list(baseline1_50m$tmb_rep(),what="Std")

estimates_bas1_75m=as.list(baseline1_75m$tmb_rep(),what="Est")
std_bas1_75m=as.list(baseline1_75m$tmb_rep(),what="Std")

estimates_bas1_100m=as.list(baseline1_100m$tmb_rep(),what="Est")
std_bas1_100m=as.list(baseline1_100m$tmb_rep(),what="Std")

#get plots

plots_bas1=baseline1$get_all_plots(baseline=NULL,show_CI="simultaneous")
plots_bas1_50m=baseline1_50m$get_all_plots(baseline=NULL,show_CI="pointwise")
plots_bas1_75m=baseline1_75m$get_all_plots(baseline=NULL,show_CI="pointwise")
plots_bas1_100m=baseline1_100m$get_all_plots(baseline=NULL,show_CI="pointwise")

```

```{r display plots baseline1}

plots_bas1
plots_bas1_50m
plots_bas1_75m
plots_bas1_100m
```

### Baseline with AngleNormal and DistanceShore in omega

```{r fit baseline2}

#initial parameters
par0 <- c(0,0,1,4,0)
init_lambda=c(1,1,0.1,1,1,1)

formulas <- list(mu1 = ~1 ,mu2 =~1,tau =~s(ID,bs="re"),nu=~s(ID,bs="re"),
                 omega=~ti(DistanceShore,k=5,bs="cs")+ti(AngleNormal,k=5,bs="cp")+ti(AngleNormal,DistanceShore,k=c(5,5),bs="cs"))

# Fit baseline 
baseline2<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("log_sigma_obs0"=log(0.035)),fixpar=c("mu1","mu2"))

baseline2$update_map(new_map=list(log_lambda=factor(c("tau.s(ID)","nu.s(ID)",rep(NA,4)))))
baseline2$fit()

# Fit baseline with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline2_50m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",
                    response = c("x","y"),par0 = par0,other_data=list("H"=H),
                    fixpar=c("mu1","mu2"))

baseline2_50m$update_lambda(init_lambda)
baseline2_50m$update_map(new_map=list(log_lambda=factor(c("tau.s(ID)","nu.s(ID)",rep(NA,4)))))
baseline2_50m$fit()

# Fit baseline with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline2_75m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",
                    response = c("x","y"),par0 = par0,other_data=list("H"=H),
                    fixpar=c("mu1","mu2"))

baseline2_75m$update_lambda(init_lambda)
baseline2_75m$update_map(new_map=list(log_lambda=factor(c("tau.s(ID)","nu.s(ID)",rep(NA,4)))))
baseline2_75m$fit()

# Fit baseline with 100m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline2_100m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",
                    response = c("x","y"),par0 = par0,other_data=list("H"=H),
                    fixpar=c("mu1","mu2"))

baseline2_100m$update_lambda(init_lambda)
baseline2_100m$update_map(new_map=list(log_lambda=factor(c("tau.s(ID)","nu.s(ID)",rep(NA,4)))))
baseline2_100m$fit()

```

```{r get results baseline2}

estimates_bas2=as.list(baseline2$tmb_rep(),what="Est")
std_bas2=as.list(baseline2$tmb_rep(),what="Std")

estimates_bas2_50m=as.list(baseline2_50m$tmb_rep(),what="Est")
std_bas2_50m=as.list(baseline2_50m$tmb_rep(),what="Std")

estimates_bas2_75m=as.list(baseline2_75m$tmb_rep(),what="Est")
std_bas2_75m=as.list(baseline2_75m$tmb_rep(),what="Std")

estimates_bas2_100m=as.list(baseline2_100m$tmb_rep(),what="Est")
std_bas2_100m=as.list(baseline2_100m$tmb_rep(),what="Std")

#plot parameters

xmin=list("DistanceShore"=0.073)
xmax=list("DistanceShore"=5)

plots_bas2=baseline2$get_all_plots(baseline=NULL,xmin=xmin,xmax=xmax,show_CI="pointwise")

plots_bas2_50m=baseline2_50m$get_all_plots(baseline=NULL,xmin=xmin,xmax=xmax,show_CI="pointwise")
plots_bas2_75m=baseline2_75m$get_all_plots(baseline=NULL,xmin=xmin,xmax=xmax,show_CI="pointwise")
plots_bas2_100m=baseline2_100m$get_all_plots(baseline=NULL,xmin=xmin,xmax=xmax,show_CI="pointwise")
```

```{r display plots baseline2}

#print plots
plots_bas2
plots_bas2_50m
plots_bas2_75m
plots_bas2_100m

```

```{r save plots baseline2}
# Create directory
if (!dir.exists("baseline2_50m")) {
    dir.create("baseline2_50m", recursive = TRUE)
}

# Use lapply to save each plot
lapply(seq_along(plots_bas2), function(i) {

    plot<-plots_bas2[[i]]
    plot_name <- names(plots_bas2)[i]
    file_path <- file.path("baseline2_50m", plot_name)

    # Check if the plot is a ggplot or plotly object
    if (inherits(plot, "ggplot")) {
      ggsave(file_path, plot, device = "pdf")
    } else if (inherits(plot, "plotly")) {
      saveWidget(plot, file_path)
  }
    })
```

```{r estimates on parameter scale baseline2}

post_coeff_bas2=baseline2$post_coeff(n_post=1000)
post_par_bas2=list("tau0"=exp(post_coeff_bas2$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_bas2$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_bas2$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_bas2$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_bas2$log_lambda[,2])))
mean_par_bas2=lapply(post_par_bas2,mean)
sd_par_bas2=lapply(post_par_bas2,sd)

post_coeff_bas2_50m=baseline2_50m$post_coeff(n_post=1000)
post_par_bas2_50m=list("tau0"=exp(post_coeff_bas2_50m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_bas2_50m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_bas2_50m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_bas2_50m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_bas2_50m$log_lambda[,2])))
mean_par_bas2_50m=lapply(post_par_bas2_50m,mean)
sd_par_bas2_50m=lapply(post_par_bas2_50m,sd)

post_coeff_bas2_75m=baseline2_75m$post_coeff(n_post=1000)
post_par_bas2_75m=list("tau0"=exp(post_coeff_bas2_75m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_bas2_75m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_bas2_75m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_bas2_75m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_bas2_75m$log_lambda[,2])))
mean_par_bas2_75m=lapply(post_par_bas2_75m,mean)
sd_par_bas2_75m=lapply(post_par_bas2_75m,sd)

#get 95% confidence intervals
post_coeff_bas2_100m=baseline2_100m$post_coeff(n_post=1000)
post_par_bas2_100m=list("tau0"=exp(post_coeff_bas2_100m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_bas2_100m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_bas2_100m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_bas2_100m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_bas2_100m$log_lambda[,2])))
mean_par_bas2_100m=lapply(post_par_bas2_100m,mean)
sd_par_bas2_100m=lapply(post_par_bas2_100m,sd)
```

```{r smooth omega baseline2}

#get angular velocity for different distances to shore

plot1=plots_bas2_50m$fe_omega_AngleNormal_q1_DistanceShore
plot2=plots_bas2_50m$fe_omega_AngleNormal_q2_DistanceShore
plot4=plots_bas2_50m$fe_omega_AngleNormal_q4_DistanceShore

omega1 <- ggplot_build(plot1)$data[[1]]
omega2 <- ggplot_build(plot2)$data[[1]]
omega4 <- ggplot_build(plot4)$data[[1]]

omega1$Category <- "Close"
omega2$Category <- "Medium"
omega4$Category <- "Far"

# Combine the data frames
combined_omega <- rbind(omega1, omega2, omega4)

ci_omega1 <- ggplot_build(plot1)$data[[2]]
ci_omega2 <- ggplot_build(plot2)$data[[2]]
ci_omega4 <- ggplot_build(plot4)$data[[2]]

ci_omega1$Category <- "Close"
ci_omega2$Category <- "Medium"
ci_omega4$Category <- "Far"

# Combine the data frames
combined_ci_omega <- rbind(ci_omega1, ci_omega2, ci_omega4)

combined_plot <- ggplot() +
  geom_line(data=combined_omega,aes(x=x,y = y,color = Category, linetype = Category)) +
  geom_ribbon(data=combined_ci_omega,aes(x=x,ymin=ymin,ymax=ymax,fill=Category),alpha=0.1)+
  labs(color = "Category", linetype = "Category",x="Theta",y="Omega") +
  geom_vline(xintercept = c(-pi/2, pi/2), linetype = "dashed", color = "grey")+
  annotate("text", x = -pi/2, y = Inf, label = expression(-pi/2), vjust = 40, color = "black") +
  annotate("text", x = pi/2, y = Inf, label = expression(pi/2), vjust = 40, color = "black")+
  theme_minimal()

ggsave("omega_DistanceShore_levels.png",plot=combined_plot,path="baseline2_50m",width = 10, height = 6, units = "in")

```

### Baseline with AngleNormal and ExpShore in omega

```{r fit baseline3}

#initial parameters
par0 <- c(0,0,1,4,0)
init_lambda=c(1,1,1,1,1,1)

#mapping for fixed smooth penalties
new_map=list("log_lambda"=factor(c("tau.s(ID)","nu.s(ID)",rep(NA,4))))

#parameters formulas
formulas <- list(mu1 = ~1 ,mu2 =~1,tau =~s(ID,bs="re"),nu=~s(ID,bs="re"),
                   omega=~ti(AngleNormal,k=5,bs="cs")+ti(ExpShore,k=5,bs="cs")+
                   ti(AngleNormal,ExpShore,k=c(5,5),bs="cs"))

# Fit baseline
baseline3<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",
                    response = c("x","y"),par0 = par0,other_data=list("log_sigma_obs0"=log(0.035)),
                    fixpar=c("mu1","mu2"))

baseline3$update_lambda(init_lambda)
baseline3$update_map(new_map)
baseline3$fit()

# Fit baseline with 30m measurement error
sigma_obs=0.03
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline3_30m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",
                    response = c("x","y"),par0 = par0,other_data=list("H"=H),
                    fixpar=c("mu1","mu2"))

baseline3_30m$update_lambda(init_lambda)
baseline3$update_map(new_map)
baseline3_30m$fit()

# Fit baseline with 40m measurement error
sigma_obs=0.04
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline3_40m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",
                    response = c("x","y"),par0 = par0,other_data=list("H"=H),
                    fixpar=c("mu1","mu2"))

baseline3_40m$update_lambda(init_lambda)
baseline3_40m$update_map(new_map)
baseline3_40m$fit()

# Fit baseline with 45m measurement error
sigma_obs=0.045
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline3_45m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",
                    response = c("x","y"),par0 = par0,other_data=list("H"=H),
                    fixpar=c("mu1","mu2"))

baseline3_45m$update_lambda(init_lambda)
baseline3_45m$update_map(new_map)
baseline3_45m$fit()

# Fit baseline with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline3_50m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",
                    response = c("x","y"),par0 = par0,other_data=list("H"=H),
                    fixpar=c("mu1","mu2"))

baseline3_50m$update_lambda(init_lambda)
baseline3_50m$update_map(new_map)
baseline3_50m$fit()

# Fit baseline with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline3_75m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",
                    response = c("x","y"),par0 = par0,other_data=list("H"=H),
                    fixpar=c("mu1","mu2"))

baseline3_75m$update_lambda(init_lambda)
baseline3_75m$update_map(new_map)
baseline3_75m$fit()

# Fit baseline with 100m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline3_100m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",
                    response = c("x","y"),par0 = par0,other_data=list("H"=H),
                    fixpar=c("mu1","mu2"))

baseline3_100m$update_lambda(init_lambda)
baseline3_100m$update_map(new_map)
baseline3_100m$fit()
```

```{r get results baseline3}

estimates_bas3=as.list(baseline3$tmb_rep(),what="Est")
std_bas3=as.list(baseline3$tmb_rep(),what="Std")

estimates_bas3_30m=as.list(baseline3_30m$tmb_rep(),what="Est")
std_bas3_30m=as.list(baseline3_30m$tmb_rep(),what="Std")

estimates_bas3_40m=as.list(baseline3_40m$tmb_rep(),what="Est")
std_bas3_40m=as.list(baseline3_40m$tmb_rep(),what="Std")

estimates_bas3_45m=as.list(baseline3_45m$tmb_rep(),what="Est")
std_bas3_45m=as.list(baseline3_45m$tmb_rep(),what="Std")

estimates_bas3_50m=as.list(baseline3_50m$tmb_rep(),what="Est")
std_bas3_50m=as.list(baseline3_50m$tmb_rep(),what="Std")

estimates_bas3_75m=as.list(baseline3_75m$tmb_rep(),what="Est")
std_bas3_75m=as.list(baseline3_75m$tmb_rep(),what="Std")

estimates_bas3_100m=as.list(baseline3_100m$tmb_rep(),what="Est")
std_bas3_100m=as.list(baseline3_100m$tmb_rep(),what="Std")

#plot parameters

xmin=list("ExpShore"=1/D_up)
xmax=list("ExpShore"=1/D_low)
link=list("ExpShore"=(\(x) 1/x))
xlabel=list("ExpShore"="Distance to shore")

plots_bas3=baseline3$get_all_plots(baseline=NULL,xmin=xmin,
                                   xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")

plots_bas3_30m=baseline3_30m$get_all_plots(baseline=NULL,xmin=xmin,xmax=xmax,
                                           link=link,xlabel=xlabel,show_CI="pointwise")

plots_bas3_40m=baseline3_40m$get_all_plots(baseline=NULL,xmin=xmin,xmax=xmax,
                                           link=link,xlabel=xlabel,show_CI="pointwise")

plots_bas3_45m=baseline3_45m$get_all_plots(baseline=NULL,xmin=xmin,xmax=xmax,
                                           link=link,xlabel=xlabel,show_CI="pointwise")

plots_bas3_50m=baseline3_50m$get_all_plots(baseline=NULL,xmin=xmin,xmax=xmax,
                                           link=link,xlabel=xlabel,show_CI="pointwise")

plots_bas3_75m=baseline3_75m$get_all_plots(baseline=NULL,xmin=xmin,xmax=xmax,
                                           link=link,xlabel=xlabel,show_CI="pointwise")

plots_bas3_100m=baseline3_100m$get_all_plots(baseline=NULL,xmin=xmin,xmax=xmax,
                                             link=link,xlabel=xlabel,show_CI="pointwise")

```

```{r display plots baseline3}

#print plots
plots_bas3
plots_bas3_50m
plots_bas3_75m
plots_bas3_100m

```

```{r}
# Create directory
if (!dir.exists("baseline3_50m")) {
    dir.create("baseline3_50m", recursive = TRUE)
}

# Use lapply to save each plot
lapply(seq_along(plots_bas2), function(i) {

    plot<-plots_bas2[[i]]
    plot_name <- names(plots_bas2)[i]
    file_path <- file.path("baseline3_50m", plot_name)

    # Check if the plot is a ggplot or plotly object
    if (inherits(plot, "ggplot")) {
      ggsave(file_path, plot, device = "pdf")
    } else if (inherits(plot, "plotly")) {
      saveWidget(plot, file_path)
  }
    })
```

```{r estimates on parameter scale baseline3}

post_coeff_bas3=baseline3$post_coeff(n_post=10000)
post_par_bas3=list("sigma_obs"=exp(post_coeff_bas3$log_sigma_obs),
              "tau0"=exp(post_coeff_bas3$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_bas3$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_bas3$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_bas3$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_bas3$log_lambda[,2])))
mean_par_bas3=lapply(post_par_bas3,mean)
sd_par_bas3=lapply(post_par_bas3,sd)
quant_par_bas3 <- lapply(post_par_bas3, quantile, probs = c(0.025, 0.975))

post_coeff_bas3_30m=baseline3_30m$post_coeff(n_post=10000)
post_par_bas3_30m=list("tau0"=exp(post_coeff_bas3_30m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_bas3_30m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_bas3_30m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_bas3_30m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_bas3_30m$log_lambda[,2])))
mean_par_bas3_30m=lapply(post_par_bas3_30m,mean)
sd_par_bas3_30m=lapply(post_par_bas3_30m,sd)
quant_par_bas3_30m <- lapply(post_par_bas3_30m, quantile, probs = c(0.025, 0.975))

post_coeff_bas3_40m=baseline3_40m$post_coeff(n_post=10000)
post_par_bas3_40m=list("tau0"=exp(post_coeff_bas3_40m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_bas3_40m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_bas3_40m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_bas3_40m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_bas3_40m$log_lambda[,2])))
mean_par_bas3_40m=lapply(post_par_bas3_40m,mean)
sd_par_bas3_40m=lapply(post_par_bas3_40m,sd)
quant_par_bas3_40m <- lapply(post_par_bas3_40m, quantile, probs = c(0.025, 0.975))

post_coeff_bas3_45m=baseline3_45m$post_coeff(n_post=10000)
post_par_bas3_45m=list("tau0"=exp(post_coeff_bas3_45m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_bas3_45m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_bas3_45m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_bas3_45m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_bas3_45m$log_lambda[,2])))
mean_par_bas3_45m=lapply(post_par_bas3_45m,mean)
sd_par_bas3_45m=lapply(post_par_bas3_45m,sd)
quant_par_bas3_45m <- lapply(post_par_bas3_45m, quantile, probs = c(0.025, 0.975))

post_coeff_bas3_50m=baseline3_50m$post_coeff(n_post=10000)
post_par_bas3_50m=list("tau0"=exp(post_coeff_bas3_50m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_bas3_50m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_bas3_50m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_bas3_50m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_bas3_50m$log_lambda[,2])))
mean_par_bas3_50m=lapply(post_par_bas3_50m,mean)
sd_par_bas3_50m=lapply(post_par_bas3_50m,sd)
quant_par_bas3_50m <- lapply(post_par_bas3_50m, quantile, probs = c(0.025, 0.975))

post_coeff_bas3_75m=baseline3_75m$post_coeff(n_post=10000)
post_par_bas3_75m=list("tau0"=exp(post_coeff_bas3_75m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_bas3_75m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_bas3_75m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_bas3_75m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_bas3_75m$log_lambda[,2])))
mean_par_bas3_75m=lapply(post_par_bas3_75m,mean)
sd_par_bas3_75m=lapply(post_par_bas3_75m,sd)
quant_par_bas3_75m <- lapply(post_par_bas3_75m, quantile, probs = c(0.025, 0.975))

post_coeff_bas3_100m=baseline3_100m$post_coeff(n_post=10000)
post_par_bas3_100m=list("tau0"=exp(post_coeff_bas3_100m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_bas3_100m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_bas3_100m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_bas3_100m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_bas3_100m$log_lambda[,2])))
mean_par_bas3_100m=lapply(post_par_bas3_100m,mean)
sd_par_bas3_100m=lapply(post_par_bas3_100m,sd)
quant_par_bas3_100m <- lapply(post_par_bas3_100m, quantile, probs = c(0.025, 0.975))

```

```{r smooth omega baseline3}
plot1=plots_bas3_50m$fe_omega_AngleNormal_q1_ExpShore
plot2=plots_bas3_50m$fe_omega_AngleNormal_q2_ExpShore
plot4=plots_bas3_50m$fe_omega_AngleNormal_q4_ExpShore

omega1 <- ggplot_build(plot1)$data[[1]]
omega2 <- ggplot_build(plot2)$data[[1]]
omega4 <- ggplot_build(plot4)$data[[1]]

omega1$Category <- "Far"
omega2$Category <- "Medium"
omega4$Category <- "Close"

# Combine the data frames
combined_omega <- rbind(omega1, omega2, omega4)

ci_omega1 <- ggplot_build(plot1)$data[[2]]
ci_omega2 <- ggplot_build(plot2)$data[[2]]
ci_omega4 <- ggplot_build(plot4)$data[[2]]

ci_omega1$Category <- "Far"
ci_omega2$Category <- "Medium"
ci_omega4$Category <- "Close"

# Combine the data frames
combined_ci_omega <- rbind(ci_omega1, ci_omega2, ci_omega4)

combined_plot <- ggplot() +
  geom_line(data=combined_omega, aes(x=x, y=y, color=Category, linetype=Category)) +
  geom_ribbon(data=combined_ci_omega, aes(x=x, ymin=ymin, ymax=ymax, fill=Category), alpha=0.1) +
  labs(color="Category", linetype="Category", x=expression(Theta), y=expression(omega)) +
  geom_vline(xintercept=c(-pi/2, pi/2), linetype="dashed", color="grey") +
  annotate("text", x = -pi/2-0.3, y = -5, label = expression(-pi/2), color = "red", size = 5) + 
  annotate("text", x = pi/2+0.2, y = -5, label = expression(pi/2), color = "red", size = 5) + 
  theme_minimal() +
  theme(
    axis.title.x = element_text(size=14),
    axis.title.y = element_text(size=14,angle=0,vjust=0.5, margin = margin(r = 10)),
    axis.text.y = element_text(size=12),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    legend.title = element_text(size=14),
    legend.text = element_text(size=12))

ggsave("omega_DistanceShore_levels.pdf", plot=combined_plot, path="baseline3", width=10, height=6, units="in")

```

## Response models with splines of ExpShip

### Response models based on baseline1

```{r fit response1}

par0 <- c(0,0,1,4,0)

#model formula
formulas <- list(mu1=~1,mu2=~1,
                 tau =~s(ExpShip,k=3,bs="cs")+s(ID,bs="re"),
                 nu=~s(ExpShip,k=3,bs="cs")+s(ID,bs="re"),
                 omega=~s(AngleNormal,k=5,bs="cs"))

#mapping for fiwed coefficients
new_map=list(coeff_re=factor(c("tau.s(ExpShip).1","tau.s(ExpShip).2",rep(NA,N),
                    "nu.s(ExpShip).1","nu.s(ExpShip).2",rep(NA,N),rep(NA,4))),coeff_fe=factor(rep(NA,5)),
                    log_lambda=factor(c("tau.s(ExpShip)",NA,"nu.s(ExpShip)",NA,NA)))
#Fit baseline
response1_sp<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("log_sigma_obs0"=log(0.05)),fixpar=c("mu1","mu2"))

new_coeff_re=c(rep(0,2),baseline1$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],rep(0,2),
               baseline1$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline1$coeff_re()[paste("omega.s(AngleNormal).",1:4,sep=""),1])
new_lambda=c(1,baseline1$lambda()["tau.s(ID)",1],1,baseline1$lambda()["nu.s(ID)",1],
             baseline1$lambda()["omega.s(AngleNormal)",1])

response1_sp$update_map(new_map)
response1_sp$update_coeff_re(new_coeff_re)
response1_sp$update_coeff_fe(baseline1$coeff_fe()[,1])

response1_sp$update_lambda(new_lambda)

response1_sp$fit()

# Fit response with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response1_sp_50m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"))

new_coeff_re=c(rep(0,2),baseline1_50m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],rep(0,2),
               baseline1_50m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline1_50m$coeff_re()[paste("omega.s(AngleNormal).",1:4,sep=""),1])
new_lambda=c(1,baseline1_50m$lambda()["tau.s(ID)",1],1,baseline1_50m$lambda()["nu.s(ID)",1],
             baseline1_50m$lambda()["omega.s(AngleNormal)",1])

response1_sp_50m$update_map(new_map)
response1_sp_50m$update_coeff_re(new_coeff_re)
response1_sp_50m$update_coeff_fe(baseline1_50m$coeff_fe()[,1])

response1_sp_50m$update_lambda(new_lambda)

response1_sp_50m$fit()

# Fit response with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response1_sp_75m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"))

new_coeff_re=c(rep(0,2),baseline1_75m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],rep(0,2),
               baseline1_75m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline1_75m$coeff_re()[paste("omega.s(AngleNormal).",1:4,sep=""),1])
new_lambda=c(1,baseline1_75m$lambda()["tau.s(ID)",1],1,baseline1_75m$lambda()["nu.s(ID)",1],
             baseline1_75m$lambda()["omega.s(AngleNormal)",1])

response1_sp_75m$update_map(new_map)
response1_sp_75m$update_coeff_re(new_coeff_re)
response1_sp_75m$update_coeff_fe(baseline1_75m$coeff_fe()[,1])

response1_sp_75m$update_lambda(new_lambda)

response1_sp_75m$fit()

# Fit response with 50m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response1_sp_100m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"))

new_coeff_re=c(rep(0,2),baseline1_100m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],rep(0,2),
               baseline1_100m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline1_100m$coeff_re()[paste("omega.s(AngleNormal).",1:4,sep=""),1])
new_lambda=c(1,baseline1_100m$lambda()["tau.s(ID)",1],1,baseline1_100m$lambda()["nu.s(ID)",1],
             baseline1_100m$lambda()["omega.s(AngleNormal)",1])

response1_sp_100m$update_map(new_map)
response1_sp_100m$update_coeff_re(new_coeff_re)
response1_sp_100m$update_coeff_fe(baseline1_100m$coeff_fe()[,1])

response1_sp_100m$update_lambda(new_lambda)

response1_sp_100m$fit()
```

```{r get results response1}

estimates_res1_sp=as.list(response1_sp$tmb_rep(),what="Est")
std_res1_sp=as.list(response1_sp$tmb_rep(),what="Std")

estimates_res1_sp_50m=as.list(response1_sp_50m$tmb_rep(),what="Est")
std_res1_sp_50m=as.list(response1_sp_50m$tmb_rep(),what="Std")

estimates_res1_sp_75m=as.list(response1_sp_75m$tmb_rep(),what="Est")
std_res1_sp_75m=as.list(response1_sp_75m$tmb_rep(),what="Std")

estimates_res1_sp_100m=as.list(response1_sp_100m$tmb_rep(),what="Est")
std_res1_sp_100m=as.list(response1_sp_100m$tmb_rep(),what="Std")

#plot parameters
xmin=list("AngleNormal"=-pi,"ExpShip"=1/45)
xmax=list("AngleNormal"=pi,"ExpShip"=1/3)
link=list("ExpShip"=(\(x) 1/x))
xlabel=list("ExpShip"="Distance to ship")

plots_res1_sp=response1_sp$get_all_plots(baseline=baseline1,xmin=xmin,
                                         xmax=xmax,link=link,xlabel=xlabel,show_CI="simultaneous")
plots_res1_sp_50m=response1_sp_50m$get_all_plots(baseline=baseline1_50m,xmin=xmin,
                                                 xmax=xmax,link=link,xlabel=xlabel,show_CI="simultaneous")
plots_res1_sp_75m=response1_sp_75m$get_all_plots(baseline=baseline1_75m,xmin=xmin,xmax=xmax,
                                                 link=link,xlabel=xlabel,show_CI="simultaneous")
plots_res1_sp_100m=response1_sp_100m$get_all_plots(baseline=baseline1_100m,xmin=xmin,
                                                   xmax=xmax,link=link,xlabel=xlabel,show_CI="simultaneous")

```

```{r display plots response1_sp}

#print plots
plots_res1_sp
plots_res1_sp_50m
plots_res1_sp_75m
plots_res1_sp_100m

```

```{r estimates on parameter scale response1}
post_coeff_res1_sp=response1_sp$post_coeff(n_post=1000)
post_par_res1_sp=list("tau0"=exp(post_coeff_res1_sp$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res1_sp$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res1_sp$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res1_sp$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res1_sp$log_lambda[,2])))
mean_par_res1_sp=lapply(post_par_res1_sp,mean)
sd_par_res1_sp=lapply(post_par_res1_sp,sd)

post_coeff_res1_sp_50m=response1_sp_50m$post_coeff(n_post=1000)
post_par_res1_sp_50m=list("tau0"=exp(post_coeff_res1_sp_50m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res1_sp_50m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res1_sp_50m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res1_sp_50m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res1_sp_50m$log_lambda[,2])))
mean_par_res1_sp_50m=lapply(post_par_res1_sp_50m,mean)
sd_par_res1_sp_50m=lapply(post_par_res1_sp_50m,sd)

post_coeff_res1_sp_75m=response1_sp_75m$post_coeff(n_post=1000)
post_par_res1_sp_75m=list("tau0"=exp(post_coeff_res1_sp_75m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res1_sp_75m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res1_sp_75m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res1_sp_75m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res1_sp_75m$log_lambda[,2])))
mean_par_res1_sp_75m=lapply(post_par_res1_sp_75m,mean)
sd_par_res1_sp_75m=lapply(post_par_res1_sp_75m,sd)

post_coeff_res1_sp_100m=response1_sp_100m$post_coeff(n_post=1000)
post_par_res1_sp_100m=list("tau0"=exp(post_coeff_res1_sp_100m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res1_sp_100m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res1_sp_100m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res1_sp_100m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res1_sp_100m$log_lambda[,2])))
mean_par_res1_sp_100m=lapply(post_par_res1_sp_100m,mean)
sd_par_res1_sp_100m=lapply(post_par_res1_sp_100m,sd)

```

### Response model with splines of ExpShip based on baseline2

```{r fit response2}

par0 <- c(0,0,1,4,0)

#model formula
formulas <- list(mu1=~1,mu2=~1,tau =~s(ExpShip,k=3,bs="cs")+s(ID,bs="re"),
                 nu=~s(ExpShip,k=3,bs="cs")+s(ID,bs="re"),
                 omega=~ti(DistanceShore,k=5,bs="cs")+ti(AngleNormal,k=5,bs="cs")+ti(AngleNormal,DistanceShore,k=c(5,5),bs="cs"))

#mapping for fixed coefficients
new_map=list(coeff_re=factor(c("tau.s(ExpShip).1","tau.s(ExpShip).2",
        rep(NA,N),"nu.s(ExpShip).1","nu.s(ExpShip).2",rep(NA,N),rep(NA,24))),
        coeff_fe=factor(rep(NA,5)),log_lambda=factor(c("tau.s(ExpShip)",NA,"nu.s(ExpShip)",NA,rep(NA,4))))

#Fit baseline
response2_sp<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("log_sigma_obs0"=log(0.05)),fixpar=c("mu1","mu2"))

new_coeff_re=c(rep(0,2),baseline2$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline2$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(1,baseline2$lambda()["tau.s(ID)",1],1,baseline2$lambda()["nu.s(ID)",1],baseline2$lambda()[3:6,1])

response2_sp$update_map(new_map)
response2_sp$update_coeff_re(new_coeff_re)
response2_sp$update_coeff_fe(baseline2$coeff_fe()[,1])
response2_sp$update_lambda(new_lambda)

response2_sp$fit()

# Fit response with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response2_sp_50m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"))

new_coeff_re=c(rep(0,2),baseline2_50m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline2_50m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2_50m$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2_50m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2_50m$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(1,baseline2_50m$lambda()["tau.s(ID)",1],1,baseline2_50m$lambda()["nu.s(ID)",1],baseline2_50m$lambda()[3:6,1])

response2_sp_50m$update_map(new_map)
response2_sp_50m$update_coeff_re(new_coeff_re)
response2_sp_50m$update_coeff_fe(baseline2_50m$coeff_fe()[,1])

response2_sp_50m$update_lambda(new_lambda)

response2_sp_50m$fit()

# Fit response with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response2_sp_75m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"))

new_coeff_re=c(rep(0,2),baseline2_75m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline2_75m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2_75m$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2_75m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2_75m$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(1,baseline2_75m$lambda()["tau.s(ID)",1],1,baseline2_75m$lambda()["nu.s(ID)",1],baseline2_75m$lambda()[3:6,1])

response2_sp_75m$update_map(new_map)
response2_sp_75m$update_coeff_re(new_coeff_re)
response2_sp_75m$update_coeff_fe(baseline2_75m$coeff_fe()[,1])

response2_sp_75m$update_lambda(new_lambda)

response2_sp_75m$fit()

# Fit response with 100m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response2_sp_100m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"))

new_coeff_re=c(rep(0,2),baseline2_100m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline2_100m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2_100m$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2_100m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2_100m$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(1,baseline2_100m$lambda()["tau.s(ID)",1],1,baseline2_100m$lambda()["nu.s(ID)",1],baseline2_100m$lambda()[3:6,1])

response2_sp_100m$update_map(new_map)
response2_sp_100m$update_coeff_re(new_coeff_re)
response2_sp_100m$update_coeff_fe(baseline2_100m$coeff_fe()[,1])

response2_sp_100m$update_lambda(new_lambda)

response2_sp_100m$fit()
```

```{r get results response2}

estimates_res2_sp=as.list(response2_sp$tmb_rep(),what="Est")
std_res2_sp=as.list(response2_sp$tmb_rep(),what="Std")

estimates_res2_sp_50m=as.list(response2_sp_50m$tmb_rep(),what="Est")
std_res2_sp_50m=as.list(response2_sp_50m$tmb_rep(),what="Std")

estimates_res2_sp_75m=as.list(response2_sp_75m$tmb_rep(),what="Est")
std_res2_sp_75m=as.list(response2_sp_75m$tmb_rep(),what="Std")

estimates_res2_sp_100m=as.list(response2_sp_100m$tmb_rep(),what="Est")
std_res2_sp_100m=as.list(response2_sp_100m$tmb_rep(),what="Std")

#plot parameters
xmin=list("AngleNormal"=-pi,"ExpShip"=1/45)
xmax=list("AngleNormal"=pi,"ExpShip"=1/3)
link=list("ExpShip"=(\(x) 1/x))
xlabel=list("ExpShip"="Distance to ship")

plot_res2_sp=response2_sp$get_all_plots(baseline=baseline2,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")
plot_res2_sp_50m=response2_sp_50m$get_all_plots(baseline=baseline2_50m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")
plot_res2_sp_75m=response2_sp_75m$get_all_plots(baseline=baseline2_75m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")
plot_res2_sp_100m=response2_sp_100m$get_all_plots(baseline=baseline2_100m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")
```

```{r display plots response2_sp}

#print plots
plot_res2_sp
plot_res2_sp_50m
plot_res2_sp_75m
plot_res2_sp_100m

```

```{r}

post_coeff_res2_sp=response2_sp$post_coeff(n_post=1000)
post_par_res2_sp=list("tau0"=exp(post_coeff_res2_sp$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res2_sp$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res2_sp$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res2_sp$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res2_sp$log_lambda[,2])))
mean_par_res2_sp=lapply(post_par_res2_sp,mean)
sd_par_res2_sp=lapply(post_par_res2_sp,sd)

post_coeff_res2_sp_50m=response2_sp_50m$post_coeff(n_post=1000)
post_par_res2_sp_50m=list("tau0"=exp(post_coeff_res2_sp_50m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res2_sp_50m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res2_sp_50m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res2_sp_50m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res2_sp_50m$log_lambda[,2])))
mean_par_res2_sp_50m=lapply(post_par_res2_sp_50m,mean)
sd_par_res2_sp_50m=lapply(post_par_res2_sp_50m,sd)

post_coeff_res2_sp_75m=response2_sp_75m$post_coeff(n_post=1000)
post_par_res2_sp_75m=list("tau0"=exp(post_coeff_res2_sp_75m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res2_sp_75m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res2_sp_75m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res2_sp_75m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res2_sp_75m$log_lambda[,2])))
mean_par_res2_sp_75m=lapply(post_par_res2_sp_75m,mean)
sd_par_res2_sp_75m=lapply(post_par_res2_sp_75m,sd)

post_coeff_res2_sp_100m=response2_sp_100m$post_coeff(n_post=1000)
post_par_res2_sp_100m=list("tau0"=exp(post_coeff_res2_sp_100m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res2_sp_100m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res2_sp_100m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res2_sp_100m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res2_sp_100m$log_lambda[,2])))
mean_par_res2_sp_100m=lapply(post_par_res2_sp_100m,mean)
sd_par_res2_sp_100m=lapply(post_par_res2_sp_100m,sd)

```

### Response model with splines of ExpShip based on baseline3

```{r fit response3}

#initialize parameters
par0=c(0,0,1.5,4.5,0)

#define model
formulas <- list(mu1=~1,mu2=~1,tau =~s(ExpShip,k=3,bs="cs")+s(ID,bs="re"),
                 nu=~s(ExpShip,k=3,bs="cs")+s(ID,bs="re"),
                 omega=~ti(AngleNormal,k=5,bs="cs")+ti(ExpShore,k=5,bs="cs")+ti(AngleNormal,ExpShore,k=c(5,5),bs="cs"))

#mapping for fixed coeffs
new_map=list(coeff_re=factor(c("tau.s(ExpShip).1","tau.s(ExpShip).2",rep(NA,N),
                    "nu.s(ExpShip).1","nu.s(ExpShip).2",rep(NA,N),rep(NA,24))),coeff_fe=factor(rep(NA,5)),
                    log_lambda=factor(c("tau.s(ExpShip)",NA,"nu.s(ExpShip)",NA,rep(NA,4))))

#Fit response
response3_sp<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("log_sigma_obs0"=log(0.07)))

new_coeff_re=c(rep(0,2),baseline3$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline3$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline3$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline3$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline3$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(1,baseline3$lambda()["tau.s(ID)",1],1,baseline3$lambda()["nu.s(ID)",1],baseline3$lambda()[3:6,1])

response3_sp$update_map(new_map)
response3_sp$update_coeff_re(new_coeff_re)
response3_sp$update_coeff_fe(baseline3$coeff_fe()[,1])

response3_sp$update_lambda(new_lambda)
response3_sp$fit()

#Fit response with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_sp_50m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_re=c(rep(0,2),baseline3_50m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline3_50m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline3_50m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline3_50m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline3_50m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(1,baseline3_50m$lambda()["tau.s(ID)",1],1,baseline3_50m$lambda()["nu.s(ID)",1],baseline3_50m$lambda()[3:6,1])

response3_sp_50m$update_map(new_map)
response3_sp_50m$update_coeff_re(new_coeff_re)
response3_sp_50m$update_coeff_fe(baseline3_50m$coeff_fe()[,1])

response3_sp_50m$update_lambda(new_lambda)
response3_sp_50m$fit()

#Fit response with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_sp_75m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_re=c(rep(0,2),baseline3_75m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline3_75m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline3_75m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline3_75m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline3_75m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(1,baseline3_75m$lambda()["tau.s(ID)",1],1,baseline3_75m$lambda()["nu.s(ID)",1],baseline3_75m$lambda()[3:6,1])

response3_sp_75m$update_map(new_map)
response3_sp_75m$update_coeff_re(new_coeff_re)
response3_sp_75m$update_coeff_fe(baseline3_75m$coeff_fe()[,1])

response3_sp_75m$update_lambda(new_lambda)
response3_sp_75m$fit()

#Fit response with 100m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_sp_100m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_re=c(rep(0,2),baseline3_100m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline3_100m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline3_100m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline3_100m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline3_100m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(1,baseline3_100m$lambda()["tau.s(ID)",1],1,baseline3_100m$lambda()["nu.s(ID)",1],baseline3_100m$lambda()[3:6,1])

response3_sp_100m$update_map(new_map)
response3_sp_100m$update_coeff_re(new_coeff_re)
response3_sp_100m$update_coeff_fe(baseline3_100m$coeff_fe()[,1])

response3_sp_100m$update_lambda(new_lambda)
response3_sp_100m$fit()
```

```{r get results response3}

estimates_res3_sp=as.list(response3_sp$tmb_rep(),what="Est")
std_res3_sp=as.list(response3_sp$tmb_rep(),what="Std")

estimates_res3_sp_50m=as.list(response3_sp_50m$tmb_rep(),what="Est")
std_res3_sp_50m=as.list(response3_sp_50m$tmb_rep(),what="Std")

estimates_res3_sp_75m=as.list(response3_sp_75m$tmb_rep(),what="Est")
std_res3_sp_75m=as.list(response3_sp_75m$tmb_rep(),what="Std")

estimates_res3_sp_100m=as.list(response3_sp_100m$tmb_rep(),what="Est")
std_res3_sp_100m=as.list(response3_sp_100m$tmb_rep(),what="Std")

#plot parameters
xmin=list("ExpShore"=1/D_up,"AngleNormal"=-pi,"ExpShip"=1/45)
xmax=list("ExpShore"=1/D_low,"AngleNormal"=pi,"ExpShip"=1/3)
link=list("ExpShore"=(\(x) 1/x),"ExpShip"=(\(x) 1/x))
xlabel=list("ExpShore"="Distance to shore","ExpShip"="Distance to ship")

plot_res3_sp=response3_sp$get_all_plots(baseline=baseline3,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")
plot_res3_sp_50m=response3_sp_50m$get_all_plots(baseline=baseline3_50m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")
plot_res3_sp_75m=response3_sp_75m$get_all_plots(baseline=baseline3_75m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")
plot_res3_sp_100m=response3_sp_100m$get_all_plots(baseline=baseline3_100m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")

```

```{r display plots response3_sp}

#print plots
plot_res3_sp
plot_res3_sp_50m
plot_res3_sp_75m
plot_res3_sp_100m

```

```{r estimates on parameter scale response3}

post_coeff_res3_sp=response3_sp$post_coeff(n_post=1000)
post_par_res3_sp=list("tau0"=exp(post_coeff_res3_sp$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res3_sp$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res3_sp$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res3_sp$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res3_sp$log_lambda[,2])))
mean_par_res3_sp=lapply(post_par_res3_sp,mean)
sd_par_res3_sp=lapply(post_par_res3_sp,sd)

post_coeff_res3_sp_50m=response3_sp_50m$post_coeff(n_post=1000)
post_par_res3_sp_50m=list("tau0"=exp(post_coeff_res3_sp_50m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res3_sp_50m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res3_sp_50m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res3_sp_50m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res3_sp_50m$log_lambda[,2])))
mean_par_res3_sp_50m=lapply(post_par_res3_sp_50m,mean)
sd_par_res3_sp_50m=lapply(post_par_res3_sp_50m,sd)

post_coeff_res3_sp_75m=response3_sp_75m$post_coeff(n_post=1000)
post_par_res3_sp_75m=list("tau0"=exp(post_coeff_res3_sp_75m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res3_sp_75m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res3_sp_75m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res3_sp_75m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res3_sp_75m$log_lambda[,2])))
mean_par_res3_sp_75m=lapply(post_par_res3_sp_75m,mean)
sd_par_res3_sp_75m=lapply(post_par_res3_sp_75m,sd)

post_coeff_res3_sp_100m=response3_sp_100m$post_coeff(n_post=1000)
post_par_res3_sp_100m=list("tau0"=exp(post_coeff_res3_sp_100m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res3_sp_100m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res3_sp_100m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res3_sp_100m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res3_sp_100m$log_lambda[,2])))
mean_par_res3_sp_100m=lapply(post_par_res3_sp_100m,mean)
sd_par_res3_sp_100m=lapply(post_par_res3_sp_100m,sd)

```

### Response model with exponential of ExpShip based on baseline2

```{r}

par0=c(0,0,1,4,0)

#define model
formulas <- list(mu1=~1,mu2=~1,tau =~ExpShip+s(ID,bs="re"),
                 nu=~ExpShip+s(ID,bs="re"),
                 omega=~ti(DistanceShore,k=5,bs="cs")+ti(AngleNormal,k=5,bs="cs")
                 +ti(AngleNormal,DistanceShore,k=c(5,5),bs="cs"))

new_map=list(coeff_re=factor(c(rep(NA,12),rep(NA,24))),coeff_fe=factor(c(NA,NA,NA,1,NA,2,NA)),
                             log_lambda=factor(c(NA,NA,NA,NA,NA,NA)))

# Fit response
response2_exp<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),
                    other_data=list("log_sigma_obs0"=log(0.05)))

new_coeff_fe=c(baseline2$coeff_fe()[1:3,1],0,baseline2$coeff_fe()[4,1],0,baseline2$coeff_fe()[5,1])
new_coeff_re=c(baseline2$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],baseline2$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(baseline2$lambda()["tau.s(ID)",1],baseline2$lambda()["nu.s(ID)",1],baseline2$lambda()[3:6,1])

response2_exp$update_map(new_map)
response2_exp$update_coeff_re(new_coeff_re)
response2_exp$update_coeff_fe(new_coeff_fe)

response2_exp$update_lambda(new_lambda)

response2_exp$fit()

# Fit response with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response2_exp_50m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_fe=c(baseline2_50m$coeff_fe()[1:3,1],0,baseline2_50m$coeff_fe()[4,1],0,baseline2_50m$coeff_fe()[5,1])
new_coeff_re=c(baseline2_50m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],baseline2_50m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2_50m$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2_50m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2_50m$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(baseline2_50m$lambda()["tau.s(ID)",1],baseline2_50m$lambda()["nu.s(ID)",1],baseline2_50m$lambda()[3:6,1])

response2_exp_50m$update_map(new_map)
response2_exp_50m$update_coeff_re(new_coeff_re)
response2_exp_50m$update_coeff_fe(new_coeff_fe)

response2_exp_50m$update_lambda(new_lambda)

response2_exp_50m$fit()

# Fit response with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response2_exp_75m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_fe=c(baseline2_75m$coeff_fe()[1:3,1],0,baseline2_75m$coeff_fe()[4,1],0,baseline2_75m$coeff_fe()[5,1])
new_coeff_re=c(baseline2_75m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],baseline2_75m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2_75m$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2_75m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2_75m$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(baseline2_75m$lambda()["tau.s(ID)",1],baseline2_75m$lambda()["nu.s(ID)",1],baseline2_75m$lambda()[3:6,1])

response2_exp_75m$update_map(new_map)
response2_exp_75m$update_coeff_re(new_coeff_re)
response2_exp_75m$update_coeff_fe(new_coeff_fe)
response2_exp_75m$update_lambda(new_lambda)

response2_exp_75m$fit()

# Fit response with 100m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response2_exp_100m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_fe=c(baseline2_100m$coeff_fe()[1:3,1],0,baseline2_100m$coeff_fe()[4,1],0,baseline2_100m$coeff_fe()[5,1])
new_coeff_re=c(baseline2_100m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],baseline2_100m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2_100m$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2_100m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2_100m$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(baseline2_100m$lambda()["tau.s(ID)",1],baseline2_100m$lambda()["nu.s(ID)",1],baseline2_100m$lambda()[3:6,1])

response2_exp_100m$update_map(new_map)
response2_exp_100m$update_coeff_re(new_coeff_re)
response2_exp_100m$update_coeff_fe(new_coeff_fe)

response2_exp_100m$update_lambda(new_lambda)

response2_exp_100m$fit()
```

```{r}

estimates_res2_exp=as.list(response2_exp$tmb_rep(),what="Est")
std_res2_exp=as.list(response2_exp$tmb_rep(),what="Std")

estimates_res2_exp_50m=as.list(response2_exp_50m$tmb_rep(),what="Est")
std_res2_exp_50m=as.list(response2_exp_50m$tmb_rep(),what="Std")

estimates_res2_exp_75m=as.list(response2_exp_75m$tmb_rep(),what="Est")
std_res2_exp_75m=as.list(response2_exp_75m$tmb_rep(),what="Std")

estimates_res2_exp_100m=as.list(response2_exp_100m$tmb_rep(),what="Est")
std_res2_exp_100m=as.list(response2_exp_100m$tmb_rep(),what="Std")

#plot parameters
xmin=list("ExpShore"=1/D_up,"AngleNormal"=-pi,"ExpShip"=1/60)
xmax=list("ExpShore"=1/D_low,"AngleNormal"=pi,"ExpShip"=1/5)
link=list("ExpShore"=(\(x) 1/x),"ExpShip"=(\(x) 1/x))
xlabel=list("ExpShore"="Distance to shore","ExpShip"="Distance to ship")

plots_res2_exp=response2_exp$get_all_plots(baseline=baseline2,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")

plots_res2_exp_50m=response2_exp_50m$get_all_plots(baseline=baseline2_50m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")

plots_res2_exp_75m=response2_exp_75m$get_all_plots(baseline=baseline2_75m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")

plots_res2_exp_100m=response2_exp_100m$get_all_plots(baseline=baseline2_100m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")

```

```{r display plots response3_sp}

#print plots
plots_res2_exp
plots_res2_exp_50m
plots_res2_exp_75m
plots_res2_exp_100m

```

### Response model with exponential of ExpShip based on baseline3

```{r}

par0=c(0,0,1,4.5,0)

#define model
formulas <- list(mu1=~1,mu2=~1,tau =~ExpShip+s(ID,bs="re"),
                 nu=~ExpShip+s(ID,bs="re"),
                 omega=~ti(AngleNormal,k=5,bs="cs")+ti(ExpShore,k=5,bs="cs")+ti(AngleNormal,ExpShore,k=c(5,5),bs="cs"))

#mapping for fixed coeffs
new_map=list(coeff_re=factor(c(rep(NA,12),rep(NA,24))),coeff_fe=factor(c(NA,NA,NA,1,NA,2,NA)),
                             log_lambda=factor(c(NA,NA,NA,NA,NA,NA)))

#Fit response
sigma_obs=exp(estimates_bas3$log_sigma_obs)
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_exp<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_fe=c(baseline3$coeff_fe()[1:3,1],0,baseline3$coeff_fe()[4,1],0,baseline3$coeff_fe()[5,1])
new_coeff_re=c(baseline3$coeff_re()[paste("tau.s(ID).",1:6,sep=""),1],baseline3$coeff_re()[paste("nu.s(ID).",1:6,sep=""),1],
               baseline3$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline3$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline3$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(baseline3$lambda()["tau.s(ID)",1],baseline3$lambda()["nu.s(ID)",1],baseline3$lambda()[3:6,1])

response3_exp$update_map(new_map)
response3_exp$update_coeff_re(new_coeff_re)
response3_exp$update_coeff_fe(new_coeff_fe)

response3_exp$update_lambda(new_lambda)
response3_exp$fit()

# Fit response with 30m measurement error
sigma_obs=0.03
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_exp_30m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_fe=c(baseline3_30m$coeff_fe()[1:3,1],0,baseline3_30m$coeff_fe()[4,1],0,baseline3_30m$coeff_fe()[5,1])
new_coeff_re=c(baseline3_30m$coeff_re()[paste("tau.s(ID).",1:6,sep=""),1],baseline3_30m$coeff_re()[paste("nu.s(ID).",1:6,sep=""),1],
               baseline3_30m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline3_30m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline3_30m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(baseline3_30m$lambda()["tau.s(ID)",1],baseline3_30m$lambda()["nu.s(ID)",1],baseline3_30m$lambda()[3:6,1])

response3_exp_30m$update_map(new_map)
response3_exp_30m$update_coeff_re(new_coeff_re)
response3_exp_30m$update_coeff_fe(new_coeff_fe)

response3_exp_30m$update_lambda(new_lambda)
response3_exp_30m$fit()

# Fit response with 40m measurement error
sigma_obs=0.04
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_exp_40m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_fe=c(baseline3_40m$coeff_fe()[1:3,1],0,baseline3_40m$coeff_fe()[4,1],0,baseline3_40m$coeff_fe()[5,1])
new_coeff_re=c(baseline3_40m$coeff_re()[paste("tau.s(ID).",1:6,sep=""),1],baseline3_40m$coeff_re()[paste("nu.s(ID).",1:6,sep=""),1],
               baseline3_40m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline3_40m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline3_40m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(baseline3_40m$lambda()["tau.s(ID)",1],baseline3_40m$lambda()["nu.s(ID)",1],baseline3_40m$lambda()[3:6,1])

response3_exp_40m$update_map(new_map)
response3_exp_40m$update_coeff_re(new_coeff_re)
response3_exp_40m$update_coeff_fe(new_coeff_fe)

response3_exp_40m$update_lambda(new_lambda)
response3_exp_40m$fit()

# Fit response with 45m measurement error
sigma_obs=0.045
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_exp_45m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_fe=c(baseline3_45m$coeff_fe()[1:3,1],0,baseline3_45m$coeff_fe()[4,1],0,baseline3_45m$coeff_fe()[5,1])
new_coeff_re=c(baseline3_45m$coeff_re()[paste("tau.s(ID).",1:6,sep=""),1],baseline3_45m$coeff_re()[paste("nu.s(ID).",1:6,sep=""),1],
               baseline3_45m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline3_45m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline3_45m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(baseline3_45m$lambda()["tau.s(ID)",1],baseline3_45m$lambda()["nu.s(ID)",1],baseline3_45m$lambda()[3:6,1])

response3_exp_45m$update_map(new_map)
response3_exp_45m$update_coeff_re(new_coeff_re)
response3_exp_45m$update_coeff_fe(new_coeff_fe)
response3_exp_45m$update_lambda(new_lambda)

response3_exp_45m$fit()

# Fit response with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_exp_50m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_fe=c(baseline3_50m$coeff_fe()[1:3,1],0,baseline3_50m$coeff_fe()[4,1],0,baseline3_50m$coeff_fe()[5,1])
new_coeff_re=c(baseline3_50m$coeff_re()[paste("tau.s(ID).",1:6,sep=""),1],baseline3_50m$coeff_re()[paste("nu.s(ID).",1:6,sep=""),1],
               baseline3_50m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline3_50m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline3_50m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(baseline3_50m$lambda()["tau.s(ID)",1],baseline3_50m$lambda()["nu.s(ID)",1],baseline3_50m$lambda()[3:6,1])

response3_exp_50m$update_map(new_map)
response3_exp_50m$update_coeff_re(new_coeff_re)
response3_exp_50m$update_coeff_fe(new_coeff_fe)

response3_exp_50m$update_lambda(new_lambda)
response3_exp_50m$fit()

# Fit response with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_exp_75m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_fe=c(baseline3_75m$coeff_fe()[1:3,1],0,baseline3_75m$coeff_fe()[4,1],0,baseline3_75m$coeff_fe()[5,1])
new_coeff_re=c(baseline3_75m$coeff_re()[paste("tau.s(ID).",1:6,sep=""),1],baseline3_75m$coeff_re()[paste("nu.s(ID).",1:6,sep=""),1],
               baseline3_75m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline3_75m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline3_75m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(baseline3_75m$lambda()["tau.s(ID)",1],baseline3_75m$lambda()["nu.s(ID)",1],baseline3_75m$lambda()[3:6,1])

response3_exp_75m$update_map(new_map)
response3_exp_75m$update_coeff_fe(new_coeff_fe)
response3_exp_75m$update_lambda(new_lambda)

response3_exp_75m$fit()

# Fit response with 100m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_exp_100m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_fe=c(baseline3_100m$coeff_fe()[1:3,1],0,baseline3_100m$coeff_fe()[4,1],0,baseline3_100m$coeff_fe()[5,1])
new_coeff_re=c(baseline3_100m$coeff_re()[paste("tau.s(ID).",1:6,sep=""),1],baseline3_100m$coeff_re()[paste("nu.s(ID).",1:6,sep=""),1],
               baseline3_100m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline3_100m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline3_100m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(baseline3_100m$lambda()["tau.s(ID)",1],baseline3_100m$lambda()["nu.s(ID)",1],baseline3_100m$lambda()[3:6,1])

response3_exp_100m$update_map(new_map)
response3_exp_100m$update_coeff_re(new_coeff_re)
response3_exp_100m$update_coeff_fe(new_coeff_fe)
response3_exp_100m$update_lambda(new_lambda)

response3_exp_100m$fit()
```

```{r}

estimates_res3_exp=as.list(response3_exp$tmb_rep(),what="Est")
std_res3_exp=as.list(response3_exp$tmb_rep(),what="Std")

estimates_res3_exp_30m=as.list(response3_exp_30m$tmb_rep(),what="Est")
std_res3_exp_30m=as.list(response3_exp_30m$tmb_rep(),what="Std")

estimates_res3_exp_40m=as.list(response3_exp_40m$tmb_rep(),what="Est")
std_res3_exp_40m=as.list(response3_exp_40m$tmb_rep(),what="Std")

estimates_res3_exp_45m=as.list(response3_exp_45m$tmb_rep(),what="Est")
std_res3_exp_45m=as.list(response3_exp_45m$tmb_rep(),what="Std")

estimates_res3_exp_50m=as.list(response3_exp_50m$tmb_rep(),what="Est")
std_res3_exp_50m=as.list(response3_exp_50m$tmb_rep(),what="Std")

estimates_res3_exp_75m=as.list(response3_exp_75m$tmb_rep(),what="Est")
std_res3_exp_75m=as.list(response3_exp_75m$tmb_rep(),what="Std")

estimates_res3_exp_100m=as.list(response3_exp_100m$tmb_rep(),what="Est")
std_res3_exp_100m=as.list(response3_exp_100m$tmb_rep(),what="Std")

#plot parameters
xmin=list("ExpShore"=1/D_up,"AngleNormal"=-pi,"ExpShip"=1/60)
xmax=list("ExpShore"=1/D_low,"AngleNormal"=pi,"ExpShip"=1/5)
link=list("ExpShore"=(\(x) 1/x),"ExpShip"=(\(x) 1/x))
xlabel=list("ExpShore"="Distance to shore","ExpShip"="Distance to ship")

plots_res3_exp=response3_exp$get_all_plots(baseline=baseline3,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")

plots_res3_exp_30m=response3_exp_30m$get_all_plots(baseline=baseline3_30m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")

plots_res3_exp_40m=response3_exp_40m$get_all_plots(baseline=baseline3_40m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")

plots_res3_exp_50m=response3_exp_50m$get_all_plots(baseline=baseline3_50m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")
plots_res3_exp_75m=response3_exp_75m$get_all_plots(baseline=baseline3_75m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")
plots_res3_exp_100m=response3_exp_100m$get_all_plots(baseline=baseline3_100m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")

```

```{r display plots response3_sp}

#print plots
plots_res3_exp
plots_res3_exp_50m
plots_res3_exp_75m
plots_res3_exp_100m

```

We select the response 3 model with 50m measurement error

```{r}
# Create directory
if (!dir.exists("response3_exp_50m")) {
    dir.create("response3_exp_50m", recursive = TRUE)
}

# Use lapply to save each plot
lapply(seq_along(plots_bas2), function(i) {

    plot<-plots_bas2[[i]]
    plot_name <- names(plots_bas2)[i]
    file_path <- file.path("response3_exp_50m", plot_name)

    # Check if the plot is a ggplot or plotly object
    if (inherits(plot, "ggplot")) {
      ggsave(file_path, plot, device = "pdf")
    } else if (inherits(plot, "plotly")) {
      saveWidget(plot, file_path)
  }
    })
```
### Plots improvements

```{r}

#get angular velocity for different distances to shore
plot_me_tau <- plots_res3_exp_50m$me_tau_ExpShip
plot_me_nu <- plots_res3_exp_50m$me_nu_ExpShip

# Update plot_me_tau
plot_me_tau <- plot_me_tau +
  labs(y = expression(tau)) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size=18),
    axis.title.y = element_text(size=18, angle=0, vjust=0.5, margin = margin(r = 10)),
    axis.text.x = element_text(size=14),
    axis.text.y = element_text(size=14),
    legend.title = element_text(size=14),
    legend.text = element_text(size=14)
  )

# Update plot_me_nu
plot_me_nu <- plot_me_nu +
  labs(y = expression(nu)) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size=18),
    axis.title.y = element_text(size=18, angle=0, vjust=0.5, margin = margin(r = 10)),
    axis.text.x = element_text(size=14),
    axis.text.y = element_text(size=14),
    legend.title = element_text(size=14),
    legend.text = element_text(size=14)
  )

# Save the updated plots
ggsave("plot_me_tau_updated.pdf", plot=plot_me_tau, width=10, height=6, units="in")
ggsave("plot_me_nu_updated.pdf", plot=plot_me_nu, width=10, height=6, units="in")

```

## Posterior samples
```{r}
post_coeff_res3_exp_50m=response3_exp_50m$post_coeff(n_post=10000)
post_par_res3_exp_50m=list(
              "tau0"=exp(post_coeff_res3_exp_50m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res3_exp_50m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res3_exp_50m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res3_exp_50m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res3_exp_50m$log_lambda[,2])),
              "tau.ExpShip"=post_coeff_res3_exp_50m$coeff_fe[,"tau.ExpShip"],
              "nu.ExpShip"=post_coeff_res3_exp_50m$coeff_fe[,"nu.ExpShip"])
mean_par_res3=lapply(post_par_res3_exp_50m,mean)
sd_par_res3=lapply(post_par_res3_exp_50m,sd)
quant_par_res3 <- lapply(post_par_res3_exp_50m, quantile, probs = c(0.025, 0.975))

```

#### Thorough uncertainty quantification for response3_exp_50m

```{r}

#posterior samples of the coefficient of the baseline model
post_coeff_re=post_coeff_bas3_50m$coeff_re[1:100,]
post_coeff_fe=post_coeff_bas3_50m$coeff_fe[1:100,]
post_coeff_lambda=exp(post_coeff_bas3_50m$log_lambda[1:100,])

# Fit response with 50m measurement error

par0=c(0,0,1,4.5,0)
formulas <- list(mu1=~1,mu2=~1,tau =~ExpShip+s(ID,bs="re"),
                 nu=~ExpShip+s(ID,bs="re"),
                 omega=~ti(AngleNormal,k=5,bs="cs")+ti(ExpShore,k=5,bs="cs")+ti(AngleNormal,ExpShore,k=c(5,5),bs="cs"))

sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))

estimates=matrix(rep(0,2*nrow(post_coeff_fe)),ncol=2)
std=matrix(rep(0,2*nrow(post_coeff_fe)),ncol=2)

for (i in 1:nrow(post_coeff_fe)) {

    response3_exp_50m_temp<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                      response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H),
                      map=list(coeff_re=factor(c(rep(NA,12),rep(NA,24))),coeff_fe=factor(c(NA,NA,NA,1,NA,2,NA)),
                               log_lambda=factor(c(NA,NA,NA,NA,NA,NA))))

    new_coeff_fe=c(post_coeff_fe[i,1:3],0,post_coeff_fe[i,4],0,post_coeff_fe[i,5])
    new_coeff_re=post_coeff_re[i,]
    new_lambda=post_coeff_lambda[i,]
    response3_exp_50m_temp$update_coeff_re(new_coeff_re)
    response3_exp_50m_temp$update_coeff_fe(new_coeff_fe)

    response3_exp_50m_temp$update_lambda(new_lambda)
    response3_exp_50m_temp$fit()

    est_alpha=as.list(response3_exp_50m_temp$tmb_rep(),what="Est")$coeff_fe[c("tau.ExpShip","nu.ExpShip"),1]
    std_alpha=as.list(response3_exp_50m_temp$tmb_rep(),what="Std")$coeff_fe[c("tau.ExpShip","nu.ExpShip"),1]

    estimates[i,]=est_alpha
    std[i,]=std_alpha
}

colnames(estimates)=c("tau.ExpShip","nu.ExpShip")
colnames(std)=c("tau.ExpShip","nu.ExpShip")

```

```{r}
#The 9th value in estimates is inconsistent, we drop it
estimates=estimates[-9,]
std=std[-9,]

mean_est=apply(X=estimates,MARGIN=2,FUN=mean)
mean_std=apply(X=na.omit(std),MARGIN=2,FUN=mean)
```