### 2024-11-08 R/application/CVM/fit_cvm.Rmd

---
title: "fit_crcvm"
author: "Alexandre Delporte"
date: "2024-06-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up

```{r packages}
set.seed(42)

library(smoothSDE)
library(ggplot2)
library(dplyr)
library(plotly)
library(htmlwidgets)
library(mgcv)
```

```{r get data}

# Set the path to the directory containing the data
par_dir=dirname(dirname(dirname(getwd()))) #parent directory 
narwhal_data_path <- file.path(par_dir,"Data", "Narwhals")

#data before exposure with first 24h removed
dataBE1=read.csv(file.path(narwhal_data_path,"DataBE1.csv"), header = TRUE,dec = ".")

#data before exposure with first 12h removed
dataBE2=read.csv(file.path(narwhal_data_path,"DataBE2.csv"), header = TRUE,dec = ".")

#data after exposure
dataAE=read.csv(file.path(narwhal_data_path,"DataAE.csv"), header = TRUE,dec = ".")

```

We choose thresholds on the distance to the shore to define the ExpShore covariate.

```{r preprocess data}

D_low=0.07
D_up=3
dataBE1[dataBE1$DistanceShore> D_up,"ExpShore"]=0
dataBE1[dataBE1$DistanceShore< D_low,"ExpShore"]=1/D_low

dataBE2[dataBE2$DistanceShore> D_up,"ExpShore"]=0
dataBE2[dataBE2$DistanceShore< D_low,"ExpShore"]=1/D_low

dataAE[dataAE$DistanceShore> D_up,"ExpShore"]=0
dataAE[dataAE$DistanceShore< D_low,"ExpShore"]=1/D_low

N=length(unique(dataBE2$ID))

n_pre=length(dataBE2$time) #number of observations before exposure
n_post=length(dataAE$time) #number of observations after exposure
```

## Baseline models

We can estimate the measurement error standard deviation directly from the data. Based on accuracy studies about fastloc GPS, it should be a few tens of metres. The estimate we get from the data before exposure is 25m (std error 1m), which seems consistent with previous studies. However, when estimating this measurement error from the post exposure data, we get a higher value of 48m, which is weird : the same GPS is device is used during all the experiment, so measurement error before or after exposure should be about the same. If we fix the measurement error to 25m when estimating the response, then we can't estimate because TMB gets a non finite value of the negative log-likelihood at initial parameters. This cannot be solved by changing the other initial parameters in the optimisation. Thus we decided to fix the measurement error to different values in the baseline and the response and compare the results we get. Higher measurement error gives higher estimates for tau, whereas the estimate of nu is not affected by the changes in the measurement error.

### Baselines with constant parameters

```{r fit baseline0}

#initial parameters
par0 <- c(0,0,1,4,0)

#model formula
formulas <- list(mu1=~1,mu2=~1,tau =~s(ID,bs="re"),nu=~s(ID,bs="re"),omega=~1)

# Fit baseline with 25m measurement error
sigma_obs=0.025
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline0_25m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"))
baseline0_25m$fit()

# Fit baseline with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline0_50m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"))
baseline0_50m$fit()

# Fit baseline with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline0_75m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"))
baseline0_75m$fit()

# Fit baseline with 100m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline0_100m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"))
baseline0_100m$fit()

```

```{r results baseline0}
estimates_bas0_25m=as.list(baseline0_25m$tmb_rep(),what="Est")
std_bas0_25m=as.list(baseline0_25m$tmb_rep(),what="Std")

estimates_bas0_50m=as.list(baseline0_50m$tmb_rep(),what="Est")
std_bas0_50m=as.list(baseline0_50m$tmb_rep(),what="Std")

estimates_bas0_75m=as.list(baseline0_75m$tmb_rep(),what="Est")
std_bas0_75m=as.list(baseline0_75m$tmb_rep(),what="Std")

estimates_bas0_100m=as.list(baseline0_100m$tmb_rep(),what="Est")
std_bas0_100m=as.list(baseline0_100m$tmb_rep(),what="Std")

```

### Baseline with AngleNormal covariate in omega

Fixing a high standard deviation (>50m) for the measurement error tends to increase the estimate of the parameter tau. Intuitively, for a low value of the measurement error, tortuous trajectory can only be caused by less persistent movement, whereas if the measurement error is high, a part of the tortuosity can be attributed to the measurement errors, which might give higher estimate of the persistence parameter.

```{r fit baseline1}

#initial parameters
par0 <- c(0,0,1,4,0)
init_lambda=c(1,1,1)

#model formula
formulas <- list(mu1=~1,mu2=~1,
                 tau =~s(ID,bs="re"),
                 nu=~s(ID,bs="re"),
                 omega=~s(AngleNormal,k=5,bs="cs"))

# Fit baseline with 25m measurement error
sigma_obs=0.025
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline1_25m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"),
                    map=list(log_lambda=factor(c("tau.s(ID)","nu.s(ID)",NA))))

baseline1_25m$update_lambda(init_lambda)

baseline1_25m$fit()

# Fit baseline with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline1_50m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"),
                    map=list(log_lambda=factor(c("tau.s(ID)","nu.s(ID)",NA))))

baseline1_50m$update_lambda(init_lambda)

baseline1_50m$fit()

# Fit baseline with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline1_75m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"),
                    map=list(log_lambda=factor(c("tau.s(ID)","nu.s(ID)",NA))))

baseline1_75m$update_lambda(init_lambda)

baseline1_75m$fit()

# Fit baseline with 100m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline1_100m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"),
                    map=list(log_lambda=factor(c("tau.s(ID)","nu.s(ID)",NA))))

baseline1_100m$update_lambda(init_lambda)

baseline1_100m$fit()

```

```{r results baseline1}
estimates_bas1_25m=as.list(baseline1_25m$tmb_rep(),what="Est")
std_bas1_25m=as.list(baseline1_25m$tmb_rep(),what="Std")

estimates_bas1_50m=as.list(baseline1_50m$tmb_rep(),what="Est")
std_bas1_50m=as.list(baseline1_50m$tmb_rep(),what="Std")

estimates_bas1_75m=as.list(baseline1_75m$tmb_rep(),what="Est")
std_bas1_75m=as.list(baseline1_75m$tmb_rep(),what="Std")

estimates_bas1_100m=as.list(baseline1_100m$tmb_rep(),what="Est")
std_bas1_100m=as.list(baseline1_100m$tmb_rep(),what="Std")

plots_bas1_25m=baseline1_25m$get_all_plots(baseline=NULL,model_name="baseline1_25m",show_CI="pointwise",save=TRUE)
plots_bas1_50m=baseline1_50m$get_all_plots(baseline=NULL,model_name="baseline1_50m",show_CI="pointwise",save=TRUE)
plots_bas1_75m=baseline1_75m$get_all_plots(baseline=NULL,model_name="baseline1_75m",show_CI="pointwise",save=TRUE)
plots_bas1_100m=baseline1_100m$get_all_plots(baseline=NULL,model_name="baseline1_100m",show_CI="pointwise",save=TRUE)

```

### Baseline with AngleNormal and DistanceShore in omega

```{r fit baseline2}

#initial parameters
par0 <- c(0,0,1,4,0)
init_lambda=c(1/0.2^2,1/0.2^2,1,0.01,1,1)

formulas <- list(mu1 = ~1 ,mu2 =~1,tau =~s(ID,bs="re"),nu=~s(ID,bs="re"),
                 omega=~ti(DistanceShore,k=5,bs="cs")+ti(AngleNormal,k=5,bs="cs")+ti(AngleNormal,DistanceShore,k=c(5,5),bs="cs"))

# Fit baseline with 25m measurement error
sigma_obs=0.025
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline2_25m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",
                    response = c("x","y"),par0 = par0,other_data=list("H"=H),
                    fixpar=c("mu1","mu2"),map=list("log_lambda"=factor(c(1,2,rep(NA,4)))))

baseline2_25m$update_lambda(init_lambda)
baseline2_25m$fit()

# Fit baseline with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline2_50m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",
                    response = c("x","y"),par0 = par0,other_data=list("H"=H),
                    fixpar=c("mu1","mu2"),map=list("log_lambda"=factor(c(1,2,rep(NA,4)))))

baseline2_50m$update_lambda(init_lambda)
baseline2_50m$fit()

# Fit baseline with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline2_75m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",
                    response = c("x","y"),par0 = par0,other_data=list("H"=H),
                    fixpar=c("mu1","mu2"),map=list("log_lambda"=factor(c(1,2,rep(NA,4)))))

baseline2_75m$update_lambda(init_lambda)
baseline2_75m$fit()

# Fit baseline with 100m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline2_100m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",
                    response = c("x","y"),par0 = par0,other_data=list("H"=H),
                    fixpar=c("mu1","mu2"),map=list("log_lambda"=factor(c(1,2,rep(NA,4)))))

baseline2_100m$update_lambda(init_lambda)
baseline2_100m$fit()

```

```{r get results baseline2}

estimates_bas2_25m=as.list(baseline2_25m$tmb_rep(),what="Est")
std_bas2_25m=as.list(baseline2_25m$tmb_rep(),what="Std")

estimates_bas2_50m=as.list(baseline2_50m$tmb_rep(),what="Est")
std_bas2_50m=as.list(baseline2_50m$tmb_rep(),what="Std")

estimates_bas2_75m=as.list(baseline2_75m$tmb_rep(),what="Est")
std_bas2_75m=as.list(baseline2_75m$tmb_rep(),what="Std")

estimates_bas2_100m=as.list(baseline2_100m$tmb_rep(),what="Est")
std_bas2_100m=as.list(baseline2_100m$tmb_rep(),what="Std")

#plot parameters

xmin=list("DistanceShore"=0.07)
xmax=list("DistanceShore"=5)

plots_bas2_25m=baseline2_25m$get_all_plots(baseline=NULL,model_name="baseline2_25m",xmin=xmin,xmax=xmax,show_CI="pointwise",save=TRUE)
plots_bas2_50m=baseline2_50m$get_all_plots(baseline=NULL,model_name="baseline2_50m",xmin=xmin,xmax=xmax,show_CI="pointwise",save=TRUE)
plots_bas2_75m=baseline2_75m$get_all_plots(baseline=NULL,model_name="baseline2_75m",xmin=xmin,xmax=xmax,show_CI="pointwise",save=TRUE)
plots_bas2_100m=baseline2_100m$get_all_plots(baseline=NULL,model_name="baseline2_100m",xmin=xmin,xmax=xmax,show_CI="pointwise",save=TRUE)
```

```{r estimates on parameter scale baseline2}

post_coeff_bas2_25m=baseline2_25m$post_coeff(n_post=1000)
post_par_bas2_25m=list("tau0"=exp(post_coeff_bas2_25m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_bas2_25m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_bas2_25m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_bas2_25m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_bas2_25m$log_lambda[,2])))
mean_par_bas2_25m=lapply(post_par_bas2_25m,mean)
sd_par_bas2_25m=lapply(post_par_bas2_25m,sd)

post_coeff_bas2_50m=baseline2_50m$post_coeff(n_post=1000)
post_par_bas2_50m=list("tau0"=exp(post_coeff_bas2_50m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_bas2_50m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_bas2_50m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_bas2_50m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_bas2_50m$log_lambda[,2])))
mean_par_bas2_50m=lapply(post_par_bas2_50m,mean)
sd_par_bas2_50m=lapply(post_par_bas2_50m,sd)

post_coeff_bas2_75m=baseline2_75m$post_coeff(n_post=1000)
post_par_bas2_75m=list("tau0"=exp(post_coeff_bas2_75m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_bas2_75m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_bas2_75m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_bas2_75m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_bas2_75m$log_lambda[,2])))
mean_par_bas2_75m=lapply(post_par_bas2_75m,mean)
sd_par_bas2_75m=lapply(post_par_bas2_75m,sd)

#get 95% confidence intervals
post_coeff_bas2_100m=baseline2_100m$post_coeff(n_post=1000)
post_par_bas2_100m=list("tau0"=exp(post_coeff_bas2_100m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_bas2_100m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_bas2_100m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_bas2_100m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_bas2_100m$log_lambda[,2])))
mean_par_bas2_100m=lapply(post_par_bas2_100m,mean)
sd_par_bas2_100m=lapply(post_par_bas2_100m,sd)
```

```{r smooth omega baseline2}

#get angular velocity for different distances to shore

plot1=plots_bas2_50m$fe_omega_AngleNormal_q1_DistanceShore
plot2=plots_bas2_50m$fe_omega_AngleNormal_q2_DistanceShore
plot4=plots_bas2_50m$fe_omega_AngleNormal_q4_DistanceShore

omega1 <- ggplot_build(plot1)$data[[1]]
omega2 <- ggplot_build(plot2)$data[[1]]
omega4 <- ggplot_build(plot4)$data[[1]]

omega1$Category <- "Close"
omega2$Category <- "Medium"
omega4$Category <- "Far"

# Combine the data frames
combined_omega <- rbind(omega1, omega2, omega4)

ci_omega1 <- ggplot_build(plot1)$data[[2]]
ci_omega2 <- ggplot_build(plot2)$data[[2]]
ci_omega4 <- ggplot_build(plot4)$data[[2]]

ci_omega1$Category <- "Close"
ci_omega2$Category <- "Medium"
ci_omega4$Category <- "Far"

# Combine the data frames
combined_ci_omega <- rbind(ci_omega1, ci_omega2, ci_omega4)

combined_plot <- ggplot() +
  geom_line(data=combined_omega,aes(x=x,y = y,color = Category, linetype = Category)) +
  geom_ribbon(data=combined_ci_omega,aes(x=x,ymin=ymin,ymax=ymax,fill=Category),alpha=0.1)+
  labs(color = "Category", linetype = "Category",x="Theta",y="Omega") +
  geom_vline(xintercept = c(-pi/2, pi/2), linetype = "dashed", color = "grey")+
  annotate("text", x = -pi/2, y = Inf, label = expression(-pi/2), vjust = 40, color = "black") +
  annotate("text", x = pi/2, y = Inf, label = expression(pi/2), vjust = 40, color = "black")+
  theme_minimal()

ggsave("omega_DistanceShore_levels.png",plot=combined_plot,path="baseline2_50m",width = 10, height = 6, units = "in")

```

### Baseline with AngleNormal and ExpShore in omega

```{r fit baseline3}

#initial parameters
par0 <- c(0,0,1,4,0)
init_lambda=c(1,1,0.01,1,1,1)

formulas <- list(mu1 = ~1 ,mu2 =~1,tau =~s(ID,bs="re"),nu=~s(ID,bs="re"),
                   omega=~ti(AngleNormal,k=5,bs="cs")+ti(ExpShore,k=5,bs="cs")+ti(AngleNormal,ExpShore,k=c(5,5),bs="cs"))

# Fit baseline with 25m measurement error
sigma_obs=0.025
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline3_25m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",
                    response = c("x","y"),par0 = par0,other_data=list("H"=H),
                    fixpar=c("mu1","mu2"),map=list("log_lambda"=factor(c(1,2,rep(NA,4)))))

baseline3_25m$update_lambda(init_lambda)
baseline3_25m$fit()

# Fit baseline with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline3_50m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",
                    response = c("x","y"),par0 = par0,other_data=list("H"=H),
                    fixpar=c("mu1","mu2"),map=list("log_lambda"=factor(c(1,2,rep(NA,4)))))

baseline3_50m$update_lambda(init_lambda)
baseline3_50m$fit()

# Fit baseline with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline3_75m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",
                    response = c("x","y"),par0 = par0,other_data=list("H"=H),
                    fixpar=c("mu1","mu2"),map=list("log_lambda"=factor(c(1,2,rep(NA,4)))))

baseline3_75m$update_lambda(init_lambda)
baseline3_75m$fit()

# Fit baseline with 100m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_pre),dim=c(2,2,n_pre))

baseline3_100m<- SDE$new(formulas = formulas,data = dataBE2,type = "RACVM",
                    response = c("x","y"),par0 = par0,other_data=list("H"=H),
                    fixpar=c("mu1","mu2"),map=list("log_lambda"=factor(c(1,2,rep(NA,4)))))

baseline3_100m$update_lambda(init_lambda)
baseline3_100m$fit()
```
```{r get results baseline3}

estimates_bas3_25m=as.list(baseline3_25m$tmb_rep(),what="Est")
std_bas3_25m=as.list(baseline3_25m$tmb_rep(),what="Std")

estimates_bas3_50m=as.list(baseline3_50m$tmb_rep(),what="Est")
std_bas3_50m=as.list(baseline3_50m$tmb_rep(),what="Std")

estimates_bas3_75m=as.list(baseline3_75m$tmb_rep(),what="Est")
std_bas3_75m=as.list(baseline3_75m$tmb_rep(),what="Std")

estimates_bas3_100m=as.list(baseline3_100m$tmb_rep(),what="Est")
std_bas3_100m=as.list(baseline3_100m$tmb_rep(),what="Std")

#plot parameters

xmin=list("ExpShore"=1/D_up)
xmax=list("ExpShore"=1/D_low)
link=list("ExpShore"=(\(x) 1/x))
xlabel=list("ExpShore"="Distance to shore")

plots_bas3_25m=baseline3_25m$get_all_plots(baseline=NULL,model_name="baseline3_25m",
                            xmin=xmin,xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise",save=TRUE)

plots_bas3_50m=baseline3_50m$get_all_plots(baseline=NULL,model_name="baseline3_50m",
                            xmin=xmin,xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise",save=TRUE)

plots_bas3_75m=baseline3_75m$get_all_plots(baseline=NULL,model_name="baseline3_75m",
                            xmin=xmin,xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise",save=TRUE)

plots_bas3_100m=baseline3_100m$get_all_plots(baseline=NULL,model_name="baseline3_100m",
                            xmin=xmin,xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise",save=TRUE)

```

```{r estimates on parameter scale baseline3}

post_coeff_bas3_25m=baseline3_25m$post_coeff(n_post=1000)
post_par_bas3_25m=list("tau0"=exp(post_coeff_bas3_25m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_bas3_25m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_bas3_25m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_bas3_25m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_bas3_25m$log_lambda[,2])))
mean_par_bas3_25m=lapply(post_par_bas3_25m,mean)
sd_par_bas3_25m=lapply(post_par_bas3_25m,sd)

post_coeff_bas3_50m=baseline3_50m$post_coeff(n_post=1000)
post_par_bas3_50m=list("tau0"=exp(post_coeff_bas3_50m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_bas3_50m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_bas3_50m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_bas3_50m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_bas3_50m$log_lambda[,2])))
mean_par_bas3_50m=lapply(post_par_bas3_50m,mean)
sd_par_bas3_50m=lapply(post_par_bas3_50m,sd)

post_coeff_bas3_75m=baseline3_75m$post_coeff(n_post=1000)
post_par_bas3_75m=list("tau0"=exp(post_coeff_bas3_75m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_bas3_75m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_bas3_75m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_bas3_75m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_bas3_75m$log_lambda[,2])))
mean_par_bas3_75m=lapply(post_par_bas3_75m,mean)
sd_par_bas3_75m=lapply(post_par_bas3_75m,sd)

post_coeff_bas3_100m=baseline3_100m$post_coeff(n_post=1000)
post_par_bas3_100m=list("tau0"=exp(post_coeff_bas3_100m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_bas3_100m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_bas3_100m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_bas3_100m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_bas3_100m$log_lambda[,2])))
mean_par_bas3_100m=lapply(post_par_bas3_100m,mean)
sd_par_bas3_100m=lapply(post_par_bas3_100m,sd)

```

```{r smooth omega baseline3}

plot1=plots_bas3_50m$fe_omega_AngleNormal_q1_ExpShore
plot2=plots_bas3_50m$fe_omega_AngleNormal_q2_ExpShore
plot4=plots_bas3_50m$fe_omega_AngleNormal_q4_ExpShore

omega1 <- ggplot_build(plot1)$data[[1]]
omega2 <- ggplot_build(plot2)$data[[1]]
omega4 <- ggplot_build(plot4)$data[[1]]

omega1$Category <- "Far"
omega2$Category <- "Medium"
omega4$Category <- "Close"

# Combine the data frames
combined_omega <- rbind(omega1, omega2, omega4)

ci_omega1 <- ggplot_build(plot1)$data[[2]]
ci_omega2 <- ggplot_build(plot2)$data[[2]]
ci_omega4 <- ggplot_build(plot4)$data[[2]]

ci_omega1$Category <- "Far"
ci_omega2$Category <- "Medium"
ci_omega4$Category <- "Close"

# Combine the data frames
combined_ci_omega <- rbind(ci_omega1, ci_omega2, ci_omega4)

combined_plot <- ggplot() +
  geom_line(data=combined_omega,aes(x=x,y = y,color = Category, linetype = Category)) +
  geom_ribbon(data=combined_ci_omega,aes(x=x,ymin=ymin,ymax=ymax,fill=Category),alpha=0.1)+
  labs(color = "Category", linetype = "Category",x="Theta",y="Omega") +
  geom_vline(xintercept = c(-pi/2, pi/2), linetype = "dashed", color = "grey")+
  annotate("text", x = -pi/2, y = Inf, label = expression(-pi/2), vjust = 40, color = "black") +
  annotate("text", x = pi/2, y = Inf, label = expression(pi/2), vjust = 40, color = "black")+
  theme_minimal()

ggsave("omega_DistanceShore_levels.png",plot=combined_plot,path="baseline3_50m",width = 10, height = 6, units = "in")

```

## Response models with splines of ExpShip

### Response models based on baseline1

```{r fit response1}

par0 <- c(0,0,1,4,0)

#model formula
formulas <- list(mu1=~1,mu2=~1,
                 tau =~s(ExpShip,k=3,bs="cs")+s(ID,bs="re"),
                 nu=~s(ExpShip,k=3,bs="cs")+s(ID,bs="re"),
                 omega=~s(AngleNormal,k=5,bs="cs"))

# Fit response with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response1_sp_50m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"),
                    map=list(coeff_re=factor(c("tau.s(ExpShip).1","tau.s(ExpShip).2",rep(NA,N),
                                               "nu.s(ExpShip).1","nu.s(ExpShip).2",rep(NA,N),rep(NA,4))),coeff_fe=factor(rep(NA,5)),
                                                                                      log_lambda=factor(c("tau.s(ExpShip)",NA,
                                                                                                          "nu.s(ExpShip)",NA,NA))))

new_coeff_re=c(rep(0,2),baseline1_50m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],rep(0,2),
               baseline1_50m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline1_50m$coeff_re()[paste("omega.s(AngleNormal).",1:4,sep=""),1])
new_lambda=c(1,baseline1_50m$lambda()["tau.s(ID)",1],1,baseline1_50m$lambda()["nu.s(ID)",1],
             baseline1_50m$lambda()["omega.s(AngleNormal)",1])
response1_sp_50m$update_coeff_re(new_coeff_re)
response1_sp_50m$update_coeff_fe(baseline1_50m$coeff_fe()[,1])

response1_sp_50m$update_lambda(new_lambda)

response1_sp_50m$fit()

# Fit response with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response1_sp_75m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"),
                    map=list(coeff_re=factor(c("tau.s(ExpShip).1","tau.s(ExpShip).2",rep(NA,N),
                                               "nu.s(ExpShip).1","nu.s(ExpShip).2",rep(NA,N),rep(NA,4))),coeff_fe=factor(rep(NA,5)),
                                                                                      log_lambda=factor(c("tau.s(ExpShip)",NA,
                                                                                                          "nu.s(ExpShip)",NA,NA))))

new_coeff_re=c(rep(0,2),baseline1_75m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],rep(0,2),
               baseline1_75m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline1_75m$coeff_re()[paste("omega.s(AngleNormal).",1:4,sep=""),1])
new_lambda=c(1,baseline1_75m$lambda()["tau.s(ID)",1],1,baseline1_75m$lambda()["nu.s(ID)",1],
             baseline1_75m$lambda()["omega.s(AngleNormal)",1])
response1_sp_75m$update_coeff_re(new_coeff_re)
response1_sp_75m$update_coeff_fe(baseline1_75m$coeff_fe()[,1])

response1_sp_75m$update_lambda(new_lambda)

response1_sp_75m$fit()

# Fit response with 50m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response1_sp_100m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"),
                    map=list(coeff_re=factor(c("tau.s(ExpShip).1","tau.s(ExpShip).2",rep(NA,N),
                                               "nu.s(ExpShip).1","nu.s(ExpShip).2",rep(NA,N),rep(NA,4))),coeff_fe=factor(rep(NA,5)),
                                                                                      log_lambda=factor(c("tau.s(ExpShip)",NA,
                                                                                                          "nu.s(ExpShip)",NA,NA))))

new_coeff_re=c(rep(0,2),baseline1_100m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],rep(0,2),
               baseline1_100m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline1_100m$coeff_re()[paste("omega.s(AngleNormal).",1:4,sep=""),1])
new_lambda=c(1,baseline1_100m$lambda()["tau.s(ID)",1],1,baseline1_100m$lambda()["nu.s(ID)",1],
             baseline1_100m$lambda()["omega.s(AngleNormal)",1])
response1_sp_100m$update_coeff_re(new_coeff_re)
response1_sp_100m$update_coeff_fe(baseline1_100m$coeff_fe()[,1])

response1_sp_100m$update_lambda(new_lambda)

response1_sp_100m$fit()
```

```{r get results response1}

estimates_res1_sp_50m=as.list(response1_sp_50m$tmb_rep(),what="Est")
std_res1_sp_50m=as.list(response1_sp_50m$tmb_rep(),what="Std")

estimates_res1_sp_75m=as.list(response1_sp_75m$tmb_rep(),what="Est")
std_res1_sp_75m=as.list(response1_sp_75m$tmb_rep(),what="Std")

estimates_res1_sp_100m=as.list(response1_sp_100m$tmb_rep(),what="Est")
std_res1_sp_100m=as.list(response1_sp_100m$tmb_rep(),what="Std")

#plot parameters
xmin=list("AngleNormal"=-pi,"ExpShip"=1/45)
xmax=list("AngleNormal"=pi,"ExpShip"=1/3)
link=list("ExpShip"=(\(x) 1/x))
xlabel=list("ExpShip"="Distance to ship")

plots_res1_sp_50m=response1_sp_50m$get_all_plots(baseline=baseline1_50m,xmin=xmin,xmax=xmax,link=link,xlabel=xlabel,
                            model_name="response1_sp_50m",show_CI="pointwise",save=TRUE)
plots_res1_sp_75m=response1_sp_75m$get_all_plots(baseline=baseline1_75m,xmin=xmin,xmax=xmax,link=link,xlabel=xlabel,
                            model_name="response1_sp_75m",show_CI="pointwise",save=TRUE)
plots_res1_sp_100m=response1_sp_100m$get_all_plots(baseline=baseline1_100m,xmin=xmin,xmax=xmax,link=link,xlabel=xlabel,
                            model_name="response1_sp_100m",show_CI="pointwise",save=TRUE)

```

```{r estimates on parameter scale response1}

post_coeff_res1_sp_50m=response1_sp_50m$post_coeff(n_post=1000)
post_par_res1_sp_50m=list("tau0"=exp(post_coeff_res1_sp_50m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res1_sp_50m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res1_sp_50m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res1_sp_50m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res1_sp_50m$log_lambda[,2])))
mean_par_res1_sp_50m=lapply(post_par_res1_sp_50m,mean)
sd_par_res1_sp_50m=lapply(post_par_res1_sp_50m,sd)

post_coeff_res1_sp_75m=response1_sp_75m$post_coeff(n_post=1000)
post_par_res1_sp_75m=list("tau0"=exp(post_coeff_res1_sp_75m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res1_sp_75m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res1_sp_75m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res1_sp_75m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res1_sp_75m$log_lambda[,2])))
mean_par_res1_sp_75m=lapply(post_par_res1_sp_75m,mean)
sd_par_res1_sp_75m=lapply(post_par_res1_sp_75m,sd)

post_coeff_res1_sp_100m=response1_sp_100m$post_coeff(n_post=1000)
post_par_res1_sp_100m=list("tau0"=exp(post_coeff_res1_sp_100m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res1_sp_100m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res1_sp_100m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res1_sp_100m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res1_sp_100m$log_lambda[,2])))
mean_par_res1_sp_100m=lapply(post_par_res1_sp_100m,mean)
sd_par_res1_sp_100m=lapply(post_par_res1_sp_100m,sd)

```

### Response model with splines of ExpShip based on baseline2

```{r fit response2}

par0 <- c(0,0,1,4,0)

#model formula
formulas <- list(mu1=~1,mu2=~1,tau =~s(ExpShip,k=3,bs="cs")+s(ID,bs="re"),
                 nu=~s(ExpShip,k=3,bs="cs")+s(ID,bs="re"),
                 omega=~ti(DistanceShore,k=5,bs="cs")+ti(AngleNormal,k=5,bs="cs")+ti(AngleNormal,DistanceShore,k=c(5,5),bs="cs"))

# Fit response with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response2_sp_50m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"),
                    map=list(coeff_re=factor(c(1:2,rep(NA,N),3:4,rep(NA,N),rep(NA,24))),coeff_fe=factor(rep(NA,5)),
                             log_lambda=factor(c("tau.s(ExpShip)",NA,"nu.s(ExpShip)",NA,rep(NA,4)))))

new_coeff_re=c(rep(0,2),baseline2_50m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline2_50m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2_50m$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2_50m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2_50m$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(1,baseline2_50m$lambda()["tau.s(ID)",1],1,baseline2_50m$lambda()["nu.s(ID)",1],baseline2_50m$lambda()[3:6,1])
response2_sp_50m$update_coeff_re(new_coeff_re)
response2_sp_50m$update_coeff_fe(baseline2_50m$coeff_fe()[,1])

response2_sp_50m$update_lambda(new_lambda)

response2_sp_50m$fit()

# Fit response with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response2_sp_75m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"),
                    map=list(coeff_re=factor(c(1:2,rep(NA,N),3:4,rep(NA,N),rep(NA,24))),coeff_fe=factor(rep(NA,5)),
                             log_lambda=factor(c("tau.s(ExpShip)",NA,"nu.s(ExpShip)",NA,rep(NA,4)))))

new_coeff_re=c(rep(0,2),baseline2_75m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline2_75m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2_75m$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2_75m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2_75m$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(1,baseline2_75m$lambda()["tau.s(ID)",1],1,baseline2_75m$lambda()["nu.s(ID)",1],baseline2_75m$lambda()[3:6,1])
response2_sp_75m$update_coeff_re(new_coeff_re)
response2_sp_75m$update_coeff_fe(baseline2_75m$coeff_fe()[,1])

response2_sp_75m$update_lambda(new_lambda)

response2_sp_75m$fit()

# Fit response with 100m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response2_sp_100m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"),
                    map=list(coeff_re=factor(c(1:2,rep(NA,N),3:4,rep(NA,N),rep(NA,24))),coeff_fe=factor(rep(NA,5)),
                             log_lambda=factor(c("tau.s(ExpShip)",NA,"nu.s(ExpShip)",NA,rep(NA,4)))))

new_coeff_re=c(rep(0,2),baseline2_100m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline2_100m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2_100m$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2_100m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2_100m$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(1,baseline2_100m$lambda()["tau.s(ID)",1],1,baseline2_100m$lambda()["nu.s(ID)",1],baseline2_100m$lambda()[3:6,1])
response2_sp_100m$update_coeff_re(new_coeff_re)
response2_sp_100m$update_coeff_fe(baseline2_100m$coeff_fe()[,1])

response2_sp_100m$update_lambda(new_lambda)

response2_sp_100m$fit()
```

```{r get results response2}
estimates_res2_50m=as.list(response2_sp_50m$tmb_rep(),what="Est")
std_res2_50m=as.list(response2_sp_50m$tmb_rep(),what="Std")

estimates_res2_75m=as.list(response2_sp_75m$tmb_rep(),what="Est")
std_res2_75m=as.list(response2_sp_75m$tmb_rep(),what="Std")

estimates_res2_100m=as.list(response2_sp_100m$tmb_rep(),what="Est")
std_res2_100m=as.list(response2_sp_100m$tmb_rep(),what="Std")

#plot parameters
xmin=list("AngleNormal"=-pi,"ExpShip"=1/45)
xmax=list("AngleNormal"=pi,"ExpShip"=1/3)
link=list("ExpShip"=(\(x) 1/x))
xlabel=list("ExpShip"="Distance to ship")

plot_res2_sp_50m=response2_sp_50m$get_all_plots(baseline=baseline2_50m,model_name="response2_sp_50m",xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise",save=TRUE)
plot_res2_sp_75m=response2_sp_75m$get_all_plots(baseline=baseline2_75m,model_name="response2_sp_75m",xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise",save=TRUE)
plot_res2_sp_100m=response2_sp_100m$get_all_plots(baseline=baseline2_100m,model_name="response2_sp_100m",xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise",save=TRUE)
```

```{r}

post_coeff_res2_sp_50m=response2_sp_50m$post_coeff(n_post=1000)
post_par_res2_sp_50m=list("tau0"=exp(post_coeff_res2_sp_50m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res2_sp_50m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res2_sp_50m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res2_sp_50m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res2_sp_50m$log_lambda[,2])))
mean_par_res2_sp_50m=lapply(post_par_res2_sp_50m,mean)
sd_par_res2_sp_50m=lapply(post_par_res2_sp_50m,sd)

post_coeff_res2_sp_75m=response2_sp_75m$post_coeff(n_post=1000)
post_par_res2_sp_75m=list("tau0"=exp(post_coeff_res2_sp_75m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res2_sp_75m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res2_sp_75m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res2_sp_75m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res2_sp_75m$log_lambda[,2])))
mean_par_res2_sp_75m=lapply(post_par_res2_sp_75m,mean)
sd_par_res2_sp_75m=lapply(post_par_res2_sp_75m,sd)

post_coeff_res2_sp_100m=response2_sp_100m$post_coeff(n_post=1000)
post_par_res2_sp_100m=list("tau0"=exp(post_coeff_res2_sp_100m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res2_sp_100m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res2_sp_100m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res2_sp_100m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res2_sp_100m$log_lambda[,2])))
mean_par_res2_sp_100m=lapply(post_par_res2_sp_100m,mean)
sd_par_res2_sp_100m=lapply(post_par_res2_sp_100m,sd)

```

### Response model with splines of ExpShip based on baseline3

```{r fit response3}

#initialize parameters
par0=c(0,0,1,4,0)

#define model
formulas <- list(mu1=~1,mu2=~1,tau =~s(ExpShip,k=3,bs="cs")+s(ID,bs="re"),
                 nu=~s(ExpShip,k=3,bs="cs")+s(ID,bs="re"),
                 omega=~ti(AngleNormal,k=5,bs="cs")+ti(ExpShore,k=5,bs="cs")+ti(AngleNormal,ExpShore,k=c(5,5),bs="cs"))

#Fit response with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_sp_50m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H),
                    map=list(coeff_re=factor(c("tau.s(ExpShip).1","tau.s(ExpShip).2",rep(NA,N),"nu.s(ExpShip).1","nu.s(ExpShip).2",rep(NA,N),rep(NA,24))),coeff_fe=factor(rep(NA,5)),log_lambda=factor(c("tau.s(ExpShip)",NA,"nu.s(ExpShip)",NA,rep(NA,4)))))

new_coeff_re=c(rep(0,2),baseline3_50m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline3_50m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline3_50m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline3_50m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline3_50m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(1,baseline3_50m$lambda()["tau.s(ID)",1],1,baseline3_50m$lambda()["nu.s(ID)",1],baseline3_50m$lambda()[3:6,1])
response3_sp_50m$update_coeff_re(new_coeff_re)
response3_sp_50m$update_coeff_fe(baseline3_50m$coeff_fe()[,1])

response3_sp_50m$update_lambda(new_lambda)
response3_sp_50m$fit()

#Fit response with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_sp_75m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H),
                    map=list(coeff_re=factor(c("tau.s(ExpShip).1","tau.s(ExpShip).2",rep(NA,N),"nu.s(ExpShip).1","nu.s(ExpShip).2",rep(NA,N),rep(NA,24))),coeff_fe=factor(rep(NA,5)),log_lambda=factor(c("tau.s(ExpShip)",NA,"nu.s(ExpShip)",NA,rep(NA,4)))))

new_coeff_re=c(rep(0,2),baseline3_75m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline3_75m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline3_75m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline3_75m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline3_75m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(1,baseline3_75m$lambda()["tau.s(ID)",1],1,baseline3_75m$lambda()["nu.s(ID)",1],baseline3_75m$lambda()[3:6,1])
response3_sp_75m$update_coeff_re(new_coeff_re)
response3_sp_75m$update_coeff_fe(baseline3_75m$coeff_fe()[,1])

response3_sp_75m$update_lambda(new_lambda)
response3_sp_75m$fit()

#Fit response with 100m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_sp_100m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H),
                    map=list(coeff_re=factor(c("tau.s(ExpShip).1","tau.s(ExpShip).2",rep(NA,N),"nu.s(ExpShip).1","nu.s(ExpShip).2",rep(NA,N),rep(NA,24))),coeff_fe=factor(rep(NA,5)),log_lambda=factor(c("tau.s(ExpShip)",NA,"nu.s(ExpShip)",NA,rep(NA,4)))))

new_coeff_re=c(rep(0,2),baseline3_100m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline3_100m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline3_100m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline3_100m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline3_100m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(1,baseline3_100m$lambda()["tau.s(ID)",1],1,baseline3_100m$lambda()["nu.s(ID)",1],baseline3_100m$lambda()[3:6,1])
response3_sp_100m$update_coeff_re(new_coeff_re)
response3_sp_100m$update_coeff_fe(baseline3_100m$coeff_fe()[,1])

response3_sp_100m$update_lambda(new_lambda)
response3_sp_100m$fit()
```

```{r get results response3}
estimates_res3_sp_50m=as.list(response3_sp_50m$tmb_rep(),what="Est")
std_res3_sp_50m=as.list(response3_sp_50m$tmb_rep(),what="Std")

estimates_res3_sp_75m=as.list(response3_sp_75m$tmb_rep(),what="Est")
std_res3_sp_75m=as.list(response3_sp_75m$tmb_rep(),what="Std")

estimates_res3_sp_100m=as.list(response3_sp_100m$tmb_rep(),what="Est")
std_res3_sp_100m=as.list(response3_sp_100m$tmb_rep(),what="Std")

#plot parameters
xmin=list("ExpShore"=1/D_up,"AngleNormal"=-pi,"ExpShip"=1/45)
xmax=list("ExpShore"=1/D_low,"AngleNormal"=pi,"ExpShip"=1/3)
link=list("ExpShore"=(\(x) 1/x),"ExpShip"=(\(x) 1/x))
xlabel=list("ExpShore"="Distance to shore","ExpShip"="Distance to ship")

plot_res3_sp_50m=response3_sp_50m$get_all_plots(baseline=baseline3_50m,model_name="response3_sp_50m",xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise",save=TRUE)
plot_res3_sp_75m=response3_sp_75m$get_all_plots(baseline=baseline3_75m,model_name="response3_sp_75m",xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise",save=TRUE)
plot_res3_sp_100m=response3_sp_100m$get_all_plots(baseline=baseline3_100m,model_name="response3_sp_100m",xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise",save=TRUE)

```

```{r estimates on parameter scale response3}

post_coeff_res3_sp_50m=response3_sp_50m$post_coeff(n_post=1000)
post_par_res3_sp_50m=list("tau0"=exp(post_coeff_res3_sp_50m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res3_sp_50m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res3_sp_50m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res3_sp_50m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res3_sp_50m$log_lambda[,2])))
mean_par_res3_sp_50m=lapply(post_par_res3_sp_50m,mean)
sd_par_res3_sp_50m=lapply(post_par_res3_sp_50m,sd)

post_coeff_res3_sp_75m=response3_sp_75m$post_coeff(n_post=1000)
post_par_res3_sp_75m=list("tau0"=exp(post_coeff_res3_sp_75m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res3_sp_75m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res3_sp_75m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res3_sp_75m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res3_sp_75m$log_lambda[,2])))
mean_par_res3_sp_75m=lapply(post_par_res3_sp_75m,mean)
sd_par_res3_sp_75m=lapply(post_par_res3_sp_75m,sd)

post_coeff_res3_sp_100m=response3_sp_100m$post_coeff(n_post=1000)
post_par_res3_sp_100m=list("tau0"=exp(post_coeff_res3_sp_100m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res3_sp_100m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res3_sp_100m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res3_sp_100m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res3_sp_100m$log_lambda[,2])))
mean_par_res3_sp_100m=lapply(post_par_res3_sp_100m,mean)
sd_par_res3_sp_100m=lapply(post_par_res3_sp_100m,sd)

```

### Response model with exponential of ExpShip based on baseline2

```{r}

par0=c(0,0,1,4,0)

#define model
formulas <- list(mu1=~1,mu2=~1,tau =~ExpShip+s(ID,bs="re"),
                 nu=~ExpShip+s(ID,bs="re"),
                 omega=~ti(DistanceShore,k=5,bs="cs")+ti(AngleNormal,k=5,bs="cs")+ti(AngleNormal,DistanceShore,k=c(5,5),bs="cs"))

# Fit response with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response2_exp_50m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H),
                    map=list(coeff_re=factor(c(rep(NA,12),rep(NA,24))),coeff_fe=factor(c(NA,NA,NA,1,NA,2,NA)),
                             log_lambda=factor(c(NA,NA,NA,NA,NA,NA))))

new_coeff_fe=c(baseline2_50m$coeff_fe()[1:3,1],0,baseline2_50m$coeff_fe()[4,1],0,baseline2_50m$coeff_fe()[5,1])
new_coeff_re=c(baseline2_50m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],baseline2_50m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2_50m$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2_50m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2_50m$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(baseline2_50m$lambda()["tau.s(ID)",1],baseline2_50m$lambda()["nu.s(ID)",1],baseline2_50m$lambda()[3:6,1])
response2_exp_50m$update_coeff_re(new_coeff_re)
response2_exp_50m$update_coeff_fe(new_coeff_fe)

response2_exp_50m$update_lambda(new_lambda)

response2_exp_50m$fit()

# Fit response with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response2_exp_75m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H),
                    map=list(coeff_re=factor(c(rep(NA,12),rep(NA,24))),coeff_fe=factor(c(NA,NA,NA,1,NA,2,NA)),
                             log_lambda=factor(c(NA,NA,NA,NA,NA,NA))))

new_coeff_fe=c(baseline2_75m$coeff_fe()[1:3,1],0,baseline2_75m$coeff_fe()[4,1],0,baseline2_75m$coeff_fe()[5,1])
new_coeff_re=c(baseline2_75m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],baseline2_75m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2_75m$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2_75m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2_75m$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(baseline2_75m$lambda()["tau.s(ID)",1],baseline2_75m$lambda()["nu.s(ID)",1],baseline2_75m$lambda()[3:6,1])
response2_exp_75m$update_coeff_re(new_coeff_re)
response2_exp_75m$update_coeff_fe(new_coeff_fe)

response2_exp_75m$update_lambda(new_lambda)

response2_exp_75m$fit()

# Fit response with 100m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response2_exp_100m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H),
                    map=list(coeff_re=factor(c(rep(NA,12),rep(NA,24))),coeff_fe=factor(c(NA,NA,NA,1,NA,2,NA)),
                             log_lambda=factor(c(NA,NA,NA,NA,NA,NA))))

new_coeff_fe=c(baseline2_100m$coeff_fe()[1:3,1],0,baseline2_100m$coeff_fe()[4,1],0,baseline2_100m$coeff_fe()[5,1])
new_coeff_re=c(baseline2_100m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],baseline2_100m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2_100m$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2_100m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2_100m$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(baseline2_100m$lambda()["tau.s(ID)",1],baseline2_100m$lambda()["nu.s(ID)",1],baseline2_100m$lambda()[3:6,1])
response2_exp_100m$update_coeff_re(new_coeff_re)
response2_exp_100m$update_coeff_fe(new_coeff_fe)

response2_exp_100m$update_lambda(new_lambda)

response2_exp_100m$fit()
```

```{r}
estimates_res2_exp_50m=as.list(response2_exp_50m$tmb_rep(),what="Est")
std_res2_exp_50m=as.list(response2_exp_50m$tmb_rep(),what="Std")

estimates_res2_exp_75m=as.list(response2_exp_75m$tmb_rep(),what="Est")
std_res2_exp_75m=as.list(response2_exp_75m$tmb_rep(),what="Std")

estimates_res2_exp_100m=as.list(response2_exp_100m$tmb_rep(),what="Est")
std_res2_exp_100m=as.list(response2_exp_100m$tmb_rep(),what="Std")

#plot parameters
xmin=list("ExpShore"=1/D_up,"AngleNormal"=-pi,"ExpShip"=1/60)
xmax=list("ExpShore"=1/D_low,"AngleNormal"=pi,"ExpShip"=1/5)
link=list("ExpShore"=(\(x) 1/x),"ExpShip"=(\(x) 1/x))
xlabel=list("ExpShore"="Distance to shore","ExpShip"="Distance to ship")

plots_res2_exp=response2_exp_50m$get_all_plots(baseline=baseline2_50m,model_name="response2_exp_50m",xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise",save=TRUE)

plots_res2_exp_75m=response2_exp_75m$get_all_plots(baseline=baseline2_75m,model_name="response2_exp_75m",xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise",save=TRUE)

plots_res2_exp_100m=response2_exp_100m$get_all_plots(baseline=baseline2_100m,model_name="response2_exp_100m",xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise",save=TRUE)

```

### Response model with exponential of ExpShip based on baseline3

```{r}

par0=c(0,0,1,4,0)

#define model
formulas <- list(mu1=~1,mu2=~1,tau =~ExpShip+s(ID,bs="re"),
                 nu=~ExpShip+s(ID,bs="re"),
                 omega=~ti(AngleNormal,k=5,bs="cs")+ti(ExpShore,k=5,bs="cs")+ti(AngleNormal,ExpShore,k=c(5,5),bs="cs"))

# Fit response with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_exp_50m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H),
                    map=list(coeff_re=factor(c(rep(NA,12),rep(NA,24))),coeff_fe=factor(c(NA,NA,NA,1,NA,2,NA)),
                             log_lambda=factor(c(NA,NA,NA,NA,NA,NA))))

new_coeff_fe=c(baseline3_50m$coeff_fe()[1:3,1],0,baseline3_50m$coeff_fe()[4,1],0,baseline3_50m$coeff_fe()[5,1])
new_coeff_re=c(baseline3_50m$coeff_re()[paste("tau.s(ID).",1:6,sep=""),1],baseline3_50m$coeff_re()[paste("nu.s(ID).",1:6,sep=""),1],
               baseline3_50m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline3_50m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline3_50m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(baseline3_50m$lambda()["tau.s(ID)",1],baseline3_50m$lambda()["nu.s(ID)",1],baseline3_50m$lambda()[3:6,1])
response3_exp_50m$update_coeff_re(new_coeff_re)
response3_exp_50m$update_coeff_fe(new_coeff_fe)

response3_exp_50m$update_lambda(new_lambda)
response3_exp_50m$fit()

# Fit response with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_exp_75m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H),
                    map=list(coeff_re=factor(c(rep(NA,12),rep(NA,24))),coeff_fe=factor(c(NA,NA,NA,1,NA,2,NA)),
                             log_lambda=factor(c(NA,NA,NA,NA,NA,NA))))

new_coeff_fe=c(baseline3_75m$coeff_fe()[1:3,1],0,baseline3_75m$coeff_fe()[4,1],0,baseline3_75m$coeff_fe()[5,1])
new_coeff_re=c(baseline3_75m$coeff_re()[paste("tau.s(ID).",1:6,sep=""),1],baseline3_75m$coeff_re()[paste("nu.s(ID).",1:6,sep=""),1],
               baseline3_75m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline3_75m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline3_75m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(baseline3_75m$lambda()["tau.s(ID)",1],baseline3_75m$lambda()["nu.s(ID)",1],baseline3_75m$lambda()[3:6,1])
response3_exp_75m$update_coeff_re(new_coeff_re)
response3_exp_75m$update_coeff_fe(new_coeff_fe)

response3_exp_75m$update_lambda(new_lambda)
response3_exp_75m$fit()

# Fit response with 100m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_exp_100m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H),
                    map=list(coeff_re=factor(c(rep(NA,12),rep(NA,24))),coeff_fe=factor(c(NA,NA,NA,1,NA,2,NA)),
                             log_lambda=factor(c(NA,NA,NA,NA,NA,NA))))

new_coeff_fe=c(baseline3_100m$coeff_fe()[1:3,1],0,baseline3_100m$coeff_fe()[4,1],0,baseline3_100m$coeff_fe()[5,1])
new_coeff_re=c(baseline3_100m$coeff_re()[paste("tau.s(ID).",1:6,sep=""),1],baseline3_100m$coeff_re()[paste("nu.s(ID).",1:6,sep=""),1],
               baseline3_100m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline3_100m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline3_100m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(baseline3_100m$lambda()["tau.s(ID)",1],baseline3_100m$lambda()["nu.s(ID)",1],baseline3_100m$lambda()[3:6,1])
response3_exp_100m$update_coeff_re(new_coeff_re)
response3_exp_100m$update_coeff_fe(new_coeff_fe)

response3_exp_100m$update_lambda(new_lambda)
response3_exp_100m$fit()
```

```{r}
estimates_res3_exp_50m=as.list(response3_exp_50m$tmb_rep(),what="Est")
std_res3_exp_50m=as.list(response3_exp_50m$tmb_rep(),what="Std")

estimates_res3_exp_75m=as.list(response3_exp_75m$tmb_rep(),what="Est")
std_res3_exp_75m=as.list(response3_exp_75m$tmb_rep(),what="Std")

estimates_res3_exp_100m=as.list(response3_exp_100m$tmb_rep(),what="Est")
std_res3_exp_100m=as.list(response3_exp_100m$tmb_rep(),what="Std")

#plot parameters
xmin=list("ExpShore"=1/D_up,"AngleNormal"=-pi,"ExpShip"=1/60)
xmax=list("ExpShore"=1/D_low,"AngleNormal"=pi,"ExpShip"=1/5)
link=list("ExpShore"=(\(x) 1/x),"ExpShip"=(\(x) 1/x))
xlabel=list("ExpShore"="Distance to shore","ExpShip"="Distance to ship")

plots_res3_exp_50m=response3_exp_50m$get_all_plots(baseline=baseline3_50m,model_name="response3_exp_50m",xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise",save=TRUE)
plots_res3_exp_75m=response3_exp_75m$get_all_plots(baseline=baseline3_75m,model_name="response3_exp_75m",xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise",save=TRUE)
plots_res3_exp_100m=response3_exp_100m$get_all_plots(baseline=baseline3_100m,model_name="response3_exp_100m",xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise",save=TRUE)

```