### 2025-01-07 R/application/RCVM/fit_rcvm.Rmd

---
title: "Final analysis of Greenland narwhals trajectories with smoothSDE"
author: "Alexandre Delporte"
date: "2024-06-07"
output: html_document
---

In this analysis, we introduce covariates \texttt{AngleNormal} and \texttt{Ishore} ($\Theta$ and $I_{shore}$) in the SDE models for the narwhal data. This allows to better understand how the fjords shoreline influences the narwhals movement in the baseline model.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up

```{r packages,message=FALSE}
set.seed(42)

library(smoothSDE)
library(ggplot2)
library(dplyr)
library(plotly)
library(htmlwidgets)
library(mgcv)
library(here)
library(xtable)
library(sf)
library(gganimate) # animated trajectories
```

```{r get data,message=FALSE}

# Set the path to the directory containing the data
par_dir=here() #parent directory 
narwhal_data_path <- file.path(par_dir,"Data","preprocessed_data","narwhals")

#data before exposure with first 24h removed
dataBE24=read.csv(file.path(narwhal_data_path,"DataBE1.csv"), header = TRUE,dec = ".")

#data before exposure with first 12h removed
dataBE12=read.csv(file.path(narwhal_data_path,"DataBE2.csv"), header = TRUE,dec = ".")

#data after exposure
dataAE=read.csv(file.path(narwhal_data_path,"DataAE.csv"), header = TRUE,dec = ".")

```

We choose thresholds on the distance to the shore to define the Ishore covariate.

```{r preprocess data}

D_low=0.073
D_up=3
dataBE24[dataBE24$DistanceShore> D_up,"Ishore"]=0
dataBE24[dataBE24$DistanceShore< D_low,"Ishore"]=1/D_low

dataBE12[dataBE12$DistanceShore> D_up,"Ishore"]=0
dataBE12[dataBE12$DistanceShore< D_low,"Ishore"]=1/D_low

dataAE[dataAE$DistanceShore> D_up,"Ishore"]=0
dataAE[dataAE$DistanceShore< D_low,"Ishore"]=1/D_low

N=length(unique(dataBE12$ID))

n_pre12=length(dataBE12$time) #number of observations before exposure
n_pre24=length(dataBE24$time) #number of observations before exposure
n_post=length(dataAE$time) #number of observations after exposure
```

# Fit baseline models

We try sde models where the parameters are smooth function of the covariates AngleNormal and Ishore.
Here, for simplicity we will always fix the measurement error to $30$ m. This choice is based on the paper of Weensven.
We will assess the sensitivity of our results to the hyperparameter later.

## Baseline CTCRW with AngleNormal covariate in $\tau$

The first model is a CTCRW where the covariate AngleNormal influences the persistence of the movement.

```{r fit baseline_ctcrw0,message=FALSE}

#initial parameters
par0 <- c(0,0,1,4)

#model formula
formulas <- list(mu1=~1,mu2=~1,tau =~s(AngleNormal,k=10,bs="cc")+s(ID,bs="re"),nu=~s(ID,bs="re"))

#Fixed measurement error
sigma_obs=0.03
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_pre12))

#Fit baseline with measurement error estimated from the data
baseline_ctcrw0<- SDE$new(formulas = formulas,data = dataBE12,type = "CTCRW",
                          response = c("x","y"),
                          par0 = par0,
                          other_data=list("H"=H),
                          fixpar=c("mu1","mu2"))
baseline_ctcrw0$fit()

```

```{r results baseline_ctcrw0}

estimates_bas_ctcrw0=as.list(baseline_ctcrw0$tmb_rep(),what="Est")
std_bas_ctcrw0=as.list(baseline_ctcrw0$tmb_rep(),what="Std")

```

```{r render estimates table baseline_ctcrw0}

# Function to extract estimate values into a data frame format
combine_to_df <- function(estimates, std_dev, param_name) {
  data.frame(
    Parameter = param_name,
    Estimate = unlist(estimates, use.names = FALSE),
    Std_Error = unlist(std_dev, use.names = FALSE)
  )
}

# Create data frames for each set
log_sigma_obs <- combine_to_df(estimates_bas_ctcrw0$log_sigma_obs, 
                               std_bas_ctcrw0$log_sigma_obs, "log_sigma_obs")
coeff_fe <- combine_to_df(estimates_bas_ctcrw0$coeff_fe,
                          std_bas_ctcrw0$coeff_fe, 
                          rownames(estimates_bas_ctcrw0$coeff_fe))

coeff_re <- combine_to_df(estimates_bas_ctcrw0$coeff_re,
                          std_bas_ctcrw0$coeff_re, 
                          rownames(estimates_bas_ctcrw0$coeff_re))

log_lambda <- combine_to_df(estimates_bas_ctcrw0$log_lambda,
                          std_bas_ctcrw0$log_lambda, 
                          rownames(estimates_bas_ctcrw0$log_lambda))

# Combine all data frames into one
table_data <- bind_rows(log_sigma_obs, coeff_fe,coeff_re,log_lambda)

```

```{r, results='asis',message=FALSE}
latex_table <- xtable(table_data, 
                      caption = "Parameter Estimates and Standard Errors")

# Print LaTeX table
print(latex_table, type = "latex", include.rownames = FALSE)
```
We can get confidence intervals for the intercept of each parameter and the random effects standard deviations.

```{r CI baseline_ctcrw0 parameters}

post_coeff_bas_ctrcw0=baseline_ctcrw0$post_coeff(n_post=10000)
post_par_bas_ctcrw0=list(
  "tau"=exp(post_coeff_bas_ctrcw0$coeff_fe[,"tau.(Intercept)"]),
  "nu"=exp(post_coeff_bas_ctrcw0$coeff_fe[,"nu.(Intercept)"]),
  "sigma_tau"=1/sqrt(exp(post_coeff_bas_ctrcw0$log_lambda[,"tau.s(ID)"])),
  "sigma_nu"=1/sqrt(exp(post_coeff_bas_ctrcw0$log_lambda[,"nu.s(ID)"])))

mean_par_bas_ctcrw0=lapply(post_par_bas_ctcrw0,mean)
sd_par_bas_ctcrw0=lapply(post_par_bas_ctcrw0,sd)
quant_par_bas_ctcrw0 <- lapply(post_par_bas_ctcrw0, 
                               quantile, probs = c(0.025, 0.975))

```

```{r CI table baseline_ctcrw0 parameters, message=FALSE}

parameter_names <- names(mean_par_bas_ctcrw0)
estimates <- mean_par_bas_ctcrw0
conf_intervals <- sapply(parameter_names, function(param) {
  sprintf("$[%.2f; %.2f]$", quant_par_bas_ctcrw0[[param]][1], 
          quant_par_bas_ctcrw0[[param]][2])
})

table_data <- data.frame(
  Parameter = 
    c("$\\tau$", "$\\nu$","$\\sigma_{\\tau}$","$\\sigma_{\\nu}$"),
  Estimate = sprintf("$%.2f$", mean_par_bas_ctcrw0),
  CI = conf_intervals,
  stringsAsFactors = FALSE
)

```

```{r,results='asis',echo=FALSE}
xtab <- xtable(
  table_data,
  caption = "CTCRW baseline estimations with random effects",
  label = "table:ctcrw_baseline_estimations_with_random_effects"
)

print(
  xtab,type="latex",
  include.rownames = FALSE,
  sanitize.text.function = identity,  # Prevent escaping of LaTeX math symbols
  hline.after = c(-1, 0, nrow(table_data)),  
  table.placement="H"
)
```

We can now plot the smooth parameter $\tau$ as a function of the covariate \texttt{AngleNormal}.

```{r plot smooth parameters baseline_ctcrw0 }
baseline_ctcrw0$plot_par(var="AngleNormal",
                         par_names=c("tau","nu"),n_post=100)

```

```{r plot smooth parameters CI baseline_ctcrw0}
baseline_ctcrw0$get_all_plots(baseline=NULL,show_CI="simultaneous")
```

The movement is clearly more persistent along the shoreline, meaning that $\tau$ increases when \texttt{AngleNormal} is close to $\pm \frac{\pi}{2}$).

## Baseline racvm with AngleNormal covariate in tau and omega

We now decide to introduce a rotational parameter $\omega$ in the SDE to allow the velocity to rotate depending on the narwhals position with respect to the shoreline.
We start by introducing the covariate \texttt{AngleNormal} in the new smooth parameter $\omega$ to understand how the animals may rotate depending on the direction of their movement with respect to the shoreline.

```{r fit baseline_racvm0,cache=TRUE}

#initial parameters
par0 <- c(0,0,1,4,0)

#model formula
formulas <- list(mu1=~1,mu2=~1,
                 tau =~s(AngleNormal,k=10,bs="cc")+s(ID,bs="re"),
                 nu=~s(ID,bs="re"),
                 omega=~s(AngleNormal,k=10,bs="cs"))

#Fixed measurement error
sigma_obs=0.03
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_pre12))

# Fit baseline with measurement error estimated from the data
baseline_racvm0<- SDE$new(formulas = formulas,data = dataBE12,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"))

baseline_racvm0$fit()

```

```{r results baseline_racvm0}

estimates_bas_racvm0=as.list(baseline_racvm0$tmb_rep(),what="Est")
std_bas_racvm0=as.list(baseline_racvm0$tmb_rep(),what="Std")
```

```{r render estimates table baseline_racvm0 }

# Function to extract estimate values into a data frame format
combine_to_df <- function(estimates, std_dev, param_name) {
  data.frame(
    Parameter = param_name,
    Estimate = unlist(estimates, use.names = FALSE),
    Std_Error = unlist(std_dev, use.names = FALSE)
  )
}

# Create data frames for each set
log_sigma_obs <- combine_to_df(estimates_bas_racvm0$log_sigma_obs, 
                               std_bas_racvm0$log_sigma_obs, "log_sigma_obs")
coeff_fe <- combine_to_df(estimates_bas_racvm0$coeff_fe, 
                          std_bas_racvm0$coeff_fe, 
                          rownames(estimates_bas_racvm0$coeff_fe))
coeff_re <- combine_to_df(estimates_bas_racvm0$coeff_re, 
                          std_bas_racvm0$coeff_re, 
                          rownames(estimates_bas_racvm0$coeff_re))
log_lambda <- combine_to_df(estimates_bas_racvm0$log_lambda, 
                          std_bas_racvm0$log_lambda, 
                          rownames(estimates_bas_racvm0$log_lambda))

# Combine all data frames into one
table_data <- bind_rows(log_sigma_obs, coeff_fe,coeff_re,log_lambda)

```

```{r, results='asis',message=FALSE}
latex_table <- xtable(table_data, 
                      caption = "Parameter Estimates and Standard Errors")

# Print LaTeX table
print(latex_table, type = "latex", include.rownames = FALSE)
```

```{r CI baseline_racvm0 parameters}

post_coeff_bas_racvm0=baseline_racvm0$post_coeff(n_post=10000)
post_par_bas_racvm0=list(
  "tau"=exp(post_coeff_bas_racvm0$coeff_fe[,"tau.(Intercept)"]),
  "nu"=exp(post_coeff_bas_racvm0$coeff_fe[,"nu.(Intercept)"]),
  "omega"=post_coeff_bas_racvm0$coeff_fe[,"omega.(Intercept)"],
  "sigma_tau"=1/sqrt(exp(post_coeff_bas_racvm0$log_lambda[,"tau.s(ID)"])),
  "sigma_nu"=1/sqrt(exp(post_coeff_bas_racvm0$log_lambda[,"nu.s(ID)"])))

mean_par_bas_racvm0=lapply(post_par_bas_racvm0,mean)
sd_par_bas_racvm0=lapply(post_par_bas_racvm0,sd)
quant_par_bas_racvm0 <- lapply(post_par_bas_racvm0, 
                               quantile, probs = c(0.025, 0.975))

```

```{r CI table baseline_racvm0 parameters, message=FALSE}

parameter_names <- names(mean_par_bas_racvm0)
estimates <- mean_par_bas_racvm0
conf_intervals <- sapply(parameter_names, function(param) {
  sprintf("$[%.2f; %.2f]$", quant_par_bas_racvm0[[param]][1], 
          quant_par_bas_racvm0[[param]][2])
})

table_data <- data.frame(
  Parameter = 
    c("$\\tau$", "$\\nu$","$\\omega$","$\\sigma_{\\tau}$","$\\sigma_{\\nu}$"),
  Estimate = sprintf("$%.2f$", mean_par_bas_racvm0),
  CI = conf_intervals,
  stringsAsFactors = FALSE
)

```

```{r,results='asis',echo=FALSE}
xtab <- xtable(
  table_data,
  caption = "racvm0 baseline estimations",
  label = "table:racvm0_baseline_estimations_with_random_effects"
)

print(
  xtab,type="latex",
  include.rownames = FALSE,
  sanitize.text.function = identity,  # Prevent escaping of LaTeX math symbols
  hline.after = c(-1, 0, nrow(table_data)),  
  table.placement="H"
)
```

```{r plot smooth parameters baseline_racvm0}
baseline_racvm0$plot_par(var="AngleNormal",
                         par_names=c("omega","tau"),n_post=100)

```

$\omega$ close to $\pm \pi$ indicates movement toward the shoreline. Here, it seems that the narwhals have a slight tendency to rotate, presumably to move along the shoreline or change direction, when they are heading toward the shore. However, it is not very clear since the confidence intervals are wide.

```{r plot smooth parameter CI baseline_racvm0}
baseline_racvm0$get_all_plots(baseline=NULL,show_CI="pointwise")
```

## Baseline with AngleNormal and Ishore in omega and tau

We add the effect of the distance to the shore through the covariate \texttt{Ishore}.
This model has a lot of parameters and takes around 15 minutes to fit. 
Indeed, we will need to fix the penalties in the tensor splines in advance to ensure convergence of the optimizer. The standard deviation of the random effects and the smoothing penalties for the splines in $\tau$ will still be estimated from the data and not fixed. To illustrate this we first fit a model where the smoothing penalties are all estimated from the data.

```{r fit baseline_racvm1_unstable,cache=TRUE,message=FALSE}

#initial parameters
par0 <- c(0,0,1,4,0)

#Fixed measurement error
sigma_obs=0.03
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_pre12))

#parameters formulas
formulas <- list(mu1 = ~1 ,mu2 =~1,tau =~s(AngleNormal,k=10,bs="cc")+s(ID,bs="re"),nu=~s(ID,bs="re"),
                   omega=~ti(AngleNormal,k=5,bs="cs")+ti(Ishore,k=5,bs="cs")+
                   ti(AngleNormal,Ishore,k=c(5,5),bs="cs"))

# Fit baseline
baseline_racvm1_unstable<- SDE$new(formulas = formulas,data = dataBE12,type = "RACVM",
                    response = c("x","y"),par0 = par0,other_data=list("H"=H),
                    fixpar=c("mu1","mu2"))

baseline_racvm1_unstable$fit()

```

```{r get results baseline_racvm1_unstable,message=FALSE}

estimates_bas_racvm1_unstable=as.list(baseline_racvm1_unstable$tmb_rep(),what="Est")
std_bas_racvm1_unstable=as.list(baseline_racvm1_unstable$tmb_rep(),what="Std")

estimates_bas_racvm1_unstable

```
```{r std random effects baseline_racvm1_unstable}

baseline_racvm1_unstable$sdev()

```

We see that the estimate for the standard deviation $\sigma_\tau$ of the random effect for $\tau$ is very high while the estimate for the standard deviation $\sigma_\nu$ of the random effect for $\nu$ is very close to $0$. The smoothing penalties estimates are extremely high (the smoothing penalties are the exponential of \texttt{estimates\_bas\_racvm1\_unstable$log\_lambda}). This suggests that the optimization did not go well and numerical errors occurred. For this reason, we decide to fix the smoothing penalties to "reasonable" values in the tensor splines.

```{r fit baseline_racvm1,cache=TRUE}

#initial parameters
par0 <- c(0,0,1,4,0)
init_lambda=exp(c(3,3,4,-2,-2,-2,-2))

#mapping for fixed smoothing penalties
new_map=list("log_lambda"=factor(c(1,2,3,rep(NA,4))))

#Fixed measurement error
sigma_obs=0.03
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_pre12))

#parameters formulas
formulas <- list(mu1 = ~1 ,mu2 =~1,tau =~s(AngleNormal,k=10,bs="cc")+s(ID,bs="re"),nu=~s(ID,bs="re"),
                   omega=~ti(AngleNormal,k=5,bs="cs")+ti(Ishore,k=5,bs="cs")+
                   ti(AngleNormal,Ishore,k=c(5,5),bs="cs"))

# Fit baseline
baseline_racvm1<- SDE$new(formulas = formulas,data = dataBE12,type = "RACVM",
                    response = c("x","y"),par0 = par0,other_data=list("H"=H),
                    fixpar=c("mu1","mu2"))

baseline_racvm1$update_lambda(init_lambda)
baseline_racvm1$update_map(new_map)
baseline_racvm1$fit()

```

```{r render estimates table baseline_racvm1 }

# Create data frames for each set
log_sigma_obs <- combine_to_df(estimates_bas_racvm1$log_sigma_obs, 
                               std_bas_racvm1$log_sigma_obs, "log_sigma_obs")
coeff_fe <- combine_to_df(estimates_bas_racvm1$coeff_fe, 
                          std_bas_racvm1$coeff_fe, 
                          rownames(estimates_bas_racvm1$coeff_fe))
coeff_re <- combine_to_df(estimates_bas_racvm1$coeff_re, 
                          std_bas_racvm1$coeff_re, 
                          rownames(estimates_bas_racvm1$coeff_re))
log_lambda <- combine_to_df(estimates_bas_racvm1$log_lambda, 
                          std_bas_racvm1$log_lambda, 
                          rownames(estimates_bas_racvm1$log_lambda))

# Combine all data frames into one
table_data <- bind_rows(log_sigma_obs, coeff_fe,coeff_re,log_lambda)

```

```{r, results='asis',message=FALSE}
latex_table <- xtable(table_data, 
                      caption = "Parameter Estimates and Standard Errors")

# Print LaTeX table
print(latex_table, type = "latex", include.rownames = FALSE)
```

```{r CI baseline_racvm1 parameters}

post_coeff_bas_racvm1=baseline_racvm1$post_coeff(n_post=10000)
post_par_bas_racvm1=list(
  "tau"=exp(post_coeff_bas_racvm1$coeff_fe[,"tau.(Intercept)"]),
  "nu"=exp(post_coeff_bas_racvm1$coeff_fe[,"nu.(Intercept)"]),
  "omega"=post_coeff_bas_racvm1$coeff_fe[,"omega.(Intercept)"],
  "sigma_tau"=1/sqrt(exp(post_coeff_bas_racvm1$log_lambda[,"tau.s(ID)"])),
  "sigma_nu"=1/sqrt(exp(post_coeff_bas_racvm1$log_lambda[,"nu.s(ID)"])))

mean_par_bas_racvm1=lapply(post_par_bas_racvm1,mean)
sd_par_bas_racvm1=lapply(post_par_bas_racvm1,sd)
quant_par_bas_racvm1 <- lapply(post_par_bas_racvm1, 
                               quantile, probs = c(0.025, 0.975))

```

```{r CI table baseline_racvm1 parameters, message=FALSE}

parameter_names <- names(mean_par_bas_racvm1)
estimates <- mean_par_bas_racvm1
conf_intervals <- sapply(parameter_names, function(param) {
  sprintf("$[%.2f; %.2f]$", quant_par_bas_racvm1[[param]][1], 
          quant_par_bas_racvm1[[param]][2])
})

table_data <- data.frame(
  Parameter = 
    c("$\\tau$", "$\\nu$","$\\omega$","$\\sigma_{\\tau}$","$\\sigma_{\\nu}$"),
  Estimate = sprintf("$%.2f$", mean_par_bas_racvm1),
  CI = conf_intervals,
  stringsAsFactors = FALSE
)

```

```{r,results='asis',echo=FALSE}
xtab <- xtable(
  table_data,
  caption = "racvm1 baseline estimations",
  label = "table:racvm1_baseline_estimations_with_random_effects"
)

print(
  xtab,type="latex",
  include.rownames = FALSE,
  sanitize.text.function = identity,  # Prevent escaping of LaTeX math symbols
  hline.after = c(-1, 0, nrow(table_data)),  
  table.placement="H"
)
```
```{r display plots baseline_racvm1}

Ishore_values=c(1/0.25,1/2,1/10)
plots=lapply(1:length(Ishore_values),function(i){
  baseline_racvm1$plot_par(var="AngleNormal",covs=data.frame("ID"="Asgeir","Ishore"=Ishore_values[i]),
                         par_names=c("omega"),n_post=100)})
plots

```

```{r plot smooth parameters CI baseline_racvm1}

#plot parameters

xmin=list("Ishore"=1/D_up)
xmax=list("Ishore"=1/D_low)
link=list("Ishore"=(\(x) 1/x))
xlabel=list("Ishore"="Distance to shore")

plots_bas_racvm1=baseline_racvm1$get_all_plots(baseline=NULL,
      xmin=xmin, xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")

plots_bas_racvm1

```

Now, we would like to demonstrate how the Distance to shore can influence the smooth parameter $\omega$.
We plot the estimated smooth parameter $\omega$ as a function of the angle for different values of the Distance to shore.

```{r smooth omega baseline_racvm1}
plot1=plots_bas_racvm1$fe_omega_AngleNormal_q1_Ishore
plot2=plots_bas_racvm1$fe_omega_AngleNormal_q2_Ishore
plot4=plots_bas_racvm1$fe_omega_AngleNormal_q4_Ishore

omega1 <- ggplot_build(plot1)$data[[1]]
omega2 <- ggplot_build(plot2)$data[[1]]
omega4 <- ggplot_build(plot4)$data[[1]]

omega1$Category <- "Far"
omega2$Category <- "Medium"
omega4$Category <- "Close"

# Combine the data frames
combined_omega <- rbind(omega1, omega2, omega4)

ci_omega1 <- ggplot_build(plot1)$data[[2]]
ci_omega2 <- ggplot_build(plot2)$data[[2]]
ci_omega4 <- ggplot_build(plot4)$data[[2]]

ci_omega1$Category <- "Far"
ci_omega2$Category <- "Medium"
ci_omega4$Category <- "Close"

# Combine the data frames
combined_ci_omega <- rbind(ci_omega1, ci_omega2, ci_omega4)

combined_plot <- ggplot() +
  geom_line(data=combined_omega, aes(x=x, y=y, color=Category, linetype=Category)) +
  geom_ribbon(data=combined_ci_omega, aes(x=x, ymin=ymin, ymax=ymax, fill=Category), alpha=0.1) +
  labs(color="Category", linetype="Category", x=expression(Theta), y=expression(omega)) +
  geom_vline(xintercept=c(-pi/2, pi/2), linetype="dashed", color="grey") +
  annotate("text", x = -pi/2-0.3, y = -5, label = expression(-pi/2), color = "red", size = 5) + 
  annotate("text", x = pi/2+0.2, y = -5, label = expression(pi/2), color = "red", size = 5) + 
  theme_minimal() +
  theme(
    axis.title.x = element_text(size=14),
    axis.title.y = element_text(size=14,angle=0,vjust=0.5, margin = margin(r = 10)),
    axis.text.y = element_text(size=12),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    legend.title = element_text(size=14),
    legend.text = element_text(size=12))

ggsave("omega_DistanceShore_levels.pdf", plot=combined_plot, path="baseline_racvm1", width=10, height=6, units="in")

combined_plot

```

There seems to be a rotation close to the shore when the angle $\Theta \in [-\pi,-\frac{\pi}{2}] \cup [\frac{\pi}{2},\pi]$, which means movement toward the shore. However, again the confidence intervals are too wide to really conclude.
Moreover, we fixed the smoothing penalties in the tensor splines. Changing their values can influence the shape of the estimated functions : lower values will produce functions that are more wiggily in each direction.

This model in indeed very complex and involves a large number of parameters given the amount of data points we have for the trajectories before exposure.
We can try a more parsimonious model to make estimation easier.

## Baseline with AngleNormal and Ishore in omega

We decide to remove the effect of \texttt{AngleNormal} on $\tau$ and only put covariates on $\omega$.

```{r fit baseline_racvm2,cache=TRUE}

#initial parameters
par0 <- c(0,0,2,4,0)
init_lambda=c(exp(4),exp(4),10^-6,10^-1)

#mapping for fixed smoothing penalties
new_map=list("log_lambda"=factor(c(1,2,rep(NA,2))))

#Fixed measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_pre12))

#parameters formulas
formulas <- list(mu1 = ~1 ,mu2 =~1,tau =~s(ID,bs="re"),nu=~s(ID,bs="re"),
                   omega=~te(AngleNormal,DistanceShore,k=5,bs="cs"))

# Fit baseline
baseline_racvm2<- SDE$new(formulas = formulas,data = dataBE12,type = "RACVM_SSM",
                    response = c("x","y"),par0 = par0,other_data=list("H"=H),
                    fixpar=c("mu1","mu2"))

baseline_racvm2$update_lambda(init_lambda)
baseline_racvm2$update_map(new_map)
baseline_racvm2$fit()

```

```{r get results baseline_racvm2}

estimates_bas_racvm2=as.list(baseline_racvm2$tmb_rep(),what="Est")
std_bas_racvm2=as.list(baseline_racvm2$tmb_rep(),what="Std")

```

```{r render estimates table baseline_racvm2}
# Create data frames for each set
log_sigma_obs <- combine_to_df(estimates_bas_racvm2$log_sigma_obs, 
                               std_bas_racvm2$log_sigma_obs, "log_sigma_obs")
coeff_fe <- combine_to_df(estimates_bas_racvm2$coeff_fe, 
                          std_bas_racvm2$coeff_fe, 
                          rownames(estimates_bas_racvm2$coeff_fe))
coeff_re <- combine_to_df(estimates_bas_racvm2$coeff_re, 
                          std_bas_racvm2$coeff_re, 
                          rownames(estimates_bas_racvm2$coeff_re))
log_lambda <- combine_to_df(estimates_bas_racvm2$log_lambda, 
                          std_bas_racvm2$log_lambda, 
                          rownames(estimates_bas_racvm2$log_lambda))

# Combine all data frames into one
table_data <- bind_rows(log_sigma_obs, coeff_fe,coeff_re,log_lambda)

```

```{r, results='asis',message=FALSE}
latex_table <- xtable(table_data, 
                      caption = "Parameter Estimates and Standard Errors")

# Print LaTeX table
print(latex_table, type = "latex", include.rownames = FALSE)
```
```{r CI baseline_racvm2 parameters}

post_coeff_bas_racvm2=baseline_racvm2$post_coeff(n_post=10000)
post_par_bas_racvm2=list(
  "tau"=exp(post_coeff_bas_racvm2$coeff_fe[,"tau.(Intercept)"]),
  "nu"=exp(post_coeff_bas_racvm2$coeff_fe[,"nu.(Intercept)"]),
  "omega"=post_coeff_bas_racvm2$coeff_fe[,"omega.(Intercept)"],
  "sigma_tau"=1/sqrt(exp(post_coeff_bas_racvm2$log_lambda[,"tau.s(ID)"])),
  "sigma_nu"=1/sqrt(exp(post_coeff_bas_racvm2$log_lambda[,"nu.s(ID)"])))

mean_par_bas_racvm2=lapply(post_par_bas_racvm2,mean)
sd_par_bas_racvm2=lapply(post_par_bas_racvm2,sd)
quant_par_bas_racvm2 <- lapply(post_par_bas_racvm2, 
                               quantile, probs = c(0.025, 0.975))

```

```{r CI table baseline_racvm2 parameters, message=FALSE}

parameter_names <- names(mean_par_bas_racvm2)
estimates <- mean_par_bas_racvm2
conf_intervals <- sapply(parameter_names, function(param) {
  sprintf("$[%.2f; %.2f]$", quant_par_bas_racvm2[[param]][1], 
          quant_par_bas_racvm2[[param]][2])
})

table_data <- data.frame(
  Parameter = 
    c("$\\tau$", "$\\nu$","$\\omega$","$\\sigma_{\\tau}$","$\\sigma_{\\nu}$"),
  Estimate = sprintf("$%.2f$", mean_par_bas_racvm2),
  CI = conf_intervals,
  stringsAsFactors = FALSE
)

```

```{r,results='asis',echo=FALSE}
xtab <- xtable(
  table_data,
  caption = "racvm2 baseline estimations",
  label = "table:racvm2_baseline_estimations_with_random_effects"
)

print(
  xtab,type="latex",
  include.rownames = FALSE,
  sanitize.text.function = identity,  # Prevent escaping of LaTeX math symbols
  hline.after = c(-1, 0, nrow(table_data)),  
  table.placement="H"
)
```
```{r display plots baseline_racvm2}

Ishore_values=c(1/0.25,1/2,1/5)
plots=lapply(1:length(Ishore_values),function(i){
  baseline_racvm2$plot_par(var="AngleNormal",covs=data.frame("ID"="Asgeir","Ishore"=Ishore_values[i]),
                         par_names=c("omega"),n_post=100)})
plots

```

```{r plot smooth parameters CI baseline_racvm2}
#plot parameters

xmin=list("Ishore"=1/D_up)
xmax=list("Ishore"=1/D_low)
link=list("Ishore"=(\(x) 1/x))
xlabel=list("Ishore"="Distance to shore")

plots_bas_racvm2=baseline_racvm2$get_all_plots(baseline=NULL,
      xmin=xmin, xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")

plots_bas_racvm2

```

```{r save plots baseline_racvm2}
# Create directory
if (!dir.exists("baseline_racvm2")) {
    dir.create("baseline_racvm2", recursive = TRUE)
}

# Use lapply to save each plot
lapply(seq_along(plots_bas2), function(i) {

    plot<-plots_bas2[[i]]
    plot_name <- names(plots_bas2)[i]
    file_path <- file.path("baseline_racvm2", plot_name)

    # Check if the plot is a ggplot or plotly object
    if (inherits(plot, "ggplot")) {
      ggsave(file_path, plot, device = "pdf")
    } else if (inherits(plot, "plotly")) {
      saveWidget(plot, file_path)
  }
    })
```

```{r smooth omega baseline_racvm2}
plot1=plots_bas_racvm2$fe_omega_AngleNormal_q1_Ishore
plot2=plots_bas_racvm2$fe_omega_AngleNormal_q2_Ishore
plot4=plots_bas_racvm2$fe_omega_AngleNormal_q4_Ishore

omega1 <- ggplot_build(plot1)$data[[1]]
omega2 <- ggplot_build(plot2)$data[[1]]
omega4 <- ggplot_build(plot4)$data[[1]]

omega1$Category <- "Far"
omega2$Category <- "Medium"
omega4$Category <- "Close"

# Combine the data frames
combined_omega <- rbind(omega1, omega2, omega4)

ci_omega1 <- ggplot_build(plot1)$data[[2]]
ci_omega2 <- ggplot_build(plot2)$data[[2]]
ci_omega4 <- ggplot_build(plot4)$data[[2]]

ci_omega1$Category <- "Far"
ci_omega2$Category <- "Medium"
ci_omega4$Category <- "Close"

# Combine the data frames
combined_ci_omega <- rbind(ci_omega1, ci_omega2, ci_omega4)

combined_plot <- ggplot() +
  geom_line(data=combined_omega, aes(x=x, y=y, color=Category, linetype=Category)) +
  geom_ribbon(data=combined_ci_omega, aes(x=x, ymin=ymin, ymax=ymax, fill=Category), alpha=0.1) +
  labs(color="Category", linetype="Category", x=expression(Theta), y=expression(omega)) +
  geom_vline(xintercept=c(-pi/2, pi/2), linetype="dashed", color="grey") +
  annotate("text", x = -pi/2-0.3, y = -5, label = expression(-pi/2), color = "red", size = 5) + 
  annotate("text", x = pi/2+0.2, y = -5, label = expression(pi/2), color = "red", size = 5) + 
  theme_minimal() +
  theme(
    axis.title.x = element_text(size=14),
    axis.title.y = element_text(size=14,angle=0,vjust=0.5, margin = margin(r = 10)),
    axis.text.y = element_text(size=12),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    legend.title = element_text(size=14),
    legend.text = element_text(size=12))

ggsave("omega_DistanceShore_levels.pdf", plot=combined_plot, path="baseline_racvm2", width=10, height=6, units="in")

combined_plot

```

Again, even though there seem to be a rotation in the movement close to the shore when the angle $\Theta \in [-\pi,-\frac{\pi}{2}] \cup [\frac{\pi}{2},\pi]$, the confidence intervals are too wide to conclude.
Moreover, the magnitude of the rotation is too low to effectively constrained the movement : it is estimated to be at most around $\pi$ rad/h, while the value we used in the simulation study to obtain constrained trajectories is $\frac{\pi}{2}$ rad/min, that is approximately $30 \pi $ rad/h.

## Simulate trajectories before exposure

```{r simulate baseline trajectories}
#initial narwhals positions (convert to matrix to avoid bugs)
z0_BE <- as.matrix(dataBE12[!duplicated(dataBE12$ID), c("x", "y")])

sigma_obs=0.03

baseline_ctcrw0_sim=baseline_ctcrw0$simulate(z0 = z0_BE,
                                            data=baseline_ctcrw0$data(),
                                            sd_noise=sigma_obs)

baseline_racvm0_sim=baseline_racvm0$simulate(z0 = z0_BE,
                                            data=baseline_racvm0$data(),
                                            sd_noise=sigma_obs)

baseline_racvm1_sim=baseline_racvm1$simulate(z0 = z0_BE,
                                            data=baseline_racvm1$data(),
                                            sd_noise=sigma_obs)

baseline_racvm2_sim=baseline_racvm2$simulate(z0 = z0_BE,
                                            data=baseline_racvm2$data(),
                                            sd_noise=sigma_obs)
# Add a label to each dataset for plotting
baseline_ctcrw0_sim <- baseline_ctcrw0_sim %>% 
  mutate(data_type = "Simulated baseline_ctcrw0")
baseline_racvm0_sim <- baseline_racvm0_sim %>% 
  mutate(data_type = "Simulated baseline_racvm0")
baseline_racvm1_sim <- baseline_racvm1_sim %>% 
  mutate(data_type = "Simulated baseline_racvm1")
baseline_racvm2_sim <- baseline_racvm2_sim %>% 
  mutate(data_type = "Simulated baseline_racvm2")
dataBE12 <- dataBE12 %>% mutate(data_type = "Observed")

# Combine both datasets
baseline_trajectories <- bind_rows(
  baseline_ctcrw0_sim,baseline_racvm0_sim,
  baseline_racvm1_sim,baseline_racvm2_sim, dataBE12)

# Plot both in a single plot with facets
ggplot(baseline_trajectories) +
  geom_point(aes(x = x, y = y, color = ID), size = 0.6) +
  facet_wrap(~ data_type) + 
  ggtitle("Trajectories before ship exposure") +
  theme_minimal() +
  labs(color = "Narwhal ID")

```

As in the preliminary analysis, the simulated trajectories still appear more wiggily or tortuous than the observed trajectories.
In comparison, look at what type of simulated trajectories we obtain when fixing a higher value for the persistence.

```{r simulate persistent trajectory,cache=TRUE}
#initial parameters
par0 <- c(0,0,3,4.6)

#model formula
formulas <- list(mu1=~1,mu2=~1,tau =~1,nu=~1)

#Fixed measurement error
sigma_obs=0.03
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_pre12))

#Fit baseline with measurement error estimated from the data
baseline_ctcrw_fixed<- SDE$new(formulas = formulas,data = dataBE12,type = "CTCRW",
                          response = c("x","y"),
                          par0 = par0,
                          other_data=list("H"=H),
                          fixpar=c("mu1","mu2","tau","nu"))

baseline_ctcrw_fixed_sim=baseline_ctcrw_fixed$simulate(z0 = z0_BE,
                                            data=baseline_ctcrw_fixed$data(),
                                            sd_noise=sigma_obs)

baseline_ctcrw_fixed_sim <- baseline_ctcrw_fixed_sim %>% 
  mutate(data_type = "Simulated baseline_ctcrw_fixed")

# Combine both datasets
baseline_trajectories <- bind_rows(baseline_trajectories,
 baseline_ctcrw_fixed_sim)

# Plot both in a single plot with facets
ggplot(baseline_trajectories) +
  geom_point(aes(x = x, y = y, color = ID), size = 0.6) +
  facet_wrap(~ data_type) + 
  ggtitle("Trajectories before ship exposure") +
  theme_minimal() +
  labs(color = "Narwhal ID")

```
The simple ctcrw model where we fixed all parameters to some value based on observations seems to produce trajectories that match better the observations than the complicated models that we fitted. However, here we simulated with the covariates values that are in the data. But the angle and distance to shore should be recomputed as the process evolves. We now try to simulate and compute the new value of the covariates \texttt{AngleNormal} and \texttt{Ishore} at each step. We use high frequency time step, and not the low frequency times steps in the observed data to have accurate trajectories.

```{r,cache=TRUE}

# Get land polygons
par_dir=here()
greenland_data_path <- file.path(par_dir,"Data", 
                                 "preprocessed_data","greenland")  
land<-st_read(file.path(greenland_data_path,
                        "updated_scoresby_sound_utm.shp"))
land <- st_transform(land, crs = "+init=EPSG:32626 +units=km")

# Define functions to compute covariates along the way
fangle=function(z,v,p) {

  #normal vector
  normal=c(z[1]-p[1],z[2]-p[2])

  #return angle
  return (signed_angle(normal,v))
}

fIshore=function(z,v,p) {

  #normal vector
  normal=c(z[1]-p[1],z[2]-p[2])

  #distance to shore
  Dshore=sqrt(normal[1]^2+normal[2]^2)
  if (Dshore>3){
    return (0)
  }
  else if (Dshore<0.02) {
    return (50)
  }
  else {
    return (1/Dshore)
  }
}

atw=list("AngleNormal"=fangle,"Ishore"=fIshore)

# Define time steps
delta=1/60
times=unlist(sapply(unique(dataBE12$ID),function(ID) {seq(min(dataBE12[dataBE12$ID==ID,"time"]),max(dataBE12[dataBE12$ID==ID,"time"]),by=delta)}))
n=length(times)
N=length(unique(dataBE12$ID))
data_reg=data.frame("AngleNormal"=rep(0,n*N),
                    "Ishore"=rep(1,n*N),time=c(),ID=rep(unique(dataBE12$ID),each=n))

# Simulate trajectories with small time step
baseline_racvm1_trajectory=baseline_racvm1$simulate(z0=z0_BE,data=data_reg,atw=atw,land=land,omega_times = 1)
baseline_racvm1_trajectory_adjusted=baseline_racvm1$simulate(z0=z0_BE,data=data_reg,atw=atw,land=land,omega_times = 40)

baseline_racvm2_trajectory=baseline_racvm2$simulate(z0=z0_BE,data=data_reg,atw=atw,land=land,omega_times = 1)
baseline_racvm2_trajectory_adjusted=baseline_racvm2$simulate(z0=z0_BE,data=data_reg,atw=atw,land=land,omega_times = 40)

baseline_racvm1_trajectory <- baseline_racvm1_trajectory %>% 
  mutate(data_type = "Simulated baseline_racvm1")
baseline_racvm1_trajectory_adjusted <- baseline_racvm1_trajectory_adjusted %>% 
  mutate(data_type = "Simulated baseline_racvm1 adjusted")
baseline_racvm2_trajectory <- baseline_racvm2_trajectory %>% 
  mutate(data_type = "Simulated baseline_racvm2")
baseline_racvm2_trajectory_adjusted <- baseline_racvm2_trajectory_adjusted %>% 
  mutate(data_type = "Simulated baseline_racvm2 adjusted")

baseline_trajectories <- bind_rows(baseline_racvm1_trajectory,
                                   baseline_racvm1_trajectory_adjusted,
                                   baseline_racvm2_trajectory,
                                   baseline_racvm2_trajectory_adjusted,dataBE12)

```

```{r}

# Basic ggplot setup
plot_baseline_trajectories <- ggplot() +
  geom_sf(data=land$geometry,fill="grey")+
  coord_sf(datum=st_crs("+init=EPSG:32626 +units=km"))+
  geom_path(data=baseline_trajectories, 
            aes(x = x, y = y,group=ID),size = 1, 
            lineend = "round",alpha=0.15) +     
  geom_point(data=baseline_trajectories,
             aes(x = x, y = y,group=ID,shape=ID),size =4 ) + facet_wrap(~data_type) 
  labs(title = "Narwhal Trajectories before Exposure", 
       subtitle = "Time : {frame_along} h", 
       x = "X Position", 
       y = "Y Position") +
  theme_minimal()

anim <- plot_baseline_trajectories + 
  transition_reveal(along = time) + 
  labs(subtitle = "Time: {round(frame_along,1)} hours")

# Save the animation
animate(anim, nframes = 100, fps = 5, width = 800, height = 600, 
        renderer = gifski_renderer("baseline_trajectories.gif"))

```

The adjusted SDE models where the smooth parameters $\omega$ has been multiplied by a factor $40$ produces trajectories that seem to follow the shoreline and look more similar to the actual data.
However, the trajectories obtained from the non adjusted fitted model does not produce constrained trajectories.
This suggests that we are indeed underestimating the smooth parameter $\omega$ with the tensor splines. This may be due to several factors : the low frequency at which the data is observed, the measurement error, the small amount of data that we have.

We can try to fit an interpretable parametric surface on the estimated spline surface for the parameter $\omega$ in the models \texttt{baseline\_racvm1} or \texttt{baseline\_racvm2}. However the surfaces are so different that it does not work very well.

```{r}

Ishore=seq(from=1/6,to=1/0.05,length.out=30)
theta=seq(from=-pi,to=pi,length.out=30)
grid<- as.data.frame(expand.grid(Ishore,theta))
colnames(grid) <- c("Ishore","AngleNormal")
grid$ID="Asgeir"

#get model matrices
mats=baseline_racvm1$make_mat(new_data=grid)
X_fe=mats$X_fe
X_re=mats$X_re

par_mat=baseline_racvm1$par(new_data=grid,X_fe=X_fe,X_re=X_re)

#matrix of values for surface plot
z=matrix(par_mat[,"omega"],30,30)

# Define the parametric function for omega
fomega<- function(cov_data, a =5,theta0=pi/2,D0=1) {
  Dshore=cov_data$DistanceShore
  theta=cov_data$AngleNormal
  if (is.null(Dshore)){
    Dshore=1/cov_data$Ishore
  }
  a*theta*(theta-theta0)*(theta+theta0)*exp(-Dshore/D0)/Dshore
}

data=grid
data$z=as.vector(z)

#fix magnitude and slope for omega on the AngleNormal axis
a_fixed <- 1  
theta0_fixed <- pi/2

fit_optim <- optim(par = c(log_D0 = log(1)),  # Exclude a from parameters
                   fn = function(params) {
                     log_D0 <- params[1]

                     # Use fixed values
                     a <- a_fixed
                     theta0 <- theta0_fixed

                      # Compute predicted z values
                     z_pred <- fomega(data, a,theta0,exp(log_D0))

                     # Return sum of squared residuals
                     sum((data$z - z_pred)^2)
                   })

est_par=fit_optim$par

# Compute the predicted values using the starting parameters
z_pred_starting <- fomega(grid,D0=exp(est_par))

# Reshape z_pred_starting to a matrix for plotting
z_pred_matrix <- matrix(z_pred_starting, nrow = length(Ishore), ncol = length(theta))

# Create the 3D surface plot using plot_ly
plot_ly(x = ~theta, y = ~1/Ishore, z = ~z_pred_matrix) %>%
  add_surface() %>%
  layout(title = "Parametric surface for omega (starting parameters)",
         scene = list(
           yaxis = list(title = "1/Ishore"),
           xaxis = list(title = "AngleNormal"),
           zaxis = list(title = "Predicted z")
         ))

# Create the 3D surface plot using plot_ly
plot_ly(x = ~theta, y = ~1/Ishore, z = ~z) %>%
  add_surface() %>%
  layout(title = "Predicted spline surface for omega",
         scene = list(
           yaxis = list(title = "Ishore"),
           xaxis = list(title = "AngleNormal"),
         ))

```

We continue the analysis with the fitted baseline models \texttt {baseline\_racvm1} and \texttt{baseline\_racvm2}.

# Fit response models

## Try a model similar to baseline racvm2

```{r}

par0 <- c(0,0,1,4,0)

sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))

#model formula
formulas <- list(mu1=~1,mu2=~1,tau =~s(ID,bs="re"),
                 nu=~s(ID,bs="re"),
                 omega=~te(AngleNormal,Ishore,k=5,bs="cs"))

response0<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM_SSM",response = c("x","y"),
                    par0 = par0,other_data=list("log_sigma_obs0"=log(0.05)),fixpar=c("mu1","mu2"))

response0$update_map(list("log_lambda"=factor(c(1,2,NA,NA))))
response0$update_lambda(exp(c(4,4,-5,-5)))

response0$fit()
```
```{r}
response0$get_all_plots(baseline=NULL,
      xmin=xmin, xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")
```

## Response models with splines of ExpShip

### Response models based on baseline_racvm2

```{r fit response1}

par0 <- c(0,0,1,4,0)

sigma_obs=0.03
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))

#model formula
formulas <- list(mu1=~1,mu2=~1,
                 tau =~s(ExpShip,k=3,bs="cs")+s(ID,bs="re"),
                 nu=~s(ExpShip,k=3,bs="cs")+s(ID,bs="re"),
                 omega=~ti(AngleNormal,k=5,bs="cs")+ti(Ishore,k=5,bs="cs")+
                   ti(AngleNormal,Ishore,k=c(5,5),bs="cs"))

#mapping for fixed coefficients
new_map=list(coeff_re=factor(c(1,2,rep(NA,N),
                    3,4,rep(NA,N),rep(NA,24))),coeff_fe=factor(rep(NA,5)),
                    log_lambda=factor(c(5,NA,6,NA,rep(NA,4))))
#Fit baseline
response1_sp<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("log_sigma_obs0"=log(0.05)),fixpar=c("mu1","mu2"))

new_coeff_re=c(rep(0,2),baseline_racvm2$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],rep(0,2),
               baseline_racvm2$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline_racvm2$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline_racvm2$coeff_re()[paste("omega.ti(Ishore).",1:4,sep=""),1],
               baseline_racvm2$coeff_re()[paste("omega.ti(AngleNormal,Ishore).",1:16,sep=""),1])
new_lambda=c(1,baseline_racvm2$lambda()["tau.s(ID)",1],1,baseline_racvm2$lambda()["nu.s(ID)",1],
             baseline_racvm2$lambda()[3:6,1])

response1_sp$update_map(new_map)
response1_sp$update_coeff_re(new_coeff_re)
response1_sp$update_coeff_fe(baseline_racvm2$coeff_fe()[,1])

response1_sp$update_lambda(new_lambda)

response1_sp$fit()

```

```{r get results response1}

estimates_res1_sp=as.list(response1_sp$tmb_rep(),what="Est")
std_res1_sp=as.list(response1_sp$tmb_rep(),what="Std")

#plot parameters
xmin=list("AngleNormal"=-pi,"ExpShip"=1/45)
xmax=list("AngleNormal"=pi,"ExpShip"=1/3)
link=list("ExpShip"=(\(x) 1/x))
xlabel=list("ExpShip"="Distance to ship")

plots_res1_sp=response1_sp$get_all_plots(baseline=baseline_racvm2,xmin=xmin,
                                         xmax=xmax,link=link,xlabel=xlabel,show_CI="simultaneous")

```

```{r display plots response1_sp}

#print plots
plots_res1_sp

```

```{r estimates on parameter scale response1}
post_coeff_res1_sp=response1_sp$post_coeff(n_post=1000)
post_par_res1_sp=list("tau0"=exp(post_coeff_res1_sp$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res1_sp$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res1_sp$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res1_sp$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res1_sp$log_lambda[,2])))
mean_par_res1_sp=lapply(post_par_res1_sp,mean)
sd_par_res1_sp=lapply(post_par_res1_sp,sd)

```

### Response model with splines of ExpShip based on baseline2

```{r fit response2}

par0 <- c(0,0,1,4,0)

#model formula
formulas <- list(mu1=~1,mu2=~1,tau =~s(ExpShip,k=3,bs="cs")+s(ID,bs="re"),
                 nu=~s(ExpShip,k=3,bs="cs")+s(ID,bs="re"),
                 omega=~ti(DistanceShore,k=5,bs="cs")+ti(AngleNormal,k=5,bs="cs")+ti(AngleNormal,DistanceShore,k=c(5,5),bs="cs"))

#mapping for fixed coefficients
new_map=list(coeff_re=factor(c("tau.s(ExpShip).1","tau.s(ExpShip).2",
        rep(NA,N),"nu.s(ExpShip).1","nu.s(ExpShip).2",rep(NA,N),rep(NA,24))),
        coeff_fe=factor(rep(NA,5)),log_lambda=factor(c("tau.s(ExpShip)",NA,"nu.s(ExpShip)",NA,rep(NA,4))))

#Fit baseline
response2_sp<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("log_sigma_obs0"=log(0.05)),fixpar=c("mu1","mu2"))

new_coeff_re=c(rep(0,2),baseline2$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline2$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(1,baseline2$lambda()["tau.s(ID)",1],1,baseline2$lambda()["nu.s(ID)",1],baseline2$lambda()[3:6,1])

response2_sp$update_map(new_map)
response2_sp$update_coeff_re(new_coeff_re)
response2_sp$update_coeff_fe(baseline2$coeff_fe()[,1])
response2_sp$update_lambda(new_lambda)

response2_sp$fit()

# Fit response with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response2_sp_50m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"))

new_coeff_re=c(rep(0,2),baseline2_50m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline2_50m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2_50m$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2_50m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2_50m$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(1,baseline2_50m$lambda()["tau.s(ID)",1],1,baseline2_50m$lambda()["nu.s(ID)",1],baseline2_50m$lambda()[3:6,1])

response2_sp_50m$update_map(new_map)
response2_sp_50m$update_coeff_re(new_coeff_re)
response2_sp_50m$update_coeff_fe(baseline2_50m$coeff_fe()[,1])

response2_sp_50m$update_lambda(new_lambda)

response2_sp_50m$fit()

# Fit response with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response2_sp_75m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"))

new_coeff_re=c(rep(0,2),baseline2_75m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline2_75m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2_75m$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2_75m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2_75m$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(1,baseline2_75m$lambda()["tau.s(ID)",1],1,baseline2_75m$lambda()["nu.s(ID)",1],baseline2_75m$lambda()[3:6,1])

response2_sp_75m$update_map(new_map)
response2_sp_75m$update_coeff_re(new_coeff_re)
response2_sp_75m$update_coeff_fe(baseline2_75m$coeff_fe()[,1])

response2_sp_75m$update_lambda(new_lambda)

response2_sp_75m$fit()

# Fit response with 100m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response2_sp_100m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",response = c("x","y"),
                    par0 = par0,other_data=list("H"=H),fixpar=c("mu1","mu2"))

new_coeff_re=c(rep(0,2),baseline2_100m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline2_100m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2_100m$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2_100m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2_100m$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(1,baseline2_100m$lambda()["tau.s(ID)",1],1,baseline2_100m$lambda()["nu.s(ID)",1],baseline2_100m$lambda()[3:6,1])

response2_sp_100m$update_map(new_map)
response2_sp_100m$update_coeff_re(new_coeff_re)
response2_sp_100m$update_coeff_fe(baseline2_100m$coeff_fe()[,1])

response2_sp_100m$update_lambda(new_lambda)

response2_sp_100m$fit()
```

```{r get results response2}

estimates_res2_sp=as.list(response2_sp$tmb_rep(),what="Est")
std_res2_sp=as.list(response2_sp$tmb_rep(),what="Std")

estimates_res2_sp_50m=as.list(response2_sp_50m$tmb_rep(),what="Est")
std_res2_sp_50m=as.list(response2_sp_50m$tmb_rep(),what="Std")

estimates_res2_sp_75m=as.list(response2_sp_75m$tmb_rep(),what="Est")
std_res2_sp_75m=as.list(response2_sp_75m$tmb_rep(),what="Std")

estimates_res2_sp_100m=as.list(response2_sp_100m$tmb_rep(),what="Est")
std_res2_sp_100m=as.list(response2_sp_100m$tmb_rep(),what="Std")

#plot parameters
xmin=list("AngleNormal"=-pi,"ExpShip"=1/45)
xmax=list("AngleNormal"=pi,"ExpShip"=1/3)
link=list("ExpShip"=(\(x) 1/x))
xlabel=list("ExpShip"="Distance to ship")

plot_res2_sp=response2_sp$get_all_plots(baseline=baseline2,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")
plot_res2_sp_50m=response2_sp_50m$get_all_plots(baseline=baseline2_50m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")
plot_res2_sp_75m=response2_sp_75m$get_all_plots(baseline=baseline2_75m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")
plot_res2_sp_100m=response2_sp_100m$get_all_plots(baseline=baseline2_100m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")
```

```{r display plots response2_sp}

#print plots
plot_res2_sp
plot_res2_sp_50m
plot_res2_sp_75m
plot_res2_sp_100m

```

```{r}

post_coeff_res2_sp=response2_sp$post_coeff(n_post=1000)
post_par_res2_sp=list("tau0"=exp(post_coeff_res2_sp$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res2_sp$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res2_sp$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res2_sp$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res2_sp$log_lambda[,2])))
mean_par_res2_sp=lapply(post_par_res2_sp,mean)
sd_par_res2_sp=lapply(post_par_res2_sp,sd)

post_coeff_res2_sp_50m=response2_sp_50m$post_coeff(n_post=1000)
post_par_res2_sp_50m=list("tau0"=exp(post_coeff_res2_sp_50m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res2_sp_50m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res2_sp_50m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res2_sp_50m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res2_sp_50m$log_lambda[,2])))
mean_par_res2_sp_50m=lapply(post_par_res2_sp_50m,mean)
sd_par_res2_sp_50m=lapply(post_par_res2_sp_50m,sd)

post_coeff_res2_sp_75m=response2_sp_75m$post_coeff(n_post=1000)
post_par_res2_sp_75m=list("tau0"=exp(post_coeff_res2_sp_75m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res2_sp_75m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res2_sp_75m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res2_sp_75m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res2_sp_75m$log_lambda[,2])))
mean_par_res2_sp_75m=lapply(post_par_res2_sp_75m,mean)
sd_par_res2_sp_75m=lapply(post_par_res2_sp_75m,sd)

post_coeff_res2_sp_100m=response2_sp_100m$post_coeff(n_post=1000)
post_par_res2_sp_100m=list("tau0"=exp(post_coeff_res2_sp_100m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res2_sp_100m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res2_sp_100m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res2_sp_100m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res2_sp_100m$log_lambda[,2])))
mean_par_res2_sp_100m=lapply(post_par_res2_sp_100m,mean)
sd_par_res2_sp_100m=lapply(post_par_res2_sp_100m,sd)

```

### Response model with splines of ExpShip based on baseline_racvm1

```{r fit response3}

#initialize parameters
par0=c(0,0,1.5,4.5,0)

#define model
formulas <- list(mu1=~1,mu2=~1,tau =~s(ExpShip,k=3,bs="cs")+s(ID,bs="re"),
                 nu=~s(ExpShip,k=3,bs="cs")+s(ID,bs="re"),
                 omega=~ti(AngleNormal,k=5,bs="cs")+ti(ExpShore,k=5,bs="cs")+ti(AngleNormal,ExpShore,k=c(5,5),bs="cs"))

#mapping for fixed coeffs
new_map=list(coeff_re=factor(c("tau.s(ExpShip).1","tau.s(ExpShip).2",rep(NA,N),
                    "nu.s(ExpShip).1","nu.s(ExpShip).2",rep(NA,N),rep(NA,24))),coeff_fe=factor(rep(NA,5)),
                    log_lambda=factor(c("tau.s(ExpShip)",NA,"nu.s(ExpShip)",NA,rep(NA,4))))

#Fit response
response3_sp<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("log_sigma_obs0"=log(0.07)))

new_coeff_re=c(rep(0,2),baseline_racvm3$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline_racvm3$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline_racvm3$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline_racvm3$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline_racvm3$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(1,baseline_racvm3$lambda()["tau.s(ID)",1],1,baseline_racvm3$lambda()["nu.s(ID)",1],baseline_racvm3$lambda()[3:6,1])

response3_sp$update_map(new_map)
response3_sp$update_coeff_re(new_coeff_re)
response3_sp$update_coeff_fe(baseline_racvm3$coeff_fe()[,1])

response3_sp$update_lambda(new_lambda)
response3_sp$fit()

#Fit response with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_sp_50m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_re=c(rep(0,2),baseline_racvm3_50m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline_racvm3_50m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline_racvm3_50m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline_racvm3_50m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline_racvm3_50m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(1,baseline_racvm3_50m$lambda()["tau.s(ID)",1],1,baseline_racvm3_50m$lambda()["nu.s(ID)",1],baseline_racvm3_50m$lambda()[3:6,1])

response3_sp_50m$update_map(new_map)
response3_sp_50m$update_coeff_re(new_coeff_re)
response3_sp_50m$update_coeff_fe(baseline_racvm3_50m$coeff_fe()[,1])

response3_sp_50m$update_lambda(new_lambda)
response3_sp_50m$fit()

#Fit response with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_sp_75m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_re=c(rep(0,2),baseline_racvm3_75m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline_racvm3_75m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline_racvm3_75m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline_racvm3_75m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline_racvm3_75m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(1,baseline_racvm3_75m$lambda()["tau.s(ID)",1],1,baseline_racvm3_75m$lambda()["nu.s(ID)",1],baseline_racvm3_75m$lambda()[3:6,1])

response3_sp_75m$update_map(new_map)
response3_sp_75m$update_coeff_re(new_coeff_re)
response3_sp_75m$update_coeff_fe(baseline_racvm3_75m$coeff_fe()[,1])

response3_sp_75m$update_lambda(new_lambda)
response3_sp_75m$fit()

#Fit response with 100m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_sp_100m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_re=c(rep(0,2),baseline_racvm3_100m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],
               rep(0,2),baseline_racvm3_100m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline_racvm3_100m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline_racvm3_100m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline_racvm3_100m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(1,baseline_racvm3_100m$lambda()["tau.s(ID)",1],1,baseline_racvm3_100m$lambda()["nu.s(ID)",1],baseline_racvm3_100m$lambda()[3:6,1])

response3_sp_100m$update_map(new_map)
response3_sp_100m$update_coeff_re(new_coeff_re)
response3_sp_100m$update_coeff_fe(baseline_racvm3_100m$coeff_fe()[,1])

response3_sp_100m$update_lambda(new_lambda)
response3_sp_100m$fit()
```

```{r get results response3}

estimates_res3_sp=as.list(response3_sp$tmb_rep(),what="Est")
std_res3_sp=as.list(response3_sp$tmb_rep(),what="Std")

estimates_res3_sp_50m=as.list(response3_sp_50m$tmb_rep(),what="Est")
std_res3_sp_50m=as.list(response3_sp_50m$tmb_rep(),what="Std")

estimates_res3_sp_75m=as.list(response3_sp_75m$tmb_rep(),what="Est")
std_res3_sp_75m=as.list(response3_sp_75m$tmb_rep(),what="Std")

estimates_res3_sp_100m=as.list(response3_sp_100m$tmb_rep(),what="Est")
std_res3_sp_100m=as.list(response3_sp_100m$tmb_rep(),what="Std")

#plot parameters
xmin=list("ExpShore"=1/D_up,"AngleNormal"=-pi,"ExpShip"=1/45)
xmax=list("ExpShore"=1/D_low,"AngleNormal"=pi,"ExpShip"=1/3)
link=list("ExpShore"=(\(x) 1/x),"ExpShip"=(\(x) 1/x))
xlabel=list("ExpShore"="Distance to shore","ExpShip"="Distance to ship")

plot_res3_sp=response3_sp$get_all_plots(baseline=baseline_racvm3,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")
plot_res3_sp_50m=response3_sp_50m$get_all_plots(baseline=baseline_racvm3_50m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")
plot_res3_sp_75m=response3_sp_75m$get_all_plots(baseline=baseline_racvm3_75m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")
plot_res3_sp_100m=response3_sp_100m$get_all_plots(baseline=baseline_racvm3_100m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")

```

```{r display plots response3_sp}

#print plots
plot_res3_sp
plot_res3_sp_50m
plot_res3_sp_75m
plot_res3_sp_100m

```

```{r estimates on parameter scale response3}

post_coeff_res3_sp=response3_sp$post_coeff(n_post=1000)
post_par_res3_sp=list("tau0"=exp(post_coeff_res3_sp$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res3_sp$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res3_sp$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res3_sp$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res3_sp$log_lambda[,2])))
mean_par_res3_sp=lapply(post_par_res3_sp,mean)
sd_par_res3_sp=lapply(post_par_res3_sp,sd)

post_coeff_res3_sp_50m=response3_sp_50m$post_coeff(n_post=1000)
post_par_res3_sp_50m=list("tau0"=exp(post_coeff_res3_sp_50m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res3_sp_50m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res3_sp_50m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res3_sp_50m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res3_sp_50m$log_lambda[,2])))
mean_par_res3_sp_50m=lapply(post_par_res3_sp_50m,mean)
sd_par_res3_sp_50m=lapply(post_par_res3_sp_50m,sd)

post_coeff_res3_sp_75m=response3_sp_75m$post_coeff(n_post=1000)
post_par_res3_sp_75m=list("tau0"=exp(post_coeff_res3_sp_75m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res3_sp_75m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res3_sp_75m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res3_sp_75m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res3_sp_75m$log_lambda[,2])))
mean_par_res3_sp_75m=lapply(post_par_res3_sp_75m,mean)
sd_par_res3_sp_75m=lapply(post_par_res3_sp_75m,sd)

post_coeff_res3_sp_100m=response3_sp_100m$post_coeff(n_post=1000)
post_par_res3_sp_100m=list("tau0"=exp(post_coeff_res3_sp_100m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res3_sp_100m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res3_sp_100m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res3_sp_100m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res3_sp_100m$log_lambda[,2])))
mean_par_res3_sp_100m=lapply(post_par_res3_sp_100m,mean)
sd_par_res3_sp_100m=lapply(post_par_res3_sp_100m,sd)

```

### Response model with exponential of ExpShip based on baseline2

```{r}

par0=c(0,0,1,4,0)

#define model
formulas <- list(mu1=~1,mu2=~1,tau =~ExpShip+s(ID,bs="re"),
                 nu=~ExpShip+s(ID,bs="re"),
                 omega=~ti(DistanceShore,k=5,bs="cs")+ti(AngleNormal,k=5,bs="cs")
                 +ti(AngleNormal,DistanceShore,k=c(5,5),bs="cs"))

new_map=list(coeff_re=factor(c(rep(NA,12),rep(NA,24))),coeff_fe=factor(c(NA,NA,NA,1,NA,2,NA)),
                             log_lambda=factor(c(NA,NA,NA,NA,NA,NA)))

# Fit response
response2_exp<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),
                    other_data=list("log_sigma_obs0"=log(0.05)))

new_coeff_fe=c(baseline2$coeff_fe()[1:3,1],0,baseline2$coeff_fe()[4,1],0,baseline2$coeff_fe()[5,1])
new_coeff_re=c(baseline2$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],baseline2$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(baseline2$lambda()["tau.s(ID)",1],baseline2$lambda()["nu.s(ID)",1],baseline2$lambda()[3:6,1])

response2_exp$update_map(new_map)
response2_exp$update_coeff_re(new_coeff_re)
response2_exp$update_coeff_fe(new_coeff_fe)

response2_exp$update_lambda(new_lambda)

response2_exp$fit()

# Fit response with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response2_exp_50m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_fe=c(baseline2_50m$coeff_fe()[1:3,1],0,baseline2_50m$coeff_fe()[4,1],0,baseline2_50m$coeff_fe()[5,1])
new_coeff_re=c(baseline2_50m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],baseline2_50m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2_50m$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2_50m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2_50m$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(baseline2_50m$lambda()["tau.s(ID)",1],baseline2_50m$lambda()["nu.s(ID)",1],baseline2_50m$lambda()[3:6,1])

response2_exp_50m$update_map(new_map)
response2_exp_50m$update_coeff_re(new_coeff_re)
response2_exp_50m$update_coeff_fe(new_coeff_fe)

response2_exp_50m$update_lambda(new_lambda)

response2_exp_50m$fit()

# Fit response with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response2_exp_75m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_fe=c(baseline2_75m$coeff_fe()[1:3,1],0,baseline2_75m$coeff_fe()[4,1],0,baseline2_75m$coeff_fe()[5,1])
new_coeff_re=c(baseline2_75m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],baseline2_75m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2_75m$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2_75m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2_75m$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(baseline2_75m$lambda()["tau.s(ID)",1],baseline2_75m$lambda()["nu.s(ID)",1],baseline2_75m$lambda()[3:6,1])

response2_exp_75m$update_map(new_map)
response2_exp_75m$update_coeff_re(new_coeff_re)
response2_exp_75m$update_coeff_fe(new_coeff_fe)
response2_exp_75m$update_lambda(new_lambda)

response2_exp_75m$fit()

# Fit response with 100m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response2_exp_100m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_fe=c(baseline2_100m$coeff_fe()[1:3,1],0,baseline2_100m$coeff_fe()[4,1],0,baseline2_100m$coeff_fe()[5,1])
new_coeff_re=c(baseline2_100m$coeff_re()[paste("tau.s(ID).",1:N,sep=""),1],baseline2_100m$coeff_re()[paste("nu.s(ID).",1:N,sep=""),1],
               baseline2_100m$coeff_re()[paste("omega.ti(DistanceShore).",1:4,sep=""),1],
               baseline2_100m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline2_100m$coeff_re()[paste("omega.ti(AngleNormal,DistanceShore).",1:16,sep=""),1])
new_lambda=c(baseline2_100m$lambda()["tau.s(ID)",1],baseline2_100m$lambda()["nu.s(ID)",1],baseline2_100m$lambda()[3:6,1])

response2_exp_100m$update_map(new_map)
response2_exp_100m$update_coeff_re(new_coeff_re)
response2_exp_100m$update_coeff_fe(new_coeff_fe)

response2_exp_100m$update_lambda(new_lambda)

response2_exp_100m$fit()
```

```{r}

estimates_res2_exp=as.list(response2_exp$tmb_rep(),what="Est")
std_res2_exp=as.list(response2_exp$tmb_rep(),what="Std")

estimates_res2_exp_50m=as.list(response2_exp_50m$tmb_rep(),what="Est")
std_res2_exp_50m=as.list(response2_exp_50m$tmb_rep(),what="Std")

estimates_res2_exp_75m=as.list(response2_exp_75m$tmb_rep(),what="Est")
std_res2_exp_75m=as.list(response2_exp_75m$tmb_rep(),what="Std")

estimates_res2_exp_100m=as.list(response2_exp_100m$tmb_rep(),what="Est")
std_res2_exp_100m=as.list(response2_exp_100m$tmb_rep(),what="Std")

#plot parameters
xmin=list("ExpShore"=1/D_up,"AngleNormal"=-pi,"ExpShip"=1/60)
xmax=list("ExpShore"=1/D_low,"AngleNormal"=pi,"ExpShip"=1/5)
link=list("ExpShore"=(\(x) 1/x),"ExpShip"=(\(x) 1/x))
xlabel=list("ExpShore"="Distance to shore","ExpShip"="Distance to ship")

plots_res2_exp=response2_exp$get_all_plots(baseline=baseline2,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")

plots_res2_exp_50m=response2_exp_50m$get_all_plots(baseline=baseline2_50m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")

plots_res2_exp_75m=response2_exp_75m$get_all_plots(baseline=baseline2_75m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")

plots_res2_exp_100m=response2_exp_100m$get_all_plots(baseline=baseline2_100m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")

```

```{r display plots response3_sp}

#print plots
plots_res2_exp
plots_res2_exp_50m
plots_res2_exp_75m
plots_res2_exp_100m

```

### Response model with exponential of ExpShip based on baseline_racvm3

```{r}

par0=c(0,0,1,4.5,0)

#define model
formulas <- list(mu1=~1,mu2=~1,tau =~ExpShip+s(ID,bs="re"),
                 nu=~ExpShip+s(ID,bs="re"),
                 omega=~ti(AngleNormal,k=5,bs="cs")+ti(ExpShore,k=5,bs="cs")+ti(AngleNormal,ExpShore,k=c(5,5),bs="cs"))

#mapping for fixed coeffs
new_map=list(coeff_re=factor(c(rep(NA,12),rep(NA,24))),coeff_fe=factor(c(NA,NA,NA,1,NA,2,NA)),
                             log_lambda=factor(c(NA,NA,NA,NA,NA,NA)))

#Fit response
sigma_obs=exp(estimates_bas3$log_sigma_obs)
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_exp<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_fe=c(baseline_racvm3$coeff_fe()[1:3,1],0,baseline_racvm3$coeff_fe()[4,1],0,baseline_racvm3$coeff_fe()[5,1])
new_coeff_re=c(baseline_racvm3$coeff_re()[paste("tau.s(ID).",1:6,sep=""),1],baseline_racvm3$coeff_re()[paste("nu.s(ID).",1:6,sep=""),1],
               baseline_racvm3$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline_racvm3$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline_racvm3$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(baseline_racvm3$lambda()["tau.s(ID)",1],baseline_racvm3$lambda()["nu.s(ID)",1],baseline_racvm3$lambda()[3:6,1])

response3_exp$update_map(new_map)
response3_exp$update_coeff_re(new_coeff_re)
response3_exp$update_coeff_fe(new_coeff_fe)

response3_exp$update_lambda(new_lambda)
response3_exp$fit()

# Fit response with 30m measurement error
sigma_obs=0.03
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_exp_30m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_fe=c(baseline_racvm3_30m$coeff_fe()[1:3,1],0,baseline_racvm3_30m$coeff_fe()[4,1],0,baseline_racvm3_30m$coeff_fe()[5,1])
new_coeff_re=c(baseline_racvm3_30m$coeff_re()[paste("tau.s(ID).",1:6,sep=""),1],baseline_racvm3_30m$coeff_re()[paste("nu.s(ID).",1:6,sep=""),1],
               baseline_racvm3_30m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline_racvm3_30m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline_racvm3_30m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(baseline_racvm3_30m$lambda()["tau.s(ID)",1],baseline_racvm3_30m$lambda()["nu.s(ID)",1],baseline_racvm3_30m$lambda()[3:6,1])

response3_exp_30m$update_map(new_map)
response3_exp_30m$update_coeff_re(new_coeff_re)
response3_exp_30m$update_coeff_fe(new_coeff_fe)

response3_exp_30m$update_lambda(new_lambda)
response3_exp_30m$fit()

# Fit response with 40m measurement error
sigma_obs=0.04
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_exp_40m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_fe=c(baseline_racvm3_40m$coeff_fe()[1:3,1],0,baseline_racvm3_40m$coeff_fe()[4,1],0,baseline_racvm3_40m$coeff_fe()[5,1])
new_coeff_re=c(baseline_racvm3_40m$coeff_re()[paste("tau.s(ID).",1:6,sep=""),1],baseline_racvm3_40m$coeff_re()[paste("nu.s(ID).",1:6,sep=""),1],
               baseline_racvm3_40m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline_racvm3_40m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline_racvm3_40m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(baseline_racvm3_40m$lambda()["tau.s(ID)",1],baseline_racvm3_40m$lambda()["nu.s(ID)",1],baseline_racvm3_40m$lambda()[3:6,1])

response3_exp_40m$update_map(new_map)
response3_exp_40m$update_coeff_re(new_coeff_re)
response3_exp_40m$update_coeff_fe(new_coeff_fe)

response3_exp_40m$update_lambda(new_lambda)
response3_exp_40m$fit()

# Fit response with 45m measurement error
sigma_obs=0.045
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_exp_45m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_fe=c(baseline_racvm3_45m$coeff_fe()[1:3,1],0,baseline_racvm3_45m$coeff_fe()[4,1],0,baseline_racvm3_45m$coeff_fe()[5,1])
new_coeff_re=c(baseline_racvm3_45m$coeff_re()[paste("tau.s(ID).",1:6,sep=""),1],baseline_racvm3_45m$coeff_re()[paste("nu.s(ID).",1:6,sep=""),1],
               baseline_racvm3_45m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline_racvm3_45m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline_racvm3_45m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(baseline_racvm3_45m$lambda()["tau.s(ID)",1],baseline_racvm3_45m$lambda()["nu.s(ID)",1],baseline_racvm3_45m$lambda()[3:6,1])

response3_exp_45m$update_map(new_map)
response3_exp_45m$update_coeff_re(new_coeff_re)
response3_exp_45m$update_coeff_fe(new_coeff_fe)
response3_exp_45m$update_lambda(new_lambda)

response3_exp_45m$fit()

# Fit response with 50m measurement error
sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_exp_50m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_fe=c(baseline_racvm3_50m$coeff_fe()[1:3,1],0,baseline_racvm3_50m$coeff_fe()[4,1],0,baseline_racvm3_50m$coeff_fe()[5,1])
new_coeff_re=c(baseline_racvm3_50m$coeff_re()[paste("tau.s(ID).",1:6,sep=""),1],baseline_racvm3_50m$coeff_re()[paste("nu.s(ID).",1:6,sep=""),1],
               baseline_racvm3_50m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline_racvm3_50m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline_racvm3_50m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(baseline_racvm3_50m$lambda()["tau.s(ID)",1],baseline_racvm3_50m$lambda()["nu.s(ID)",1],baseline_racvm3_50m$lambda()[3:6,1])

response3_exp_50m$update_map(new_map)
response3_exp_50m$update_coeff_re(new_coeff_re)
response3_exp_50m$update_coeff_fe(new_coeff_fe)

response3_exp_50m$update_lambda(new_lambda)
response3_exp_50m$fit()

# Fit response with 75m measurement error
sigma_obs=0.075
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_exp_75m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_fe=c(baseline_racvm3_75m$coeff_fe()[1:3,1],0,baseline_racvm3_75m$coeff_fe()[4,1],0,baseline_racvm3_75m$coeff_fe()[5,1])
new_coeff_re=c(baseline_racvm3_75m$coeff_re()[paste("tau.s(ID).",1:6,sep=""),1],baseline_racvm3_75m$coeff_re()[paste("nu.s(ID).",1:6,sep=""),1],
               baseline_racvm3_75m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline_racvm3_75m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline_racvm3_75m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(baseline_racvm3_75m$lambda()["tau.s(ID)",1],baseline_racvm3_75m$lambda()["nu.s(ID)",1],baseline_racvm3_75m$lambda()[3:6,1])

response3_exp_75m$update_map(new_map)
response3_exp_75m$update_coeff_fe(new_coeff_fe)
response3_exp_75m$update_lambda(new_lambda)

response3_exp_75m$fit()

# Fit response with 100m measurement error
sigma_obs=0.1
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))
response3_exp_100m<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                    response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H))

new_coeff_fe=c(baseline_racvm3_100m$coeff_fe()[1:3,1],0,baseline_racvm3_100m$coeff_fe()[4,1],0,baseline_racvm3_100m$coeff_fe()[5,1])
new_coeff_re=c(baseline_racvm3_100m$coeff_re()[paste("tau.s(ID).",1:6,sep=""),1],baseline_racvm3_100m$coeff_re()[paste("nu.s(ID).",1:6,sep=""),1],
               baseline_racvm3_100m$coeff_re()[paste("omega.ti(AngleNormal).",1:4,sep=""),1],
               baseline_racvm3_100m$coeff_re()[paste("omega.ti(ExpShore).",1:4,sep=""),1],
               baseline_racvm3_100m$coeff_re()[paste("omega.ti(AngleNormal,ExpShore).",1:16,sep=""),1])
new_lambda=c(baseline_racvm3_100m$lambda()["tau.s(ID)",1],baseline_racvm3_100m$lambda()["nu.s(ID)",1],baseline_racvm3_100m$lambda()[3:6,1])

response3_exp_100m$update_map(new_map)
response3_exp_100m$update_coeff_re(new_coeff_re)
response3_exp_100m$update_coeff_fe(new_coeff_fe)
response3_exp_100m$update_lambda(new_lambda)

response3_exp_100m$fit()
```

```{r}

estimates_res3_exp=as.list(response3_exp$tmb_rep(),what="Est")
std_res3_exp=as.list(response3_exp$tmb_rep(),what="Std")

estimates_res3_exp_30m=as.list(response3_exp_30m$tmb_rep(),what="Est")
std_res3_exp_30m=as.list(response3_exp_30m$tmb_rep(),what="Std")

estimates_res3_exp_40m=as.list(response3_exp_40m$tmb_rep(),what="Est")
std_res3_exp_40m=as.list(response3_exp_40m$tmb_rep(),what="Std")

estimates_res3_exp_45m=as.list(response3_exp_45m$tmb_rep(),what="Est")
std_res3_exp_45m=as.list(response3_exp_45m$tmb_rep(),what="Std")

estimates_res3_exp_50m=as.list(response3_exp_50m$tmb_rep(),what="Est")
std_res3_exp_50m=as.list(response3_exp_50m$tmb_rep(),what="Std")

estimates_res3_exp_75m=as.list(response3_exp_75m$tmb_rep(),what="Est")
std_res3_exp_75m=as.list(response3_exp_75m$tmb_rep(),what="Std")

estimates_res3_exp_100m=as.list(response3_exp_100m$tmb_rep(),what="Est")
std_res3_exp_100m=as.list(response3_exp_100m$tmb_rep(),what="Std")

#plot parameters
xmin=list("ExpShore"=1/D_up,"AngleNormal"=-pi,"ExpShip"=1/60)
xmax=list("ExpShore"=1/D_low,"AngleNormal"=pi,"ExpShip"=1/5)
link=list("ExpShore"=(\(x) 1/x),"ExpShip"=(\(x) 1/x))
xlabel=list("ExpShore"="Distance to shore","ExpShip"="Distance to ship")

plots_res3_exp=response3_exp$get_all_plots(baseline=baseline_racvm3,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")

plots_res3_exp_30m=response3_exp_30m$get_all_plots(baseline=baseline_racvm3_30m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")

plots_res3_exp_40m=response3_exp_40m$get_all_plots(baseline=baseline_racvm3_40m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")

plots_res3_exp_50m=response3_exp_50m$get_all_plots(baseline=baseline_racvm3_50m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")
plots_res3_exp_75m=response3_exp_75m$get_all_plots(baseline=baseline_racvm3_75m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")
plots_res3_exp_100m=response3_exp_100m$get_all_plots(baseline=baseline_racvm3_100m,xmin=xmin,
                            xmax=xmax,link=link,xlabel=xlabel,show_CI="pointwise")

```

```{r display plots response3_sp}

#print plots
plots_res3_exp
plots_res3_exp_50m
plots_res3_exp_75m
plots_res3_exp_100m

```

We select the response 3 model with 50m measurement error

```{r}
# Create directory
if (!dir.exists("response3_exp_50m")) {
    dir.create("response3_exp_50m", recursive = TRUE)
}

# Use lapply to save each plot
lapply(seq_along(plots_bas2), function(i) {

    plot<-plots_bas2[[i]]
    plot_name <- names(plots_bas2)[i]
    file_path <- file.path("response3_exp_50m", plot_name)

    # Check if the plot is a ggplot or plotly object
    if (inherits(plot, "ggplot")) {
      ggsave(file_path, plot, device = "pdf")
    } else if (inherits(plot, "plotly")) {
      saveWidget(plot, file_path)
  }
    })
```

### Plots improvements

```{r}

#get angular velocity for different distances to shore
plot_me_tau <- plots_res3_exp_50m$me_tau_ExpShip
plot_me_nu <- plots_res3_exp_50m$me_nu_ExpShip

# Update plot_me_tau
plot_me_tau <- plot_me_tau +
  labs(y = expression(tau)) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size=18),
    axis.title.y = element_text(size=18, angle=0, vjust=0.5, margin = margin(r = 10)),
    axis.text.x = element_text(size=14),
    axis.text.y = element_text(size=14),
    legend.title = element_text(size=14),
    legend.text = element_text(size=14)
  )

# Update plot_me_nu
plot_me_nu <- plot_me_nu +
  labs(y = expression(nu)) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size=18),
    axis.title.y = element_text(size=18, angle=0, vjust=0.5, margin = margin(r = 10)),
    axis.text.x = element_text(size=14),
    axis.text.y = element_text(size=14),
    legend.title = element_text(size=14),
    legend.text = element_text(size=14)
  )

# Save the updated plots
ggsave("plot_me_tau_updated.pdf", plot=plot_me_tau, width=10, height=6, units="in")
ggsave("plot_me_nu_updated.pdf", plot=plot_me_nu, width=10, height=6, units="in")

```

## Posterior samples

```{r}
post_coeff_res3_exp_50m=response3_exp_50m$post_coeff(n_post=10000)
post_par_res3_exp_50m=list(
              "tau0"=exp(post_coeff_res3_exp_50m$coeff_fe[,"tau.(Intercept)"]),
              "nu0"=exp(post_coeff_res3_exp_50m$coeff_fe[,"nu.(Intercept)"]),
              "omega0"=post_coeff_res3_exp_50m$coeff_fe[,"omega.(Intercept)"],
              "sigma_tau"=1/sqrt(exp(post_coeff_res3_exp_50m$log_lambda[,1])),
              "sigma_nu"=1/sqrt(exp(post_coeff_res3_exp_50m$log_lambda[,2])),
              "tau.ExpShip"=post_coeff_res3_exp_50m$coeff_fe[,"tau.ExpShip"],
              "nu.ExpShip"=post_coeff_res3_exp_50m$coeff_fe[,"nu.ExpShip"])
mean_par_res3=lapply(post_par_res3_exp_50m,mean)
sd_par_res3=lapply(post_par_res3_exp_50m,sd)
quant_par_res3 <- lapply(post_par_res3_exp_50m, quantile, probs = c(0.025, 0.975))

```

#### Thorough uncertainty quantification for response3_exp_50m

```{r}

#posterior samples of the coefficient of the baseline model
post_coeff_re=post_coeff_bas3_50m$coeff_re[1:100,]
post_coeff_fe=post_coeff_bas3_50m$coeff_fe[1:100,]
post_coeff_lambda=exp(post_coeff_bas3_50m$log_lambda[1:100,])

# Fit response with 50m measurement error

par0=c(0,0,1,4.5,0)
formulas <- list(mu1=~1,mu2=~1,tau =~ExpShip+s(ID,bs="re"),
                 nu=~ExpShip+s(ID,bs="re"),
                 omega=~ti(AngleNormal,k=5,bs="cs")+ti(ExpShore,k=5,bs="cs")+ti(AngleNormal,ExpShore,k=c(5,5),bs="cs"))

sigma_obs=0.05
H=array(rep(sigma_obs^2*diag(2),n_post),dim=c(2,2,n_post))

estimates=matrix(rep(0,2*nrow(post_coeff_fe)),ncol=2)
std=matrix(rep(0,2*nrow(post_coeff_fe)),ncol=2)

for (i in 1:nrow(post_coeff_fe)) {

    response3_exp_50m_temp<- SDE$new(formulas = formulas,data = dataAE,type = "RACVM",
                      response = c("x","y"),par0 = par0,fixpar=c("mu1","mu2"),other_data=list("H"=H),
                      map=list(coeff_re=factor(c(rep(NA,12),rep(NA,24))),coeff_fe=factor(c(NA,NA,NA,1,NA,2,NA)),
                               log_lambda=factor(c(NA,NA,NA,NA,NA,NA))))

    new_coeff_fe=c(post_coeff_fe[i,1:3],0,post_coeff_fe[i,4],0,post_coeff_fe[i,5])
    new_coeff_re=post_coeff_re[i,]
    new_lambda=post_coeff_lambda[i,]
    response3_exp_50m_temp$update_coeff_re(new_coeff_re)
    response3_exp_50m_temp$update_coeff_fe(new_coeff_fe)

    response3_exp_50m_temp$update_lambda(new_lambda)
    response3_exp_50m_temp$fit()

    est_alpha=as.list(response3_exp_50m_temp$tmb_rep(),what="Est")$coeff_fe[c("tau.ExpShip","nu.ExpShip"),1]
    std_alpha=as.list(response3_exp_50m_temp$tmb_rep(),what="Std")$coeff_fe[c("tau.ExpShip","nu.ExpShip"),1]

    estimates[i,]=est_alpha
    std[i,]=std_alpha
}

colnames(estimates)=c("tau.ExpShip","nu.ExpShip")
colnames(std)=c("tau.ExpShip","nu.ExpShip")

```

```{r}
#The 9th value in estimates is inconsistent, we drop it
estimates=estimates[-9,]
std=std[-9,]

mean_est=apply(X=estimates,MARGIN=2,FUN=mean)
mean_std=apply(X=na.omit(std),MARGIN=2,FUN=mean)
```