### 2025-03-17 R/simulation_study/examples/illustrative_trajectories.Rmd
try some new smooth functions omega

---
title: "Illustrative trajectories"
author: "Alexandre Delporte"
date: "2024-11-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this document, we simulate from constrained correlated velocity models with smooth parameters $omega$ and $tau$. The smooth functions will be parametric and the parameters will have direct interpretation in terms of movement.

# Set up

```{r set up}
cat("\014")                 # Clears the console

#seed for reproducibility 
seed=42

#parallel computing
library(foreach)
library(doParallel)

#for pipeline operator
library(dplyr)

#plots
library(ggplot2)
library(plotly)
library(htmlwidgets)
library(ggpubr)

#data managing
library(tidyr)

#progress bar
library(progress)

#for land polygons
library(sf)

#for spline estimation
library(mgcv)

#to get root of git repo
library(here)

#animated trajectories
library(gganimate)

#get functions to simulate trajectories
source(file.path(here("R","simulation_study","CVM_functions.R")))

```

# Define initial velocity and location for trajectories

```{r initial velocity and positions}
set.seed(seed)
# Initial velocity and location in fjords -------------

# Set the path to the directory containing the data
greenland_data_path <- here("Data","preprocessed_data","greenland")

#get the land and coastline geometries from the geojson file
land_border<-st_read(file.path(greenland_data_path,"updated_scoresby_sound_utm.shp"))

land_border <- st_transform(land_border, crs = "+init=EPSG:32626 +units=km")

v0=c(0,0)

#generate random initial points
x0=matrix(rep(NA,12),ncol=2)
colnames(x0)=c("x1","x2")

N=6 #number of trajectories
i=1
while (i<=N) {
  #choose location uniformly in the map
  x=c(runif(1,min=430,max=510),runif(1,min=7760,max=7900))
  p=nearest_shore_point(st_point(x),land_border)
  Dshore=(x[1]-p[1])^2+(x[2]-p[2])^2
  if (Dshore<1.5 && Dshore >0.3) {
    x0[i,]=x
    i=i+1
  }
}
```

# Define domains

```{r rectangular domain}

# Rectangular domain around the initial location ---------------------

x.max=max(x0[,1])+5
x.min=min(x0[,1])-5
y.max=max(x0[,2])+5
y.min=min(x0[,2])-5

large_rectangle_coords <- matrix(c(x.min, y.min, x.min, y.max, x.max, y.max, x.max, y.min, x.min,  y.min), ncol = 2, byrow = TRUE)
small_rectangle_coords <- matrix(c(x.min+1, y.min+1, x.min+1, y.max-1, x.max-1, y.max-1, x.max-1, y.min+1, x.min+1,  y.min+1), ncol = 2, byrow = TRUE)

# Create an sf object representing the rectangles
large_rectangle <- st_sfc(st_polygon(list(large_rectangle_coords)))
large_rectangle<-st_sf(geometry = large_rectangle)

small_rectangle <- st_sfc(st_polygon(list(small_rectangle_coords)))
small_rectangle<-st_sf(geometry = small_rectangle)

#change its crs
rect_border<- st_sym_difference(large_rectangle, small_rectangle)

```

```{r circular domain}

# Circular domain around the initial position -------
middle=c(mean(x0[,1]),mean(x0[,2]))

# Create circle geometries
large_circle <- st_buffer(st_sfc(st_point(middle)), dist = 55)
small_circle <- st_buffer(st_sfc(st_point(middle)), dist = 54)

# Convert circles to polygons
large_circle_polygon <- st_cast(large_circle, "POLYGON")
small_circle_polygon <- st_cast(small_circle, "POLYGON")

# Create sf objects
large_circle_sf <- st_sf(geometry = large_circle_polygon)
small_circle_sf <- st_sf(geometry = small_circle_polygon)

# Calculate the symmetric difference to get the border
circ_border <- st_sym_difference(large_circle_sf, small_circle_sf)

```

```{r circular domains with islands}

#### Circular domain with islands

# Create small circles borders within the big circle
circle1 <- st_buffer(st_sfc(st_point(middle-c(15,0))), dist = 10)

# Convert circles to polygons
circle_polygon1 <- st_cast(circle1, "POLYGON")

# Create sf objects
circle_sf1 <- st_sf(geometry = circle_polygon1)

# Create small circles borders within the big circle
circle2 <- st_buffer(st_sfc(st_point(middle-c(0,15))), dist = 10)

# Convert circles to polygons
circle_polygon2 <- st_cast(circle2, "POLYGON")

# Create sf objects
circle_sf2 <- st_sf(geometry = circle_polygon2)

circ_island_border <- st_union(st_union(circ_border, circle_sf1), circle_sf2)

```

```{r plot domains}

# Plot using ggplot2
ggplot() +
  # Rectangular border
  geom_sf(data = rect_border, fill = "lightblue", color = "black") +
  ggtitle("Rectangular Border") +
  theme_minimal()

ggplot() +
  # Circular border
  geom_sf(data = circ_border, fill = "lightgreen", color = "black") +
  ggtitle("Circular Border") +
  theme_minimal()

ggplot() +
  # Circular domain with islands
  geom_sf(data = circ_island_border, fill = "lightcoral", color = "black") +
  ggtitle("Circular Border with Islands") +
  theme_minimal()
```

# Define parametric functions for smooth parameters $\tau$

First define the function for a constant value of $\tau$. By default, we set this value to $2$ h.

```{r function for constant tau}

#constant tau
ftau_constant=function(cov_data,tau=2) {
  return (tau)
}

```

Then we define smooth functions of the angle $\theta$ to model higher persistence along the shoreline. The function need to have bumps around $\frac{\pi}{2}$ and $-\frac{\pi}{2}$. This type of function can be obtained in multiple ways, using hyperbolic tangent or gaussian bumps for instance. We give an example with hyperbolic tangent.

```{r tanh tau}

ftau_tanh=function(cov_data,alpha=1,tau0=0.5,tau1=4) {
  #cov_data : dataframe of covariate with column for angle theta
  #alpha : steepness of the bump
  #tau0 : base value of tau
  #tau1 : maximum value of tau
  epsilon=0.1
  theta=cov_data$theta
  tau=tau0+(tau1-tau0)/2/tanh(epsilon/alpha)*(tanh((theta+(pi/2+epsilon))/alpha)-tanh((theta+(pi/2-epsilon))/alpha))+
         (tau1-tau0)/2/tanh(epsilon/alpha)*(tanh((theta+-pi/2+epsilon)/alpha)-tanh((theta-pi/2-epsilon)/alpha))

  return (tau)
}

```

```{r plot tanh tau tau1}
# Plot curves for different values of tau1
tau1_values <- c(2,4)

cov_data=data.frame("theta"=seq(from=-pi+0.1,to=pi-0.1,length.out=100))

plot_data_tau1 <- data.frame(
  theta = rep(cov_data$theta, times = length(tau1_values)),
  tau1 = rep(tau1_values, each = length(cov_data$theta)))
plot_data_tau1$tau=unlist(lapply(tau1_values,function(tau1) {ftau_tanh(cov_data,tau1=tau1)}))
# Plot using ggplot
ggplot(plot_data_tau1, aes(x = theta, y = tau, color = as.factor(tau1))) +
  geom_line() +
 ylab(expression(tau)) + 
  xlab(expression(Theta)) + 
  geom_vline(xintercept = c(-pi/2, pi/2), linetype = "dashed", color = "black") + 
  annotate("text", x = -pi/2-0.3, y = 0.3, label = expression(-pi/2), color = "black", size = 6) + 
  annotate("text", x = pi/2+0.2, y = 0.3, label = expression(pi/2), color = "black", size = 6) + 
  scale_color_manual(values = c("blue", "red"), 
                     labels = c("tau1 = 2", "tau1 = 4"))+
  theme_minimal() +
  theme(
    axis.title = element_text(size = 18, face = "bold"),
    axis.text = element_text(size = 12, face = "bold"),
    axis.text.x = element_blank(),  # Remove x-axis text
    axis.ticks.x = element_blank(),  # Remove x-axis ticks
    axis.title.y = element_text(angle = 0)  # Rotate the y-axis label and adjust horizontal justification
  )

```

```{r plot tanh tau alpha}
# Plot curves for different values of alpha
alpha_values<-c(0.5,1)

plot_data_alpha <- data.frame(
  theta = rep(cov_data$theta, times = length(tau1_values)),
  alpha = rep(alpha_values, each = length(cov_data$theta)))
plot_data_alpha$tau=unlist(lapply(alpha_values,function(alpha) {ftau_tanh(cov_data,alpha=alpha)}))
# Plot using ggplot
ggplot(plot_data_alpha, aes(x = theta, y = tau, color = as.factor(alpha))) +
  geom_line() +
 ylab(expression(tau)) + 
  xlab(expression(Theta)) + 
  geom_vline(xintercept = c(-pi/2, pi/2), linetype = "dashed", color = "black") + 
  annotate("text", x = -pi/2-0.3, y = 0.3, label = expression(-pi/2), color = "black", size = 6) + 
  annotate("text", x = pi/2+0.2, y = 0.3, label = expression(pi/2), color = "black", size = 6) + 
  scale_color_manual(values = c("blue", "red"), 
                     labels = c("alpha = 0.5", "alpha = 1"))+
  theme_minimal() +
  theme(
    axis.title = element_text(size = 18, face = "bold"),
    axis.text = element_text(size = 12, face = "bold"),
    axis.text.x = element_blank(),  # Remove x-axis text
    axis.ticks.x = element_blank(),  # Remove x-axis ticks
    axis.title.y = element_text(angle = 0)  # Rotate the y-axis label and adjust horizontal justification
  )

```

However this parametrization is quite complex and we can do better and easier with gaussian bumps.

```{r define tau gaussian}

ftau_gaussian=function(cov_data,sigma_bump=0.3,sigma_dip=0.5,tau0=1,dtau_bump=1,dtau_dip=0.5) {
  #cov_data : dataframe of covariates with column for angle theta
  #sigma_bump : width of the bumps at -pi/2 and pi/2
  #sigma_dip : width of the dip at 0
  #tau0 : base value of tau
  #dtau_bump : height of the bump
  #dtau_dip : depth of the dip

  theta=cov_data$theta
  bump1 <- dtau_bump*exp(-((theta - pi/2)^2) / (2 * sigma_bump^2))
  bump2 <- dtau_bump*exp(-((theta + pi/2)^2) / (2 * sigma_bump^2))
  dip<- dtau_dip*exp(-((theta)^2) / (2 * sigma_dip^2))
  tau=tau0+bump1+bump2-dip

  return (tau)
}

```

```{r plot tau gaussian dtau_bump}
dtau_bump_values=c(1,3)

plot_data_dtau <- data.frame(
  theta = rep(cov_data$theta, times = length(dtau_bump_values)),
  dtau = rep(dtau_bump_values, each = length(cov_data$theta)))
plot_data_dtau$tau=unlist(lapply(dtau_bump_values,function(dtau) {ftau_gaussian(cov_data,dtau_bump=dtau)}))

#ggplot for tau
plot_tau_gaussian_dtau <- ggplot(data=plot_data_dtau) + 
  geom_line(aes(x = theta, y = tau,color = as.factor(dtau))) + 
  ylab(expression(tau)) + 
  xlab(expression(Theta)) + 
  geom_vline(xintercept = c(-pi/2, pi/2), linetype = "dashed", color = "black") + 
  annotate("text", x = -pi/2-0.3, y = 0.3, label = expression(-pi/2), color = "black", size = 6) + 
  annotate("text", x = pi/2+0.2, y = 0.3, label = expression(pi/2), color = "black", size = 6) + 
  scale_color_manual(values = c("blue", "red"), 
                     labels = c("dtau_bump = 1", "dtau_bump = 3"))+
  theme_minimal() +
  theme(
    axis.title = element_text(size = 18, face = "bold"),
    axis.text = element_text(size = 12, face = "bold"),
    axis.text.x = element_blank(),  # Remove x-axis text
    axis.ticks.x = element_blank(),  # Remove x-axis ticks
    axis.title.y = element_text(angle = 0)  # Rotate the y-axis label and adjust horizontal justification
  )

plot_tau_gaussian_dtau

```

```{r plot tau gaussian sigma_bump}
sigma_bump_values=c(0.2,0.4)

plot_data_sigma_bump <- data.frame(
  theta = rep(cov_data$theta, times = length(sigma_bump_values)),
  sigma = rep(sigma_bump_values, each = length(cov_data$theta)))
plot_data_sigma_bump$tau=unlist(lapply(sigma_bump_values,function(sigma) {ftau_gaussian(cov_data,sigma_bump=sigma)}))

#ggplot for tau
plot_tau_gaussian_sigma <- ggplot(data=plot_data_sigma_bump) + 
  geom_line(aes(x = theta, y = tau,color = as.factor(sigma))) + 
  ylab(expression(tau)) + 
  xlab(expression(Theta)) + 
  geom_vline(xintercept = c(-pi/2, pi/2), linetype = "dashed", color = "black") + 
  annotate("text", x = -pi/2-0.3, y = 0.3, label = expression(-pi/2), color = "black", size = 6) + 
  annotate("text", x = pi/2+0.2, y = 0.3, label = expression(pi/2), color = "black", size = 6) + 
  scale_color_manual(values = c("blue", "red"), 
                     labels = c("sigma_bump = 0.2", "sigma_bump = 0.4"))+
  theme_minimal() +
  theme(
    axis.title = element_text(size = 18, face = "bold"),
    axis.text = element_text(size = 12, face = "bold"),
    axis.text.x = element_blank(),  # Remove x-axis text
    axis.ticks.x = element_blank(),  # Remove x-axis ticks
    axis.title.y = element_text(angle = 0)  # Rotate the y-axis label and adjust horizontal justification
  )

plot_tau_gaussian_sigma
```

# Define parametric functions for smooth parameter $\omega$

The parameter $\omega$ will drive how the process rotates depending on the current distance to the boundary and the current angle. It is crucial to rotate fast enough and in the right direction to avoid hitting the boundary. Obviously, what we mean by "fast enough" and "right direction" depends on the shape of the boundary. What is fast enough for a smooth boundary might not be fast enough in a corner for instance. $\vert \omega \vert$ should increase fast as $\vert \theta \vert$ ranges through $\left[ \frac{\pi}{2},\pi \right]$. The sign of $\omega$ depends on the sign of $\theta$. For positive angles $\theta \in \left[ \frac{\pi}{2},\pi \right]$, we need the velocity to rotate in the inverse trigonometric sense. Since the rotation induced in the velocity during $\Delta$ hours is $-\omega \Delta$, this means that we need a positive $\omega$ to effectively rotate in the inverse trigonometric sense. Similarly, for negative $\theta \in \left[ -\pi,-\frac{\pi}{2} \right]$, we need a negative $\omega$.

This type of behavior can be obtained with several function.
We start by using the hyperbolic tangent function.

```{r parametric tanh function for omega}

fomega_tanh=function(cov_data,D0=0.5,omega0=30*pi/2,lambda=2,kappa=0.2) {
  Dshore=cov_data$BoundaryDistance
  theta=cov_data$BoundaryAngle
  coeff=exp(-kappa*(Dshore/D0)^2)
  omega=omega0/2*(tanh(lambda*(theta+pi/2-atanh(-0.9)/lambda))+tanh(lambda*(theta-(pi/2-atanh(-0.9)/lambda))))*coeff
  return(omega)
}
```

```{r plot omega tanh}

get_omega_plot<-function(fomega,Dshore,theta) {

  #grid of values
  grid<- as.data.frame(expand.grid(Dshore,theta))
  colnames(grid)=c("BoundaryDistance","BoundaryAngle")

  #smooth omega
  npoints=length(Dshore)
  omega=matrix(fomega(grid),npoints,npoints)

  # dynamic plots for omega 
  plot_omega <- plot_ly(type = "surface",
                           contours = list(z = list(show = TRUE, start = min(omega), end = max(omega), size = 5,color="black")),
                             x = ~theta,y = ~Dshore,z=~omega,
                           colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
  layout(title=paste("Omega"),
         scene=list(xaxis = list(title = "Theta",showgrid = F),
                    yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "omega")))

  return (plot_omega)
}

Dshore=seq(from=0.05,to=4,length.out=30)
theta=seq(from=-pi+0.1,to=pi-0.1,length.out=30)

get_omega_plot(fomega_tanh,Dshore,theta)

```

A similar smooth function may also be obtained with the inverse of a square root near $-\pi$ and $\pi$.

```{r define omega  sqrt}

fomega_sqrt <- function(cov_data, a=10*pi, p = 2,D0=0.5) {

  Dshore=cov_data$BoundaryDistance
  theta=cov_data$BoundaryAngle

  epsilon <- pmax(0, Dshore - D0)^2  # Damping term
  ifelse(
    abs(theta) <= pi / 2,
    0,  # Flat in the middle
    sign(theta) * a * ((abs(theta) - pi / 2)^p) /
      sqrt((pi - abs(theta)) + epsilon)
  )
}

get_omega_plot(fomega_sqrt,Dshore,theta)

```

These smooth functions $\omega$ will produce a SDE model that is constrained within a restricted domain of $\R^2$.
The trajectories avoid the shoreline, but are not attracted to it. To model a tendency to follow the shoreline, we need to modify 
a bit the smooth function to get a rotation of the velocity toward the shore when it is on the verge of leaving it. 
We can do this in several ways. First, we use a sin function.

```{r define and plot omega sin}
# Define the extended function f(theta)
fomega_sin<- function(cov_data, a =4*pi,b=10*pi,D0=0.8) {
  Dshore=cov_data$BoundaryDistance
  theta=cov_data$BoundaryAngle
  ifelse(abs(theta) < pi/2, 
         -a * sin(2*theta)*exp(-(Dshore/D0)^2),
         ifelse(theta>0,
               sign(theta)*b*abs(theta-pi/2)^2*exp(-(Dshore/D0)^2),
               sign(theta)*b*abs(theta+pi/2)^2*exp(-(Dshore/D0)^2)))
}

get_omega_plot(fomega_sin,Dshore,theta)

```

In these functions, there is no mid range effect of the shore on the drift of the process, which means
that when the process is far enough from the shore, it cannot be attracted back.
To resolve this, we add a gaussian kernel in the functions.

```{r}

fomega_gaussian= function(cov_data,b=1,D1=1.5,theta1=pi/2/sqrt(3),sigma_D=1,sigma_theta=0.6) {
  Dshore=cov_data$BoundaryDistance
  theta=cov_data$BoundaryAngle
  b*(exp(-1/2*(theta+theta1)^2/sigma_theta^2)-
           exp(-1/2*(theta-theta1)^2/(sigma_theta)^2))*exp(-1/2*(Dshore-D1)^2/sigma_D^2)
}

get_omega_plot(fomega_gaussian,Dshore,theta)

```

We define a product of sine and cosine with an additional gaussian attractive term

```{r define and plot omega sin}
# Define the extended function f(theta)
fomega_sin_cos<- function(cov_data,omega_0=10,D0=0.8,p=1,b=1,D1=1,theta1=pi/4,sigma_D=1,sigma_theta=0.5) {
  Dshore=cov_data$BoundaryDistance
  theta=cov_data$BoundaryAngle

  attraction <- fomega_gaussian(cov_data,b,D1,theta1,sigma_D,sigma_theta)

  if (p%%2==0) {
    repulsion <- -omega_0*sign(theta)*sin(theta/2)^p*cos(theta)*D0/Dshore*exp(-Dshore/D0)
  }
  else {
    repulsion <- -omega_0*sin(theta/2)^p*cos(theta)*D0/Dshore*exp(-Dshore/D0)
  }
  attraction+repulsion
}

get_omega_plot(fomega_sin_cos,Dshore,theta)

```

Now we use a polynomial with an additional gaussian kernel to model attraction toward the shore.

```{r}
fomega_poly=function(cov_data,a=0.2,b=5,D0=0.3,D1=1.5,sigma_D=1,sigma_theta=pi/8,
                     theta1=pi/2/sqrt(3),p=3) {
  Dshore=cov_data$BoundaryDistance
  theta=cov_data$BoundaryAngle

  if (p%%2==0) {
    repulsion<- a*sign(theta)*theta^p*(theta-pi/2)*(theta+pi/2)*D0/Dshore*exp(-Dshore/D0)
  }
  else {
    repulsion<- a*theta^p*(theta-pi/2)*(theta+pi/2)*D0/Dshore*exp(-Dshore/D0)
  }

  attraction<- fomega_gaussian(cov_data,b,D1,theta1,sigma_D,sigma_theta)

  return (repulsion+attraction)
}

get_omega_plot(fomega_poly,Dshore,theta)
```
```{r}
fomega_quadratic=function(cov_data,a=0.2,D0=0.3,D1=1.5) {
  Dshore=cov_data$BoundaryDistance
  theta=cov_data$BoundaryAngle

    return (a*(theta-pi/2)*(theta+pi/2)*D0/Dshore*exp(-Dshore/D0))
}

get_omega_plot(fomega_quadratic,Dshore,theta)
```

We can also geometrically compute what should be the angular veolcity to avoid land. And we get the expression of $\omega$ like this.

```{r}

fomega_minimal=function(cov_data,nu=4,D0=0.5,D1=1.5,a=1,b=10,sigma_D=0.5,sigma_theta=pi/8) {
  Dshore=cov_data$BoundaryDistance
  theta=cov_data$BoundaryAngle
  ifelse(abs(theta)>pi/2,a*sign(theta)*(pi/2-abs(theta))*nu*cos(theta)/Dshore*exp(-Dshore/D0),-sign(theta)*b*
    exp(-1/2*(theta-sign(theta)*pi/4)^2/sigma_theta^2)*exp(-Dshore/D1))
}

get_omega_plot(fomega_minimal,Dshore,theta)
```

We create and save the perpective plot for the paper.

```{r persp plot cubic parametric omega}

#Define function omega with adequate parameters
fomega_cubic<-function(cov_data) {
  fomega_poly(cov_data,a=2,b=5,D0=0.3,D1=0.8,sigma_D=0.3,sigma_theta=pi/8,
                     theta1=pi/2/sqrt(3),p=1)
}

# Perspective plots for omega

# Grid 
Dshore = seq(from = 0.05, to = 2, length.out = 30)
theta = seq(from = -pi + 0.1, to = pi - 0.1, length.out = 30)
grid <- as.data.frame(expand.grid(Dshore, theta))
colnames(grid) = c("BoundaryDistance", "BoundaryAngle")

# Smooth omega
z = matrix(fomega_cubic(grid), 30, 30)

# Create a function interpolating colors in the range of specified colors
jet.colors <- colorRampPalette( c("blue", "green") )
# Generate the desired number of colors from this palette
nbcol <- 100
color <- jet.colors(nbcol)
# Compute the z-value at the facet centres
ncz=ncol(z)
nrz=nrow(z)
zfacet <- z[-1, -1] + z[-1, -ncz] + z[-nrz, -1] + z[-nrz, -ncz]
# Recode facet z-values into color indices
facetcol <- cut(zfacet, nbcol)

# Open PNG device 
png(file.path("perspective_plots.png"), width = 1000, height = 500)

par(mfrow = c(1, 2), mar = c(3, 3, 3, 3))  #

# First plot
res=persp(Dshore, theta, z, 
      phi = 15,
      axes = FALSE,
      col = color[facetcol], 
      cex.lab = 1.5, 
      ticktype = "simple",
      xlim = c(min(Dshore), max(Dshore)),  # Explicit xlim
      ylim = c(min(theta), max(theta)))

#lines(trans3d(seq(0.05,2,length.out=5), y = rep(-pi+0.1,5), z = rep(min(z),5), pmat = res), col = "black")
lines(trans3d(rep(1,5), y = rep(-pi+0.1,5), z = seq(min(z)-20,min(z)+1030,5), pmat = res), col = "grey",lty=2)
lines(trans3d(rep(0.5,5), y = rep(-pi+0.1,5), z = seq(min(z)-20,min(z)+1030,5), pmat = res), col = "grey",lty=2)
lines(trans3d(rep(0.1,5), y = rep(-pi+0.1,5), z = seq(min(z)-20,min(z)+1030,5), pmat = res), col = "grey",lty=2)

text(trans3d(0.1, -pi + 0.05, min(z)+30, pmat = res), labels = "0.1", pos = 4, cex = 1.2, col = "black")  # Line 4
text(trans3d(0.5, -pi + 0.05, min(z)+30, pmat = res), labels = "0.5", pos = 4, cex = 1.2, col = "black")  # Line 2
text(trans3d(1, -pi + 0.05, min(z)+30, pmat = res), labels = "1", pos = 4, cex = 1.2, col = "black")  # Line 3

mtext(expression(D^{shore}), side = 1, line = -2, cex = 1.5)
mtext(expression(omega), side = 2, line = 1, cex = 1.5, las = 2)

# Second plot
persp(Dshore, theta, z, 
      phi = 15,theta=-90,
      axes = FALSE,
      col = color[facetcol], 
      cex.lab = 1.5,  #
      ticktype = "simple")
mtext(expression(Theta), side = 1, line = -2, cex = 1.5)
mtext(expression(omega), side = 2, line = 1, cex = 1.5, las = 2)

#lines(trans3d(seq(0.05,2,length.out=5), y = rep(-pi+0.1,5), z = rep(min(z),5), pmat = res), col = "black")
lines(trans3d(rep(0.5,5), y = rep(-pi+0.1,5), z = seq(min(z)-20,min(z)+1030,5), pmat = res), col = "grey",lty=2)
lines(trans3d(rep(1,5), y = rep(-pi+0.1,5), z = seq(min(z)-20,min(z)+1030,5), pmat = res), col = "grey",lty=2)
lines(trans3d(rep(1.5,5), y = rep(-pi+0.1,5), z = seq(min(z)-20,min(z)+1030,5), pmat = res), col = "grey",lty=2)

text(trans3d(0.5,-pi+0.05, min(z)+30, pmat = res), labels = expression(-pi/2), pos = 4, cex = 1.2, col = "black") 
text(trans3d(1, -pi + 0.05, min(z)+30, pmat = res), labels = "0", pos = 4, cex = 1.2, col = "black")  
text(trans3d(1.5, -pi + 0.05, min(z)+30, pmat = res), labels = expression(pi/2), pos = 4, cex = 1.2, col = "black")

dev.off()

```

We will keep the velocity parameter $\nu$ constant in our simulations.

```{r constant function for nu}

#constant nu
fnu_constant=function(cov_data,nu=4) {
  return (nu)
}
```

# Trajectories in a rectangular domain.

To start we inspect what the trajectories look like in the rectangular domain. 
We will try the different parametric function that we used before.

We simulate over 3 days with a constant time step equal to 10 seconds.

```{r set time steps for simulation}
TMAX=3*24
dt=1/360
times=seq(0,TMAX,by=dt)

```

```{r simulate with omega tanh in rectangle,cache=TRUE}
# CRCVM with tanh smooth omega

simulate_paralell<-function(ftau,fnu,fomega,v0=v0,x0=x0,times,border,n_cores=6,seed=42) {

  #start cluster
  cl <- makeCluster(n_cores) 
  registerDoParallel(cl)

  N<-nrow(x0)

  #paralell loop
  data=foreach (i=1:N,.combine='rbind',.export=c("sim_CRCVM","fomega_tanh","fomega_poly",
                "fomega_sin","fomega_sin_cos","fomega_sqrt","fomega_gaussian","fomega_minimal","fomega_quadratic"),
                .packages=c("progress","MASS","sf","mgcv","smoothSDE")) %dopar% {

    set.seed((seed-1)*N+i)

    res=sim_CRCVM(ftau=ftau,fomega=fomega,fnu=fnu,v0=v0,x0=x0[i,],
                      times=times,land=border,verbose=FALSE)

    data_sim=res$sim
    data_sim$ID=factor(rep(i,length(data_sim$y1)))
    data_shore=res$boundary[,c("p1","p2")]
    cbind(data_sim,data_shore)
  }

  #stop cluster
  stopCluster(cl)

  return (data)
}

data_rect_tanh<-simulate_paralell(ftau_constant,fnu_constant,fomega_tanh,v0,x0,times,rect_border)

```

```{r plot rect_tanh, message=FALSE}

get_trajectory_plot<-function(data,border,by=60) {

  #subsample to 10 minutes
  data_sub=data[seq(1,nrow(data),by=by),]

  #create plot
  plot<- ggplot() +
    geom_sf(data = border$geometry, fill = "grey", inherit.aes = FALSE) +
    coord_sf(datum = st_crs("+init=EPSG:32626 +units=km")) +
    geom_path(data = data_sub, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = data_sub, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = x0, aes(x = x1, y = x2), shape = 3, size = 2, col = "red") +
    theme_minimal() +
    xlab("x") + ylab("y") +
    theme(legend.position = "none")

  return (plot)
}

get_trajectory_plot(data_rect_tanh,rect_border)
```

The process is avoiding land by rotating when it approaches the boundary. 
Now, to have trajectories that tends to remain close to the boundary and move along it we can use the function \texttt{fomega_poly} with 
a non zero attractive gaussian term instead.

```{r simulate with omega cubic attractive in rectangle,cache=TRUE}

fomega_cubic_rect=function(cov_data) {

  fomega_poly(cov_data,a=1,b=5,D0=0.5,D1=1.5,sigma_D=0.5,sigma_theta=pi/8,p=1)

}

data_rect_cubic<-simulate_paralell(ftau_constant,fnu_constant,fomega_cubic_rect,v0,x0,times,rect_border)

```

```{r}

get_trajectory_plot(data_rect_cubic,rect_border)
```

We remove the attractive term in the function by taking parameter $b=0$.

```{r simulate with omega cubic repulsive in rectangle,cache=TRUE,message=FALSE}

fomega_cubic_repulsive_rect=function(cov_data) {

  fomega_poly(cov_data,a=1,b=0,D0=0.5,D1=1.5,sigma_D=0.5,sigma_theta=pi/8,p=1)

}

data_rect_cubic_repulsive<-simulate_paralell(ftau_constant,fnu_constant,fomega_cubic_repulsive_rect,
                                             v0,x0,times,rect_border)

```

```{r}
get_trajectory_plot(data_rect_cubic_repulsive,rect_border)
```

We can also increase the parameter $p$ in the polynomial omega function to get a steeper function
at the extremities.

```{r simulate with omega cubic attractive in rectangle,cache=TRUE,message=FALSE}

fomega_poly_rect=function(cov_data) {

  fomega_poly(cov_data,a=1,b=0,D0=0.5,D1=1.5,sigma_D=0.5,sigma_theta=pi/8,p=2)

}

data_rect_poly<-simulate_paralell(ftau_constant,fnu_constant,fomega_poly_rect,v0,x0,times,rect_border)

```

```{r}

get_trajectory_plot(data_rect_poly,rect_border)
```

We now try our \texttt{fome_sin_cos} function defined with a product of sin and cosine.

```{r simulate with omega cubic attractive in rectangle,cache=TRUE,message=FALSE}

fomega_sin_cos_rect=function(cov_data) {

  fomega_sin_cos(cov_data,omega_0=10,D0=0.8,p=1,b=0,D1=1,theta1=pi/4,sigma_D=1,sigma_theta=0.5)

}

data_rect_sin_cos<-simulate_paralell(ftau_constant,fnu_constant,fomega_sin_cos_rect,v0,x0,times,rect_border)

```

```{r}

get_trajectory_plot(data_rect_sin_cos,rect_border)
```

```{r simulate with omega minimal in rectangle,cache=TRUE,message=FALSE}

fomega_minimal_rect=function(cov_data) {

  fomega_minimal(cov_data,nu=4,D0=0.5,D1=1.5,a=3,b=10,sigma_D=0.5,sigma_theta=pi/8)
}

data_rect_minimal<-simulate_paralell(ftau_constant,fnu_constant,fomega_minimal_rect,v0,x0,times,rect_border)

```

```{r}

get_trajectory_plot(data_rect_minimal,rect_border)
```

Until now, we kept the parameter $\tau$ constant. We could emphasize the tendency to move along the boundary by using a smooth $\tau$ given by the functions \texttt{ftau_tanh} or \texttt{ftau_gaussian}.

Finally, we compare this to Hanks simulation technique for reflected SDE.

```{r simulate hanks reflected in rectangle,cache=TRUE}

# Generate samples ---------------
cores=detectCores()
cl <- makeCluster(6) #not to overload your computer
registerDoParallel(cl)

data_rect_hanks=foreach (i=1:N,.combine='rbind',.packages=c("progress","MASS","sf","mgcv")) %dopar% {

  set.seed((seed-1)*N+i)

  data_sim=sim_HANKSreflected(mu1=0,mu2=0,beta=0.5,sigma=8/sqrt(pi*2),x0[i,],v0,dt,TMAX,land=rect_border)
  data_sim$ID=factor(rep(i,length(data_sim$y1)))
  data_sim
}

#stop cluster
stopCluster(cl)

```

```{r}

get_trajectory_plot(data_rect_hanks,rect_border)
```

# Trajectories in a circular domain.

The values for the parameter of the smooth function $\omega$ and $\tau$ that produce constrained trajectories in a domain may not work for another domain.

```{r simulate with omega cubic tau gaussian in circle,cache=TRUE}

#change parameter values 
fomega_cubic_attractive_circ=function(cov_data) {

   fomega_poly(cov_data,a=30,D0=0.5,D1=3,sigma_theta=pi/10,sigma_D=1,b=5,p=1,theta1=pi/2)
}

data_circ_cubic_attractive <-simulate_paralell(ftau_constant,fnu_constant,fomega_cubic_attractive_circ,v0,x0,times,circ_border)

```

```{r}

get_trajectory_plot(data_circ_cubic_attractive,circ_border)
```

```{r simulate with omega cubic tau gaussian in circle,cache=TRUE}

#change parameter values 
fomega_cubic_repulsive_circ=function(cov_data) {

   fomega_poly(cov_data,a=3,D0=0.5,D1=1.5,sigma_theta=pi/10,sigma_D=0.5,b=0,theta=pi/2,p=1)
}

data_circ_cubic_repulsive <-simulate_paralell(ftau_constant,fnu_constant,fomega_cubic_repulsive_circ,v0,x0,times,circ_border)

```

```{r}

get_trajectory_plot(data_circ_cubic_repulsive,circ_border)
```

Finally, we compare it to Hanks simulation technique for reflected SDE.

```{r simulate hanks reflected in rectangle,cache=TRUE}

# Generate samples ---------------
cores=detectCores()
cl <- makeCluster(6) #not to overload your computer
registerDoParallel(cl)

data_circ_hanks=foreach (i=1:N,.combine='rbind',.packages=c("progress","MASS","sf","mgcv")) %dopar% {

  set.seed((seed-1)*N+i)

  data_sim=sim_HANKSreflected(mu1=0,mu2=0,beta=0.5,sigma=8/sqrt(pi*2),x0[i,],v0,dt,TMAX,land=circ_border)
  data_sim$ID=factor(rep(i,length(data_sim$y1)))
  data_sim
}

#stop cluster
stopCluster(cl)

```

```{r}
get_trajectory_plot(data_circ_hanks,circ_border)
```

# Trajectories in a circular domain with islands

```{r simulate with omega cubic and tau gaussian in circle with islands,cache=TRUE}

data_circ_island_cubic_attractive=simulate_paralell(ftau_constant,fnu_constant,fomega_cubic_attractive_circ,v0,x0,times,circ_island_border)

```

```{r}

get_trajectory_plot(data_circ_island_cubic_attractive,circ_island_border)
```

```{r simulate with omega cubic and tau gaussian in circle with islands,cache=TRUE}
data_circ_island_cubic_repulsive<- simulate_paralell(ftau_constant,fnu_constant,fomega_cubic_repulsive_circ,v0,x0,times,circ_island_border)

```

```{r}
get_trajectory_plot(data_circ_island_cubic_repulsive,circ_island_border)
```

# Trajectories in Scoresby Sound fjords

Now we simulate trajectories in the fjords.

```{r simulate with omega cubic and tau gaussian in circle with islands,cache=TRUE}
data_fjords_tanh<- simulate_paralell(ftau_constant,fnu_constant,fomega_tanh,v0,x0,times,land_border)
```

```{r}
get_trajectory_plot(data_fjords_tanh,land_border)
```

```{r define function to get animated trajectories,eval=FALSE}

animate_trajectory_plot<- function(data,border,by=60,anim_name="sim.gif") {

  #subsample 
  data_sub=data[seq(1,nrow(data),by=by),]

  # Basic ggplot setup
  plot <- ggplot() +
    geom_sf(data=border$geometry,fill="grey")+
    coord_sf(datum=st_crs("+init=EPSG:32626 +units=km"))+
    geom_path(data=data_sub, 
              aes(x = y1, y = y2,group=ID),size = 1, 
              lineend = "round",alpha=0.15) +     
    geom_point(data=data_sub,
               aes(x = y1, y = y2,shape=ID),size =4 ) +  
    labs(title = "Simulated trajectory in Scoresby Sound ", 
         subtitle = "Time : {frame_along} h", 
         x = "X Position", 
         y = "Y Position") +
    theme_minimal()

  anim <- plot + 
    transition_reveal(along = time) + 
    labs(subtitle = "Time: {round(frame_along,1)} hours")

  # Save the animation
  animate(anim, nframes = 100, fps = 5, width = 800, height = 600, 
          renderer = gifski_renderer(anim_name))
  }

```

```{r}
animate_trajectory_plot(data_fjords_tanh,land_border,by=60,anim_name="sim_fjords_tanh.gif")
```

```{r simulate with omega cubic in fjords,cache=TRUE}

#set parameters values
fomega_fjords_cubic_attractive<- function(cov_data){
 fomega_poly(cov_data,a=3,D0=0.5,D1=1,sigma_theta=pi/8,sigma_D=0.5,b=4,theta1=pi/2,p=1)
}

data_fjords_cubic_attractive<- simulate_paralell(ftau_constant,fnu_constant,fomega_fjords_cubic_attractive,v0,x0,times,land_border)
```

```{r}
get_trajectory_plot(data_fjords_cubic_attractive,land_border)
```

```{r}
animate_trajectory_plot(data_fjords_cubic_attractive,land_border,anim_name="sim_fjords_cubic_attractive.gif")
```

![Simulated Trajectories](sim_fjords_cubic_attractive.gif)

We now simulate trajectories without the attractive gaussian term in $\omega$.

```{r simulate with omega cubic repulsive in fjords,cache=TRUE}

fomega_cubic_fjords_repulsive=function(cov_data) {
  fomega_cubic(cov_data,a=1,D0=0.3,D1=0.8,sigma_theta=pi/8,sigma_D=0.3,b=0)
}

data_fjords_cubic_repulsive<- simulate_paralell(ftau_constant,fnu_constant,fomega_fjords_cubic_repulsive,v0,x0,times,land_border)
```

```{r}
get_trajectory_plot(data_fjords_cubic_repulsive,land_border)
```

```{r}
animate_trajectory_plot(data_fjords_cubic_repulsive,land_border,anim_name="sim_fjords_cubic_repulsive.gif")
```

![Simulated Trajectories](sim_fjords_cubic_repulsive.gif)

```{r simulate with omega quadratic in fjords,cache=TRUE}

fomega_fjords_quadratic=function(cov_data) {
  fomega_quadratic(cov_data,a=0.5,D0=3)
}

data_fjords_quadratic<- simulate_paralell(ftau_constant,fnu_constant,fomega_fjords_quadratic,v0,x0,times,land_border)
```

```{r}
get_trajectory_plot(data_fjords_quadratic,land_border)
```
```{r simulate with omega sin cos in fjords,cache=TRUE}

fomega_fjords_sin_cos=function(cov_data) {
  fomega_sin_cos(cov_data,omega_0=10,D0=0.8,p=1,b=0)
}

data_fjords_sin_cos<- simulate_paralell(ftau_constant,fnu_constant,fomega_fjords_sin_cos,v0,x0,times,land_border)
```

```{r}
get_trajectory_plot(data_fjords_sin_cos,land_border)
```

Finally, we compare it to Hanks simulation technique for reflected SDE.

```{r simulate hanks reflected in rectangle,cache=TRUE}
# Generate samples ---------------
cores=detectCores()
cl <- makeCluster(6) #not to overload your computer
registerDoParallel(cl)

data_fjords_hanks=foreach (i=1:N,.combine='rbind',.packages=c("progress","MASS","sf","mgcv")) %dopar% {

  set.seed((seed-1)*N+i)

  data_sim=sim_HANKSreflected(mu1=0,mu2=0,beta=0.5,sigma=8/sqrt(pi*2),x0[i,],v0,dt,TMAX,land=land_border)
  data_sim$ID=factor(rep(i,length(data_sim$y1)))
  data_sim
}

#stop cluster
stopCluster(cl)
```

```{r}
get_trajectory_plot(data_fjords_hanks,land_border)
```

```{r}
animate_trajectory_plot(data_fjords_hanks,land_border,anim_name="sim_fjords_hanks.gif")
```

```{r simulate with omega sin cos in fjords,cache=TRUE}

#set parameters values
fomega_fjords_sin_cos_attractive<- function(cov_data){
 fomega_sin_cos(cov_data,omega_0=20,D0=0.5,p=2,b=3,D1=1,theta1=pi/2/sqrt(3),sigma_D=1,sigma_theta=0.5)
}

data_fjords_sin_cos_attractive<- simulate_paralell(ftau_constant,fnu_constant,
                                                   fomega_fjords_sin_cos_attractive,v0,x0,times,land_border)
```

```{r}
get_trajectory_plot(data_fjords_sin_cos_attractive,land_border)
```
We try a higher order polynomial.

```{r simulate with omega cubic repulsive in fjords,cache=TRUE}

fomega_fjords_poly_attractive=function(cov_data) {
  fomega_poly(cov_data,a=1,D0=0.3,D1=1,sigma_theta=pi/8,sigma_D=1,b=3,p=3)
}

data_fjords_poly_attractive<- simulate_paralell(ftau_constant,fnu_constant,
                                                fomega_fjords_poly_attractive,v0,x0,times,land_border)
```

```{r, message=FALSE}
plot=get_trajectory_plot(data_fjords_poly_attractive,land_border)
```

```{r}
animate_trajectory_plot(data_fjords_cubic_repulsive,land_border,anim_name="sim_fjords_cubic_repulsive.gif")
```
# Create final plots for the paper

```{r zoom in fjords}

#Zoom on Fjords for the plot --------------------

x.min= min(c(data_sim_fjords_cubic_attractive_sub$y1,
             data_sim_fjords_hanks_sub$y1))-1
x.max = max(c(data_sim_fjords_cubic_attractive_sub$y1,
              data_sim_fjords_hanks_sub$y1))+1
y.min = min(c(data_sim_fjords_cubic_attractive_sub$y2,
              data_sim_fjords_hanks_sub$y2))-1
y.max = max(c(data_sim_fjords_cubic_attractive_sub$y2,
              data_sim_fjords_hanks_sub$y2))+1

bbox <- st_bbox(c(xmin=x.min, xmax = x.max, ymin =y.min, ymax = y.max), 
                crs = st_crs(land_border))

bbox_polygon <- st_as_sfc(st_bbox(bbox)
                          , crs = st_crs(land_border))

cropped_land<- land_border%>% st_intersection(bbox_polygon)

```

```{r get final combined plot }

# Create the plots ----------------------

# Ensure all geometries have the same CRS
st_crs(cropped_land$geometry) <- st_crs("+init=EPSG:32626 +units=km")
st_crs(rect_border$geometry) <- st_crs("+init=EPSG:32626 +units=km")
st_crs(circ_border$geometry) <- st_crs("+init=EPSG:32626 +units=km")

# Combine geometries into a single sf object
geometries <- bind_rows(
  st_as_sf(data.frame(geometry = cropped_land$geometry, domain = "Fjords")),
  st_as_sf(data.frame(geometry = rect_border$geometry, domain = "Rect")),
  st_as_sf(data.frame(geometry = circ_border$geometry, domain = "Circ"))
)

# Combine all data into one dataframe
data_combined <- bind_rows(
  data_sim_fjords_cubic_attractive_sub %>% mutate(domain = "Fjords", model = "shore-attractive CVM"),
  data_sim_fjords_hanks_sub %>% mutate(domain = "Fjords", model = "shore-reflected CVM"),
  data_sim_rect_cubic_attractive_sub %>% mutate(domain = "Rect", model = "shore-attractive CVM"),
  data_sim_rect_hanks_sub %>% mutate(domain = "Rect", model = "shore-reflected CVM"),
  data_sim_circ_cubic_attractive_sub %>% mutate(domain = "Circ", model = "shore-attractive CVM"),
  data_sim_circ_hanks_sub %>% mutate(domain = "Circ", model = "shore-reflected CVM")

)

# Function to create plots with individual scales
create_plot <- function(domain, model) {
  data_filtered <- data_combined %>% filter(domain == !!domain & model == !!model)

  ggplot() +
    geom_sf(data = geometries %>% filter(domain == !!domain), aes(geometry = geometry), fill = "grey", inherit.aes = FALSE) +
    coord_sf(datum = st_crs("+init=EPSG:32626 +units=km")) +
    geom_path(data = data_filtered, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = data_filtered, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = x0, aes(x = x1, y = x2), shape = 3, size = 2, col = "red") +
     scale_color_viridis_d(labels=c("A1","A2","A3","A4","A5","A6"))+
    theme_minimal() + 
    xlab("x") + ylab("y") +
    theme(legend.position = "none") +
  theme(
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 14),
    legend.text = element_text(size = 18),
    axis.title.y = element_text(angle = 0)  # Rotate the y-axis label and adjust horizontal justification
  )
}

# Get unique combinations of domain and model
unique_combinations <- expand.grid(model = unique(data_combined$model),domain = unique(data_combined$domain))

# Create a list of plots
plot_list <- lapply(seq_len(nrow(unique_combinations)), function(i) {
  domain <- unique_combinations$domain[i]
  model <- unique_combinations$model[i]
  create_plot(domain, model)
})

final_plot=ggarrange(plotlist=plot_list,
                     ncol=2, nrow=3, common.legend = TRUE, legend="bottom",
                     labels= c("shore-attractive CVM","shore-reflected CVM"))

ggsave("illustrative_trajectories.pdf", plot = final_plot, width = 14, height = 14)

```

```{r print final plot}
final_plot

```

```{r}

#plot only trajectory in fjords

data_combined_fjords=data_combined[data_combined$domain=="Fjords",]

# Create a list of plots
plot_list <- lapply(seq_len(length(unique(data_combined$model))), function(i) {
  model <- unique_combinations$model[i]
  create_plot(domain="Fjords",model)
})

final_plot_fjords=ggarrange(plotlist=plot_list,
                     ncol=1, nrow=2, common.legend = TRUE, legend="bottom",
                     labels= c("shore-attractive CVM","shore-reflected CVM"))

final_plot_fjords

ggsave("illustrative_trajectories_fjords.pdf", plot = final_plot_fjords, width = 14, height = 14)

```